### 2021-04-27 eco_footprint.Rmd
download ecological footprint data

---
title: "Ecological footprint"
author: "Juan Rocha"
date: "4/26/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(plotly)
library(patchwork)

theme_set(theme_light())
```

A short document to remind myself how to get ecological footpint data. The global footprint network has an API documented [here](https://data.footprintnetwork.org/#/api). Through their service one can download all the time series of the ecological foot print per country. The API uses `curl`, so one can do it from `bash` directly or using `Rcurl`.

Here is an example of the command I used on bash (with the key from email):

```{bash}
# curl -u juan:<api-key-here> http://api.footprintnetwork.org/v1/years -o years.json
```

That way I got the `types`, `countries` and `years` json files. Now I parse them into data frames to extract the relevant variables:

```{r}
countries <- jsonlite::read_json("data/countries.json", simplifyVector = TRUE) %>% 
    as_tibble()
years <- jsonlite::read_json("data/years.json", simplifyVector = TRUE) %>% 
    as_tibble()
types <- jsonlite::read_json("data/types.json", simplifyVector = TRUE) %>% 
    as_tibble()

```

```{bash}
#curl -u juan:<api-key-here>  "https://api.footprintnetwork.org/v1/data/all/[1961-2017]" -o "all_#1.json"

```

```{r}
test <- jsonlite::read_json("data/ecofootprint/all_2000.json", simplifyVector = TRUE) %>% 
    as_tibble()
```

```{r}
ecofp <- list()
fls <- list.files("data/ecofootprint/")
ecofp <- map(fls, function(x) {
    jsonlite::read_json(
        path = paste0("data/ecofootprint/", x),
        simplifyVector = TRUE)
    }) %>% 
    bind_rows()
```

```{r}
ecofp
```

```{r}
ecofp %>% 
    ggplot(aes(year, value)) +
    geom_line(aes(color = shortName), show.legend = FALSE) +
    facet_wrap(~record, scales = "free_y" ) +
    theme_light(base_size = 9)
```

### 2021-04-27 trilemma_data.R

@@ -91,4 +91,6 @@ foot %>% names()
 
 ### this file is only for the last year... it does not have time series
 ### I've emailed the footprint network to see if we can get access to the time series.
-foot 
+
+countries <- jsonlite::read_json("data/countries.json", simplifyVector = TRUE) %>% 
+    as_tibble()

### 2021-05-22 eco_footprint.Rmd
update missing values analysis

@@
 date: "4/26/2021"
-output: html_document
+output:
+  html_notebook:
+      theme:
+          code_font:
+            google: Fira Code
 ---
@@
 ecofp
+
+## share file with Tong and Maike:
+# write_csv(ecofp, file = "eco_footprint_processed.csv")
 ```

### 2021-05-22 missing_values.Rmd
update missing values analysis

---
title: "Missing values"
author: "Juan Rocha"
date: "May 2021"
output:
  html_notebook:
      toc: true
      toc_float: true
      highlight: tango
      code_folding: hide
      theme: 
          code_font:
            google: Fira Code
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(naniar)
library(plotly)
library(patchwork)

theme_set(theme_light())
```

## Ecological footprint

```{r}
ecofp <- list()
fls <- list.files("data/ecofootprint/")
ecofp <- map(fls, function(x) {
    jsonlite::read_json(
        path = paste0("data/ecofootprint/", x),
        simplifyVector = TRUE)
    }) %>% 
    bind_rows()

ecofp
```

```{r}
ecofp %>% 
    ggplot(aes(year, value)) +
    geom_line(aes(color = shortName), show.legend = FALSE) +
    facet_wrap(~record, scales = "free_y" ) +
    theme_light(base_size = 9)
```

```{r}
ecofp %>%
    filter(record == "EFConsPerCap") %>%
    mutate(missing = is.na(value)) %>% 
    ggplot(aes(year, countryName)) +
    geom_tile(aes(fill = value)) +
    scale_fill_viridis_c(option = "C") +
    theme(axis.text.y = element_text(size = 4), legend.position = "top")
```

```{r}
g1 <- ecofp %>%
    filter(record == "EFConsPerCap") %>%
    select(countryName, year, value) %>% 
    pivot_wider(id_col = countryName, names_from = year, values_from = value) %>% 
    pivot_longer(cols = 2:last_col(), names_to = "year", values_to = "value") %>% 
    mutate(missing = is.na(value), year = as.numeric(year)) %>% 
    group_by(year) %>% 
    summarise(n_nas  = sum(missing)) %>% 
    ggplot(aes(year, n_nas)) + 
    geom_col(aes(fill = (year>1995)), alpha = 0.5) + 
    coord_flip() +
    theme_light(base_size = 7) +
    theme(legend.position = c(0.8,0.8)) + labs(title = "Missing values per year")

g2 <- ecofp %>%
    filter(record == "EFConsPerCap", year > 1995) %>%
    select(countryName, year, value) %>% 
    pivot_wider(id_col = countryName, names_from = year, values_from = value) %>% 
    pivot_longer(cols = 2:last_col(), names_to = "year", values_to = "value") %>% 
    mutate(missing = is.na(value), year = as.numeric(year)) %>% 
    group_by(countryName) %>% 
    summarise(n_nas  = sum(missing)) %>% 
    filter(n_nas > 0) %>% 
    mutate(countryName = as_factor(countryName) %>% fct_reorder(., n_nas)) %>% 
    ggplot(aes(countryName, n_nas)) + 
    geom_col(aes(fill = (n_nas < 11)), alpha = 0.5) + coord_flip() +
    theme_light(base_size = 7) + 
    labs(title = "Missing values per country after 1995") +
    theme(legend.position = c(0.8, 0.1))

g1 + g2
```

An alternative is use the data after 1995 where `NA`s start decreasing but still give us a couple of decades worth of data to work with. Another is limiting to countries where the count of `NA`s is less than 10.

## Gini: Inequality

```{r}
inq <- readxl::read_xlsx(
    path = "~/Documents/Projects/DATA/WIID_World_Income_Inequality_DB/WIID_06MAY2020.xlsx")

inq %>% 
    #filter(resource == "Income (net/gross)") %>% 
    ggplot(aes(year, gini_reported)) +
    geom_line(aes(color = country), show.legend = FALSE) +
    facet_grid(region_un ~ resource_detailed ) +
    theme_light(base_size = 6)

```
Need to read the docs carefully, it seems many measurements per year and it's hard to tell what is the unit per time that we should use for the analysis.

```{r}
inq %>% 
    ggplot(aes(year, gini_reported)) +
    geom_line(aes(color = country), show.legend = FALSE) +
    facet_grid(region_un ~ resource)
```
It seems that `Income (net)` is the variable with most observations. Let's see:

```{r}
g <- inq %>% 
    filter(year > 1950, resource == "Income (gross)") %>% 
    ggplot(aes(year, gini_reported)) +
    geom_line(aes(color = country), size = 0.1, show.legend = FALSE) +
    facet_wrap(~resource_detailed)

g
```
```{r}
inq %>% 
    filter(year > 1950, resource_detailed == "Income, gross") %>% 
    ggplot(aes(gini_reported, palma)) +
    geom_miss_point(size = 1, alpha = 0.5) +
    theme(legend.position = c(0.1,0.8))
```

```{r}
inq %>% 
    filter(year > 1950, resource_detailed == "Income, gross") %>% 
    filter(country == "Australia")
```
Note there are multiple estimates per year-country because the database report different sources.

```{r}
inq %>% 
    filter(year > 1950, resource_detailed == "Income, gross") %>% 
    group_by(country, year) %>% 
    summarize(gini_mean = mean(gini_reported, na.rm = TRUE)) %>% 
    pivot_wider(id_cols = country, names_from = year, values_from = gini_mean) %>% 
    pivot_longer(2:last_col(), names_to = "year", values_to = "gini") %>% 
    mutate(year = as.numeric(year)) %>% 
    ggplot(aes(year, country)) +
    geom_tile(aes(fill = gini)) +
    scale_fill_viridis_c(option = "C") +
    theme(axis.text.y = element_text(size = 4), legend.position = "top")
```
```{r}
inq %>% 
    filter(!is.na(palma), year > 1990) %>% select(country, year, palma) %>% 
    group_by(country, year) %>% 
    summarize(palma_mean = mean(palma, na.rm = TRUE)) %>% 
    pivot_wider(id_cols = country, names_from = year, values_from = palma_mean) %>% 
    pivot_longer(2:last_col(), names_to = "year", values_to = "palma_mean") %>%
    mutate(year = as.numeric(year)) %>%
    ggplot(aes(year, country)) +
    geom_tile(aes(fill = palma_mean)) +
    scale_fill_viridis_c(option = "C") +
    theme(axis.text.y = element_text(size = 4), legend.position = "top")
```
To be fair with Gini, let's to the same over all Gini's reported. Even if Gini comes computed by different studies and in different currencies, the coefficient is always between 0-100, meaning it is comparable.

```{r}
inq %>% 
    filter(!is.na(gini_reported), year > 1990) %>% select(country, year, gini_reported) %>% 
    group_by(country, year) %>% 
    summarize(gini_mean = mean(gini_reported, na.rm = TRUE)) %>% 
    pivot_wider(id_cols = country, names_from = year, values_from = gini_mean) %>% 
    pivot_longer(2:last_col(), names_to = "year", values_to = "gini_mean") %>%
    mutate(year = as.numeric(year)) %>%
    ggplot(aes(year, country)) +
    geom_tile(aes(fill = gini_mean)) +
    scale_fill_viridis_c(option = "C") +
    theme(axis.text.y = element_text(size = 4), legend.position = "top")
```
So how many country / years can we use? Remember the data has `r inq$country %>% unique() %>% length()` countries

```{r}
inq %>% 
    filter(!is.na(gini_reported), year > 1990) %>% select(country, year, gini_reported) %>% 
    group_by(country, year) %>% 
    summarize(gini_mean = mean(gini_reported, na.rm = TRUE)) %>% 
    pivot_wider(id_cols = country, names_from = year, values_from = gini_mean) %>% 
    pivot_longer(2:last_col(), names_to = "year", values_to = "gini_mean") %>%
    mutate(year = as.numeric(year), missing = is.na(gini_mean)) %>% 
    group_by(year) %>% 
    summarise(n_nas = sum(missing), missingness = n_nas / 200)
```
We are dealing with ~50% of missing values on the best years, at best we can work with 100 countries. Which ones?

```{r}
inq %>% 
    filter(!is.na(gini_reported), year > 1990) %>% select(country, year, gini_reported) %>% 
    group_by(country, year) %>% 
    summarize(gini_mean = mean(gini_reported, na.rm = TRUE)) %>% 
    pivot_wider(id_cols = country, names_from = year, values_from = gini_mean) %>% 
    pivot_longer(2:last_col(), names_to = "year", values_to = "gini_mean") %>%
    mutate(year = as.numeric(year), missing = is.na(gini_mean)) %>% 
    group_by(country) %>% 
    summarise(n_nas = sum(missing), missingness = n_nas / (2018-1990)) %>% 
    arrange(missingness)
```

If we allow for 60% missing values, we get 80 countries, 30% NAs is ~60 countries. If needed one can impute `NA`s with a mean trend. But how much is that messing with the real trends? Double click on the legend to select particular countries:

```{r fig.width=7}
g <- inq %>% 
    filter(!is.na(gini_reported), year > 1990) %>% select(country, year, gini_reported) %>% 
    group_by(country, year) %>% 
    summarize(gini_mean = mean(gini_reported, na.rm = TRUE)) %>% 
    pivot_wider(id_cols = country, names_from = year, values_from = gini_mean) %>% 
    pivot_longer(2:last_col(), names_to = "year", values_to = "gini_mean") %>%
    mutate(year = as.numeric(year), missing = is.na(gini_mean)) %>% 
    ggplot(aes(year, gini_mean)) +
    geom_line(aes(color = country), show.legend = FALSE, size = 0.5) 
    # geom_miss_point(show.legend = FALSE)
ggplotly(g)
```

This justifies why the step-wise approach that Tong first proposed makes sense. If we restrict the analysis to good-data countries, we miss key actors like South Africa, the country with highest inequality in the world.

## GNI: Prosperity

```{r}
dat1 <- read_csv(
    "~/Documents/Projects/DATA/WorldBank/SDG_csv/SDGData.csv") %>% 
    janitor::clean_names()
```

```{r}
dat1 %>% 
    select(-x33) %>% 
    filter(str_detect(indicator_name, "GNI")) %>% 
    pivot_longer(cols = x1990:x2017, names_to = "year", values_to = "value") %>% 
    mutate(year = str_remove(year, "x")) %>% 
    mutate(year = as.numeric(year)) %>% 
    ggplot(aes(year, value)) +
    geom_line(aes(color = country_name), show.legend = FALSE) +
    #geom_miss_point(show.legend = FALSE) +
    facet_wrap(~indicator_name, scales = "free_y")
```

```{r}
## How many missing values? 10506, but who has more?
missingness <- dat1 %>% 
    select(-x33) %>% 
    filter(str_detect(indicator_name, "GNI")) %>% 
    pivot_longer(cols = x1990:x2017, names_to = "year", values_to = "value") %>% 
    mutate(year = str_remove(year, "x")) %>% 
    mutate(year = as.numeric(year)) %>% 
    mutate(ismissing = is.na(value)) %>% 
    group_by(country_code, indicator_name, country_name) %>% 
    summarize(missingness = sum(ismissing)/n())

missingness %>% 
    ggplot(aes(country_code, indicator_name)) +
    geom_tile(aes(fill = missingness)) + 
    scale_fill_viridis_c(option = "C") +
    theme(axis.text.x = element_text(angle =90), legend.position = "top")
```

Decide who to exclude and which variable to use

```{r}
missingness %>% 
    ungroup() %>% 
    group_by(indicator_name) %>% 
    summarize(x = mean(missingness)) %>% 
    arrange(x)
```

```{r}
missingness %>% 
    ungroup() %>% 
    group_by(country_name) %>% 
    summarize(x = mean(missingness)) %>% 
    arrange(desc(x)) 
```
It seems that `GNI per capita (current US$)` is the one with less missing values. Tong recomments using the one with Atlas method wich is close on missingness, only 18%. A threshold of 50% missigness per country will left out places like Somalia, Iraq, Costa Rica, Zambia, China or Chile. Perhaps 60% is tolerable?

### 2021-05-22 lit_review.R

@@ -1,0 +1,3 @@
+library(tidyverse)
+
+dat <- read_csv(file = "https://api.box.com/2.0/files/uoii9orvhkoijk1hzg4ahqwdydwhuxlb")

### 2021-05-22 trilemma_data.R

@@ -71,7 +71,7 @@ inq %>%
     #filter(resource == "Income (net/gross)") %>% 
     ggplot(aes(year, gini_reported)) +
     geom_line(aes(color = country), show.legend = FALSE) +
-    facet_grid(resource_detailed ~ region_un)
+    facet_grid(region_un ~ resource_detailed )
 
 ## Need to read the docs carefully, it seems many measurements per year and it's hard to tell what is the unit per time that we should use for the analysis.
 

### 2021-05-25 lit_review.Rmd
update notebooks for quaterly meeting

---
title: "Literature review: Inequality and the biosphere"
author: "Juan Rocha"
date: "May 2021"
output:
  html_notebook:
    toc: true
    toc_float: true
    highlight: tango
    code_folding: hide
    theme: 
      bootswatch: cosmo
      code_font:
        google: Fira Code
    selfcontained: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  cache = TRUE, warning = FALSE, message = FALSE, echo = TRUE,
  fig.width = 8, fig.height = 5
)

library(tidyverse)
library(tidytext)
library(tm)
library (topicmodels)
library(lda)
library(naniar)
library(plotly)
library(patchwork)

theme_set(theme_light(10))
```

Below I read the data and organize the data by most citations with the mock data (N=2000) that Emmy provided:
```{r}
dat <- read_csv(file = "data/April27_1-2000.csv") %>% 
    janitor::clean_names()
dat %>% 
    select(title, authors, cited_by, year, author_keywords, abbreviated_source_title) %>% 
    arrange(desc(cited_by))
```

## Topic model

The topic model below is a quick and rough analysis. A proper one would need a sensitivity analysis over a number of choices that I'm making here arbirtrarily such as the algorithm to use, number of topics to fit, etc. Works for brainstorming, do not trust the end result.

```{r}
## extra words
too_words <- tibble(
    word = c("paper", "study", "aim", "aims", "objective", "purpose")
)

dtm <- dat %>% 
    select(abstract, title, year) %>%
    filter(!is.na(abstract)) %>%
    unique() %>% # 2193
    unnest_tokens(word, abstract) %>% 
    filter(!is.na(word)) %>% ## Books don't have abstract so they get dropped (n = 8)
    anti_join(stop_words) %>% 
    anti_join(too_words) %>%
    filter(!str_detect(word, "[:digit:]")) %>%
    group_by(title) %>%
    count(word, sort = TRUE) %>%
    cast_dtm(document = title, term = word, value = n)

dtm

```

Now we have a document term matrix with `r dim(dtm)[1]` documents (abstracts) and `r dim(dtm)[2]` words, after cleaning for common words, stop words and numbers. I pre-computed the topic models outside this notebook because it takes some time and memory, so not suitable for computing at runing time. Here are some visualizatins of the preliminary results:

```{r}
load("data/topic_models_gibbs.RData")
k <- c(5,10,25,50,100)
df_topic_number <- tibble(
    topic_number = k,
    entropy = map_dbl (topicNumber.TM, function (x)
        mean(apply(posterior(x)$topics, 1, function (z) - sum(z * log(z)))) # maximize Entropy
    ),
    alpha = map_dbl(topicNumber.TM, slot, "alpha"),
    log_lik = map_dbl(topicNumber.TM, logLik) #,  #maximize loglik
    #perplexity = map_dbl(topicNumber.TM, perplexity) #minimize perplexity
)

df_stats <- tibble(
    model = names(lapply(tset.TM, logLik)),
    loglik = as.numeric(lapply(tset.TM, logLik)), #maximize loglik
    entropy = lapply (tset.TM, function (x) 
        mean(apply(posterior(x)$topics,
                   1, function (z) - sum(z * log(z))))) %>% as.numeric()#maximize ENTROPY

)

perp <-  lapply(tset.TM[c(1,2,4)], perplexity)
perp$Gibbs <- NA
# pretty names:
df_stats$model <- c("VEM alpha", "VEM fixed", "Gibbs", "CTM")

g1 <- df_stats %>%
    add_column(perplexity = as.numeric(perp[c(1,2,4,3)])) %>% #minimize perplexity
    pivot_longer(cols = 2:4, names_to = "measure", values_to = "value") %>%
    ggplot(aes(model, value)) + 
    geom_col() + 
    # scale_y_continuous(labels = scales::label_scientific) +
    facet_wrap(.~measure, scales = "free_y") +
    labs(x = "Algorithm", y = "Value", tag = "A") +
    theme_light(base_size = 8) + 
    theme(axis.text.x = element_text(size = 5))

g2 <- df_topic_number %>%
    # mutate(alpha_log = log10(alpha)) %>%
    pivot_longer(cols = 2:last_col(), names_to = "measure", values_to = "value") %>%
    # filter(measure != "alpha") %>%
    ggplot(aes(as.factor(topic_number), value)) +
    geom_col() + 
    # scale_y_continuous(labels = scales::label_scientific) +
    labs(x = "Number of topics", y = "Value", tag = "B") +
    facet_wrap(.~measure, scales = "free", ncol = 4, nrow = 1) +
    theme_light(base_size = 8)

g1/g2
```

The best algorithm should maximize entropy and the log-likelihood while reducing perplexity. The latter cannot be calculated for Gibbs sampling, but with the small dataset we have, it seems to be the best performing algorithm. The optimal number of topics is something between 25 and 50, but not higher than 100, where the log-likelihood starts to decay again. For now I will be using 25 topics and Gibbs sampling for the rest of the analysis.

The heatmap below shows how a few documents (on the rows) have a high content of certain topics (on columns).

```{r}
gplots::heatmap.2(
    posterior(topicNumber.TM[[3]])$topics, 
    key=T, trace="none", margins=c(5,5), 
    cexRow=0.7, cexCol=0.7, col="topo.colors", keysize=1, key.par = list(cex = 0.5),
    labRow = FALSE, key.title = NA, key.xlab = "probability", key.ylab = "count" )
```

And below are the topics. A topic is the (posterior) probability distributions of words being likely to be representing together certain hidden structures (topics) in the text. All words in the vocabulary (N = `r dim(dtm)[2]` words) have a probability of explaining a topic, here I'm only showing the top 10 words per topic:

```{r fig.width=8, fig.height=6}
df_topics25 <- tidy(topicNumber.TM[[3]], matrix = "beta")

g_25 <- df_topics25 %>%
    group_by(topic) %>%
    top_n(10, beta) %>% 
    ungroup() %>%
    arrange(topic, - beta) %>%
    mutate(term = reorder_within(term, beta, topic)) %>%
    ggplot(aes(term, beta)) +
    geom_col(aes(fill = as.factor(topic)),show.legend = FALSE) +
    coord_flip() + 
    scale_x_reordered() +
    labs(tag = "D", y = "Probability of word explaining the topic", x = "Word ranking") +
    facet_wrap(.~ topic, scales = "free_y", ncol = 5) +
    theme_light(base_size = 7) + 
    theme(axis.text.x = element_text(size = 5))

g_25
```

Now let's see how topics change over time:

```{r}
df_documents <- tidy(topicNumber.TM[[3]], matrix = "gamma")

papers_yr <- dat %>%
    #filter(year > 2000, year < 2021) %>%
    group_by(year) %>%
    tally() %>% rename(total_papers = n)

g_stack <-  df_documents %>% 
    left_join(
        dat %>% 
            select(abstract, title, year) %>%
            filter(!is.na(abstract)) %>%
            unique(),
        by = c("document" = "title") ) %>%
    mutate(year = as.numeric(year)) %>% 
    filter(year > 2000, year <2021) %>%
    group_by(topic, year) %>% 
    summarize(gamma_yr_tp = sum(gamma), .groups = "drop") %>% 
    group_by(year) %>%
    mutate(total = sum(gamma_yr_tp)) %>%
    ungroup() %>% 
    mutate(proportion = gamma_yr_tp / total,
           topic = as.factor(topic)) %>%
    ggplot(aes(x = year, y = proportion, group = topic)) +
    geom_area(aes(fill = topic, color = topic), 
              position = "stack", show.legend = FALSE, alpha = 0.5, size = 0.25) +
    labs(y = "Relative proportion of topics", x = "Year") +
    theme_classic(base_size = 10)

g_stack
```

If you're interested in learning more about the method or applications, you can learn more [here](https://arxiv.org/abs/2012.14312) or the references therein.

## Co-author networks

Below some code to construct the network. Note however that it's very slow. With over 4500 authors we have a network of co-authorship with over 232k links. The network alone is over 200Mb in RAM so at this stage any expensive computation is not worthed until we get the final data set (e.g. centrality, community detection, etc). I leave the graph here as demonstration, but in the explorative phase it is perhaps more useful tables with high impact papers per topic, and key authors per topic to follow up with the qualitative analysis.

```{r fig.width=8}
library(network)
authors <- dat %>% 
    select(authors, doi) %>% 
    mutate(authors = str_split(authors, pattern = ", ")) %>% 
    unnest(cols = authors) %>% 
    mutate(presence = 1)

mat <- authors %>% unique() %>% 
    pivot_wider(names_from = authors, values_from = presence, values_fill = 0) %>% 
    select(-doi) %>%
    as.matrix()

mat <- t(mat) %*% (mat)

df_authors <- as_tibble(mat, rownames = "author")

df_authors <- df_authors %>% 
    pivot_longer(2:last_col(), names_to = "co_author", values_to = "n_papers") %>% 
    filter(author != co_author, n_papers > 0)

net <- df_authors %>% 
    # filtering reduces from 232k links and >4.5k authors to 708 links shared by 233 authors
    filter(n_papers > 1) %>% 
    network(., directed = FALSE, matrix.type = "edgelist")

networkD3::simpleNetwork(
    Data= df_authors %>% filter(n_papers >1) ,
    Source = "author", Target = "co_author", 
    zoom = TRUE, linkColour = "grey", nodeColour = "blue"
)

## free memory
rm(mat, authors)

```

### 2021-05-25 missing_values.Rmd
update notebooks for quaterly meeting

@@
 output:
-  html_notebook:
-      toc: true
-      toc_float: true
+  html_document:
+    toc: yes
+    toc_float: yes
     highlight: tango
     code_folding: hide
+    df_print: paged
     theme: 
+      bootswatch: cosmo
+      code_font:
+        google: Fira Code
+    self_contained: true
+
+  html_notebook:
+    toc: yes
+    toc_float: yes
+    highlight: tango
+    code_folding: hide
+    df_print: paged
+    theme: 
+      bootswatch: cosmo
       code_font:
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_chunk$set(
+  cache = TRUE, warning = FALSE, message = FALSE, echo = TRUE,
+  fig.width = 8, fig.height = 5
+)
+
 library(tidyverse)
@@
 
-theme_set(theme_light())
+theme_set(theme_light(10))
 ```
@@
     geom_line(aes(color = shortName), show.legend = FALSE) +
-    facet_wrap(~record, scales = "free_y" ) +
-    theme_light(base_size = 9)
+    facet_wrap(~record, scales = "free_y" ) 
 ```
@@
 
-```{r}
+```{r fig.height=3, fig.width=8}
 g1 <- ecofp %>%
@@
     labs(title = "Missing values per country after 1995") +
-    theme(legend.position = c(0.8, 0.1))
+    theme(legend.position = c(0.8, 0.2))
 
@@
     path = "~/Documents/Projects/DATA/WIID_World_Income_Inequality_DB/WIID_06MAY2020.xlsx")
-
-inq %>% 
-    #filter(resource == "Income (net/gross)") %>% 
-    ggplot(aes(year, gini_reported)) +
-    geom_line(aes(color = country), show.legend = FALSE) +
-    facet_grid(region_un ~ resource_detailed ) +
-    theme_light(base_size = 6)
-
+inq
 ```
@@
 ```
-It seems that `Income (net)` is the variable with most observations. Let's see:
 
-```{r}
-g <- inq %>% 
-    filter(year > 1950, resource == "Income (gross)") %>% 
-    ggplot(aes(year, gini_reported)) +
-    geom_line(aes(color = country), size = 0.1, show.legend = FALSE) +
-    facet_wrap(~resource_detailed)
-
-g
-```
 ```{r}
@@
 
-```{r fig.width=7}
+```{r fig.width=8}
 g <- inq %>% 
@@
     geom_line(aes(color = country), show.legend = FALSE, size = 0.5) 
-    # geom_miss_point(show.legend = FALSE)
-ggplotly(g)
+ggplotly(g, tooltip = "country") 
 ```
@@
 Perhaps 60% is tolerable?
+
+## Analysis
+
+The bottle neck is centaintly our proxies of inequality, no matter what we use. Below a mock up example with some arbitrary choices.
+
+```{r}
+
+dat <- ecofp %>%
+  filter(record == "EFConsPerCap") %>%
+  select(countryName, isoa2, year, value) %>% 
+  rename(EFconsPerCap = value, country = countryName) %>% 
+  filter(year >= 1990)
+
+dat <- dat %>% 
+  left_join(
+    .,
+    inq %>% 
+      filter(!is.na(gini_reported), year >= 1990) %>% 
+      select(country, isoa2 = c2, country_code = c3, year, gini_reported) %>% 
+      group_by(country, year, isoa2, country_code) %>% 
+      summarize(gini_mean = mean(gini_reported, na.rm = TRUE))
+
+  )
+
+dat <- dat %>% 
+  left_join(
+    ., 
+    dat1 %>% 
+      select(-x33) %>% 
+      filter(indicator_name == "GNI per capita (current US$)") %>% 
+      pivot_longer(cols = x1990:x2017, names_to = "year", values_to = "value") %>%
+      mutate(year = str_remove(year, "x")) %>% 
+      mutate(year = as.numeric(year)) %>% 
+      mutate(ismissing = is.na(value)) %>% 
+      filter(year >= 1990) %>% 
+      rename(country = country_name, gni = value)
+  )
+
+dat
+
+```
+Each path is a country trajectory over time for the data we have available since 1990.
+Here are the three phase space graphs, one for each combination of variables in the trilemma.
+Ideally, we want to understand if being in a particular corner of the space force you to certain trajectories in other dimensions (e.g.
+being poor country forces you to treat the environment badly while you reach a minimum of equality or economic activity).
+The red dots are missing values and remind us why it is difficult to perform a full dynamic analysis, and many dots that are not missing are left alone, mining it will be hard to reconstruct time series for some countries.
+
+```{r}
+p1 <- dat %>% 
+  ggplot(aes(EFconsPerCap, gini_mean, group = country)) +
+  geom_miss_point(size = 0.5) +
+  geom_path(size = 0.2, color = "gray50") +
+  theme(legend.position = c(0.8, 0.8))
+
+p2<- dat %>% 
+  ggplot(aes(gni, gini_mean, group = country)) +
+  geom_miss_point(size = 0.5, show.legend = FALSE) +
+  geom_path(size = 0.2, color = "gray50") +
+  scale_x_log10()
+
+p3<- dat %>% 
+  ggplot(aes(gni, EFconsPerCap, group = country)) +
+  geom_miss_point(size = 0.5, show.legend = FALSE) +
+  geom_path(size = 0.2, color = "gray50") +
+  scale_x_log10()
+
+p1 + p2 + p3
+```
+
+```{r}
+# g <- dat %>% 
+#   ggplot(aes(EFconsPerCap, gini_mean, group = country)) +
+#   geom_point(aes(color = country), size = 0.5, alpha = 0.5) +
+#   geom_path(aes(color = country), size = 0.3) +
+#   theme_light(base_size = 7)
+# 
+# ggplotly(g, tooltip = c("country", "gini_mean", "EFconsPerCap"),
+#          width = 640, height = 480)
+```
+
+```{r}
+g <- dat %>% 
+  ggplot(aes(gni, gini_mean, group = country)) +
+  geom_point(aes(color = country), size = 0.5, alpha = 0.5) +
+  geom_path(aes(color = country), size = 0.3) +
+  scale_x_log10() +
+  theme_light(base_size = 7)
+
+ggplotly(g)
+```
+
+```{r}
+g <- dat %>% 
+  ggplot(aes(gni, EFconsPerCap, group = country)) +
+  geom_point(aes(color = country), size = 0.5, alpha = 0.5) +
+  geom_path(aes(color = country), size = 0.3) +
+  scale_x_log10()+
+  theme_light(base_size = 7)
+
+ggplotly(g)
+```

### 2021-05-25 lit_review.R

@@ -1,3 +1,167 @@
 library(tidyverse)
+library(tidytext)
+library(tm)
+library (topicmodels)
+library(lda)
+library(tictoc)
+library(network)
 
-dat <- read_csv(file = "https://api.box.com/2.0/files/uoii9orvhkoijk1hzg4ahqwdydwhuxlb")
+# not working, need manual download
+# dat <- read_csv(file = "https://stockholmuniversity.box.com/s/uoii9orvhkoijk1hzg4ahqwdydwhuxlb")
+# "https://api.box.com/2.0/files/uoii9orvhkoijk1hzg4ahqwdydwhuxlb"
+dat <- read_csv(file = "data/April27_1-2000.csv") %>% 
+    janitor::clean_names()
+
+## clean data: extra words
+too_words <- tibble(
+    word = c("paper", "study", "aim", "aims", "objective", "purpose")
+)
+
+dtm <- dat %>% 
+    select(abstract, title, year) %>%
+    filter(!is.na(abstract)) %>%
+    unique() %>% # 2193
+    unnest_tokens(word, abstract) %>% 
+    filter(!is.na(word)) %>% ## Books don't have abstract so they get dropped (n = 8)
+    anti_join(stop_words) %>% 
+    anti_join(too_words) %>%
+    filter(!str_detect(word, "[:digit:]")) %>%
+    group_by(title) %>%
+    count(word, sort = TRUE) %>%
+    cast_dtm(document = title, term = word, value = n)
+
+dtm
+
+## Choosing best algorithm
+SEED <- 2021
+k <- 10
+
+tset.TM <- list (
+    VEM0 = LDA(dtm, k=k, control = list ( seed = SEED)),
+    VEM_fixed= LDA(dtm, k=k, control= list (estimate.alpha = F, seed = SEED)),
+    Gibbs = LDA (dtm, k=k, method ="Gibbs", control = list (
+        seed = SEED, burnin= 1000, thin = 100, iter= 1000)),
+    CTM = CTM (dtm, k=k, control = list(seed = SEED, var= list (tol= 10^-4), em= list (tol = 10^-3))))
+
+sapply (tset.TM[1:3], slot, "alpha")
+
+#Finding number of topics
+k <- c(5,10,25,50,100)
+
+topicNumber.TM <- map(
+    .x = k,
+    .f = function(x) {
+        LDA(dtm, k = x, control= list (seed = SEED), method = "Gibbs")
+    })
+
+#Finding number of topics
+k <- c(5,10,25,50,100)
+
[…]

### 2021-10-27 Trilemma_clustering.Rmd
updating gitignore

---
title: "Trilemma"
author: "Juan Rocha"
date: "Updated `r Sys.Date()`"
output:
  html_notebook:
    toc: yes
    toc_float: yes
    highlight: tango
    code_folding: hide
    df_print: paged
    theme: 
      bootswatch: cosmo
      code_font:
        google: Fira Code

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  cache = TRUE, warning = FALSE, message = FALSE, echo = TRUE,
  fig.width = 8, fig.height = 5
)

library(tidyverse)
library(naniar)
library(plotly)
library(patchwork)

# time series
library(tsibble)
library(fable)
library(imputeTS)
library(slider)

# # clustering
# library(vegan)
# library(NbClust)
# library(clValid)
# library(mclust)

# theme_set(theme_light(base_size = 10))
```

## Data

```{r}
ecofp <- list()
fls <- list.files("data/ecofootprint/")
ecofp <- map(fls, function(x) {
    jsonlite::read_json(
        path = paste0("data/ecofootprint/", x),
        simplifyVector = TRUE)
    }) %>% 
    bind_rows()

ecofp
```

```{r}
inq <- readxl::read_xlsx(
    path = "~/Documents/Projects/DATA/WIID_World_Income_Inequality_DB/WIID_06MAY2020.xlsx")
inq
```

```{r}
dat1 <- read_csv(
    "~/Documents/Projects/DATA/WorldBank/SDG_csv/SDGData.csv") %>% 
    janitor::clean_names()
```

## Analysis

The bottle neck is certaintly our proxies of inequality, no matter what we use. Below a mock up example with some arbitrary choices.

```{r}

dat <- ecofp %>%
  filter(record == "EFConsPerCap") %>%
  select(countryName, isoa2, year, value) %>% 
  rename(EFconsPerCap = value, country = countryName) %>% 
  filter(year >= 1990)

dat <- dat %>% 
  left_join(
    .,
    inq %>% 
      filter(!is.na(gini_reported), year >= 1990) %>% 
      select(country, isoa2 = c2, country_code = c3, year, gini_reported) %>% 
      group_by(country, year, isoa2, country_code) %>% 
      summarize(gini_mean = mean(gini_reported, na.rm = TRUE))

  )

dat <- dat %>% 
  left_join(
    ., 
    dat1 %>% 
      select(-x33) %>% 
      filter(indicator_name == "GNI per capita (current US$)") %>% 
      pivot_longer(cols = x1990:x2017, names_to = "year", values_to = "value") %>%
      mutate(year = str_remove(year, "x")) %>% 
      mutate(year = as.numeric(year)) %>% 
      mutate(ismissing = is.na(value)) %>% 
      filter(year >= 1990) %>% 
      rename(country = country_name, gni = value)
  )

dat

```

Compute missing values for inequality:
```{r}
missingness <- dat %>% 
    mutate(ismissing = is.na(gini_mean)) %>% 
    group_by(country) %>% 
    summarize(missingness = sum(ismissing)/n()) %>% 
    arrange(missingness)

key_countries <- missingness %>% 
    filter(missingness < 0.3) %>% 
    pull(country)
```

```{r}
#library(tsibbledata)
df_dat <- dat %>% 
    select(-starts_with("indicator")) %>% 
    filter(country %in% key_countries) %>% 
    as_tsibble(key = country, index = year)

df_dat <- df_dat %>% 
    fill_gaps(.full = TRUE)

```

```{r}

df_dat <- df_dat %>% 
    group_by(country) %>% 
    mutate(
        gini = na_interpolation(gini_mean),
        gni2 = na_interpolation(gni),
        EFconsPerCap = na_interpolation(EFconsPerCap)
    ) 
```

```{r message=FALSE, warning=FALSE}
df_dat %>% 
    select(year, EFconsPerCap, gni2, gini) %>% 
    GGally::ggpairs(
        columns = 3:5 
    # lower = list(continuous = GGally::wrap("points", alpha = 0.5, size = 0.5)),
    # diag = list(continuous = GGally::wrap("densityDiag", alpha = 0.2)),
    # upper = list(continuous = GGally::wrap("cor", size = 2))
    ) + theme_light(base_size = 16)
```

```{r}
df_dat <- df_dat %>% 
    select(EFconsPerCap, gini, gni2) %>% 
    filter(!is.na(gni2))   # lost 2 more countries due to NAs

```

```{r}
df_dat %>% 
    pivot_longer(cols = c("EFconsPerCap", "gini", "gni2"),
                 names_to = "variable", values_to = "value") %>% 
    ggplot(aes(year, value)) +
    geom_line(aes(group = country)) +
    facet_wrap(~variable, ncol = 1, scales = "free") +
    theme_light(base_size = 15)
```

```{r}
p1 <- df_dat %>% 
  ggplot(aes(EFconsPerCap, gini, group = country)) +
  geom_point(size = 0.5) +
  geom_path(size = 0.2, aes(color = year)) +
  theme(legend.position = c(0.8, 0.8))

p2<- df_dat %>% 
  ggplot(aes(gni2, gini, group = country)) +
  geom_point(size = 0.5, show.legend = FALSE) +
  geom_path(size = 0.2, aes(color = year)) +
  scale_x_log10()

p3<- df_dat %>% 
  ggplot(aes(gni2, EFconsPerCap, group = country)) +
  geom_point(size = 0.5, show.legend = FALSE) +
  geom_path(size = 0.2, aes(color = year)) +
  scale_x_log10()

p1 + p2 + p3 + plot_layout(guides = "collect") & theme_light(base_size = 15)
```

```{r}
df_dat %>% 
    as_tibble() %>% 
    group_by(country) %>% 
    summarize(
        mean_ef = mean(EFconsPerCap),
        mean_gni = mean(gni2),
        mean_gini = mean(gini)
    ) %>% 
    arrange(desc(mean_ef))
```
### Experimenting with phase plane

```{r}
df_dat |> 
    mutate(
        d_EF = slide_dbl(EFconsPerCap, diff, .before = 1, .after = 0, .complete = TRUE),
        d_gini = slide_dbl(gini, diff, .before = 1, .after = 0, .complete = TRUE),
        d_gni2 = slide_dbl(gni2, diff, .before = 1, .after = 0, .complete = TRUE)
    ) |> 
    mutate(
        theta = atan(d_gini/d_EF),
        gamma = atan(d_gini/d_gni2)
    )
```

## Clustering

I'm not sure if the clustering requires one column per variable * year. Since really the unit of observation is country. Need to read upon it.

```{r}
m <- "ward.D2" # minimize the total within-cluster variance

clust_num <- NbClust::NbClust( 
    data = df_dat %>% as.data.frame() %>% 
        select(where(is.numeric), -year),
    #distance = "maximum",
    min.nc = 2, max.nc = 10, 
    method = m, 
    index = 'all')
```

It is a bit contested because the numeric indices favours 2 clusters with 6 of them, but 3 and 4 clusters are favoured by 5 metrics each. The Hubert index (which is graphical) favours 4 indexes as well, tieing up with the 2 cluster option.

```{r}
## Stability & Internal validation
stab <- clValid::clValid(
    obj = df_dat %>% as.data.frame() %>%
        select(where(is.numeric), -year) %>%
        as.matrix(),
    nClust = c(2:9),
    clMethods = c("hierarchical", "kmeans",  #"som", "model", "diana", "fanny",
                 "sota", "pam", "clara", "agnes"),
    validation = c('stability', "internal"),
    #metric = "manhattan",
    method = "ward",
    verbose = FALSE
) ## Hierarchical is the optimal method

clValid::optimalScores(stab)
```
The stability analysis also supports the choice of either 2 or 9 clusters. Partition around medioids does perform well with 9 clusters, but others are supporting the 2 clusters.

Here is an attempt with another package `kml` that take into consideration the fact that our data is longitudinal. However, looking into the documentation, it seems that it only handles a variable at the time. For the sake of testing:

```{r}
# library(kml)

test_dat <- df_dat %>% 
  select(country, gini, year) %>% 
  as_tibble() %>% 
  pivot_wider(names_from = year, values_from = gini, names_prefix = "yr_") %>% 
  as_data_frame() %>% 
  kml::cld(idAll = "country", timeInData = 2:29)

```
An error I don't understand.

### 2021-10-27 lit_review.Rmd
updating gitignore

@@
 ---
-title: "Literature review: Inequality and the biosphere"
+title: 'Literature review: Inequality and the biosphere'
 author: "Juan Rocha"
-date: "May 2021"
+date: "September 2021"
 output:
+  html_document:
+    df_print: paged
+    code_folding: hide
+    toc: true
+    toc_float:
+      collapsed: true
+      smooth_control: true
+    toc_depth: 3
+    fig_caption: true
+    fig_width: 8
+    fig_height: 5
+    highlight: tango
+    self_contained: true
+    theme: "paper"
+
   html_notebook:
-    toc: true
-    toc_float: true
+    toc: yes
+    toc_float: yes
     highlight: tango
@@
         google: Fira Code
-    selfcontained: true
 ---
@@
 library(tm)
-library (topicmodels)
+library(topicmodels)
 library(lda)
@@
 
-theme_set(theme_light(10))
+theme_set(theme_light())
 ```
 
-Below I read the data and organize the data by most citations with the mock data (N=2000) that Emmy provided:
-```{r}
-dat <- read_csv(file = "data/April27_1-2000.csv") %>% 
-    janitor::clean_names()
+Below I read the data and organize the data by most citations with the data that Emmy provided (N=86430 papers):
+```{r message = FALSE, warning=FALSE}
+
+fls <- list.files(path = "data/lit_review/")
+
+dat_xls <- paste0("data/lit_review/", str_subset(fls, ".xlsx")) %>% 
+    map(readxl::read_excel) %>% 
+    map(janitor::clean_names) %>% 
+    map(function(x) x %>% select(abstract, title, year, doi, cited_by) %>% as_tibble()) %>% 
+    map(function(x) x %>% mutate(across(.cols = everything(), .fns = as.character))) %>% 
+    bind_rows()
+
+dat_csv  <- paste0("data/lit_review/", str_subset(fls, ".csv")) %>% # head() %>% 
+    map(read.csv) %>% 
+    map(janitor::clean_names) %>% 
+    map(function(x) x %>% select(abstract, title, year, doi, cited_by) %>% as_tibble()) %>% 
+    map(function(x) x %>% mutate(across(.cols = everything(), .fns = as.character))) %>% 
+    bind_rows()
+
+dat <- bind_rows(dat_csv, dat_xls)
+rm(dat_csv, dat_xls)
+
 dat %>% 
-    select(title, authors, cited_by, year, author_keywords, abbreviated_source_title) %>% 
+  select(title, year, cited_by, doi, abstract) %>% 
   arrange(desc(cited_by))
@@
 
-The topic model below is a quick and rough analysis.
-A proper one would need a sensitivity analysis over a number of choices that I'm making here arbirtrarily such as the algorithm to use, number of topics to fit, etc.
-Works for brainstorming, do not trust the end result.
+I have done the topic modelling on the data Emmy downloaded. Here some prelim results:
 
 ```{r}
-## extra words
-too_words <- tibble(
-    word = c("paper", "study", "aim", "aims", "objective", "purpose")
-)
-
-dtm <- dat %>% 
-    select(abstract, title, year) %>%
-    filter(!is.na(abstract)) %>%
-    unique() %>% # 2193
-    unnest_tokens(word, abstract) %>% 
-    filter(!is.na(word)) %>% ## Books don't have abstract so they get dropped (n = 8)
-    anti_join(stop_words) %>% 
-    anti_join(too_words) %>%
-    filter(!str_detect(word, "[:digit:]")) %>%
-    group_by(title) %>%
-    count(word, sort = TRUE) %>%
-    cast_dtm(document = title, term = word, value = n)
-
-dtm
+## This does not compile to HTML due to memory, I compute stuff offline and add figs manually
+# load("data/210903_topic_models_gibbs.RData")
+# dtm
 
@@
 
-Now we have a document term matrix with `r dim(dtm)[1]` documents (abstracts) and `r dim(dtm)[2]` words, after cleaning for common words, stop words and numbers.
-I pre-computed the topic models outside this notebook because it takes some time and memory, so not suitable for computing at runing time.
-Here are some visualizatins of the preliminary results:
+Now we have a document term matrix with 73622 documents (abstracts) and 164098 words, after cleaning for common words, stop words and numbers.
+I pre-computed the topic models outside this notebook because it takes some time (days) and memory, so not suitable for computing at running time.
+Here are some visualizations of the preliminary results:
 
 ```{r}
-load("data/topic_models_gibbs.RData")
-k <- c(5,10,25,50,100)
-df_topic_number <- tibble(
-    topic_number = k,
-    entropy = map_dbl (topicNumber.TM, function (x)
-        mean(apply(posterior(x)$topics, 1, function (z) - sum(z * log(z)))) # maximize Entropy
-    ),
-    alpha = map_dbl(topicNumber.TM, slot, "alpha"),
-    log_lik = map_dbl(topicNumber.TM, logLik) #,  #maximize loglik
-    #perplexity = map_dbl(topicNumber.TM, perplexity) #minimize perplexity
-)
 
-df_stats <- tibble(
-    model = names(lapply(tset.TM, logLik)),
-    loglik = as.numeric(lapply(tset.TM, logLik)), #maximize loglik
-    entropy = lapply (tset.TM, function (x) 
-        mean(apply(posterior(x)$topics,
-                   1, function (z) - sum(z * log(z))))) %>% as.numeric()#maximize ENTROPY
+# k <- c(25,50,100,250,500)
+# df_topic_number <- tibble(
+#     topic_number = k,
+#     entropy = map_dbl (topicNumber.TM, function (x)
+#         mean(apply(posterior(x)$topics, 1, function (z) - sum(z * log(z)))) # maximize Entropy
+#     ),
+#     alpha = map_dbl(topicNumber.TM, slot, "alpha"),
+#     log_lik = map_dbl(topicNumber.TM, logLik) #,  #maximize loglik
+#     #perplexity = map_dbl(topicNumber.TM, perplexity) #minimize perplexity
+# )
+# 
+# df_stats <- tibble(
+#     model = names(lapply(tset.TM, logLik)),
+#     loglik = as.numeric(lapply(tset.TM, logLik)), #maximize loglik
+#     entropy = lapply (tset.TM, function (x) 
+#         mean(apply(posterior(x)$topics,
+#                    1, function (z) - sum(z * log(z))))) %>% as.numeric()#maximize ENTROPY
+#     
+# )
+# 
+# 
+# perp <-  lapply(tset.TM[c(1,2,4)], perplexity)
+# perp$Gibbs <- NA
+# # pretty names:
+# df_stats$model <- c("VEM alpha", "VEM fixed", "Gibbs", "CTM")
+# 
+# g1 <- df_stats %>%
+#     add_column(perplexity = as.numeric(perp[c(1,2,4,3)])) %>% #minimize perplexity
+#     pivot_longer(cols = 2:4, names_to = "measure", values_to = "value") %>%
+#     ggplot(aes(model, value)) + 
+#     geom_col() + 
+#     # scale_y_continuous(labels = scales::label_scientific) +
+#     facet_wrap(.~measure, scales = "free_y") +
+#     labs(x = "Algorithm", y = "Value", tag = "A") +
+#     theme_light(base_size = 15) + 
+#     theme(axis.text.x = element_text(size = 5))
+# 
+# g2 <- df_topic_number %>%
+#     # mutate(alpha_log = log10(alpha)) %>%
+#     pivot_longer(cols = 2:last_col(), names_to = "measure", values_to = "value") %>%
+#     # filter(measure != "alpha") %>%
+#     ggplot(aes(as.factor(topic_number), value)) +
+#     geom_col() + 
+#     # scale_y_continuous(labels = scales::label_scientific) +
+#     labs(x = "Number of topics", y = "Value", tag = "B") +
+#     facet_wrap(.~measure, scales = "free", ncol = 4, nrow = 1) +
+#     theme_light(base_size = 15)
+# 
+# g1/g2
 
-)
-
-perp <-  lapply(tset.TM[c(1,2,4)], perplexity)
-perp$Gibbs <- NA
-# pretty names:
-df_stats$model <- c("VEM alpha", "VEM fixed", "Gibbs", "CTM")
-
-g1 <- df_stats %>%
-    add_column(perplexity = as.numeric(perp[c(1,2,4,3)])) %>% #minimize perplexity
-    pivot_longer(cols = 2:4, names_to = "measure", values_to = "value") %>%
-    ggplot(aes(model, value)) + 
-    geom_col() + 
-    # scale_y_continuous(labels = scales::label_scientific) +
-    facet_wrap(.~measure, scales = "free_y") +
-    labs(x = "Algorithm", y = "Value", tag = "A") +
-    theme_light(base_size = 8) + 
-    theme(axis.text.x = element_text(size = 5))
-
-g2 <- df_topic_number %>%
-    # mutate(alpha_log = log10(alpha)) %>%
-    pivot_longer(cols = 2:last_col(), names_to = "measure", values_to = "value") %>%
-    # filter(measure != "alpha") %>%
-    ggplot(aes(as.factor(topic_number), value)) +
-    geom_col() + 
-    # scale_y_continuous(labels = scales::label_scientific) +
-    labs(x = "Number of topics", y = "Value", tag = "B") +
-    facet_wrap(.~measure, scales = "free", ncol = 4, nrow = 1) +
-    theme_light(base_size = 8)
-
-g1/g2
 ```
 
+![Algorithm selection](figures/topicstats.png)
+
 The best algorithm should maximize entropy and the log-likelihood while reducing perplexity.
-The latter cannot be calculated for Gibbs sampling, but with the small dataset we have, it seems to be the best performing algorithm.
-The optimal number of topics is something between 25 and 50, but not higher than 100, where the log-likelihood starts to decay again.
-For now I will be using 25 topics and Gibbs sampling for the rest of the analysis.
+The latter cannot be calculated for Gibbs sampling, but it seems to be the best performing algorithm.
+The optimal number of topics is 100, where the log-likelihood starts to decay again.
+For now I will be using 50 topics and Gibbs sampling for the rest of the analysis (reason: simpler visualizations).
 
-The heatmap below shows how a few documents (on the rows) have a high content of certain topics (on columns).
+A topic is the (posterior) probability distributions of words being likely to be representing together certain hidden structures (topics) in the text.
+All words in the vocabulary (N = 164098 words) have a probability of explaining a topic, here I'm only showing the top 10 words per topic:
 
-```{r}
-gplots::heatmap.2(
-    posterior(topicNumber.TM[[3]])$topics, 
-    key=T, trace="none", margins=c(5,5), 
-    cexRow=0.7, cexCol=0.7, col="topo.colors", keysize=1, key.par = list(cex = 0.5),
-    labRow = FALSE, key.title = NA, key.xlab = "probability", key.ylab = "count" )
+```{r fig.width=8, fig.height=10}
+# df_topics25 <- tidy(topicNumber.TM[[3]], matrix = "beta")
+# 
+# g_25 <- df_topics25 %>%
+#     group_by(topic) %>%
+#     top_n(25, beta) %>%
+#     ungroup() %>%
+#     arrange(topic, - beta) %>%
+#     mutate(term = reorder_within(term, beta, topic)) %>%
+#     ggplot(aes(term, beta)) +
+#     geom_col(aes(fill = as.factor(topic)),show.legend = FALSE) +
+#     coord_flip() +
+#     scale_x_reordered() +
+#     labs(y = "Probability of word explaining the topic", x = "Word ranking") +
+#     facet_wrap(.~ topic, scales = "free_y", ncol = 10) +
+#     theme_light(base_size = 5) +
+#     theme(axis.text.x = element_text(size = 5))
+# 
+# 
+# g_25
+# 
+# ggsave(
+#     filename = "topics_100.png",
+#     path = "figures/",
+#     plot = g_25,
+#     device = "png",
+#     width = 10, height = 12, units = "in",
+#     dpi = 400,
+#     bg = "white"
+# )
 ```
-
-And below are the topics.
-A topic is the (posterior) probability distributions of words being likely to be representing together certain hidden structures (topics) in the text.
-All words in the vocabulary (N = `r dim(dtm)[2]` words) have a probability of explaining a topic, here I'm only showing the top 10 words per topic:
-
-```{r fig.width=8, fig.height=6}
-df_topics25 <- tidy(topicNumber.TM[[3]], matrix = "beta")
-
-g_25 <- df_topics25 %>%
-    group_by(topic) %>%
-    top_n(10, beta) %>% 
-    ungroup() %>%
-    arrange(topic, - beta) %>%
-    mutate(term = reorder_within(term, beta, topic)) %>%
-    ggplot(aes(term, beta)) +
-    geom_col(aes(fill = as.factor(topic)),show.legend = FALSE) +
-    coord_flip() + 
-    scale_x_reordered() +
-    labs(tag = "D", y = "Probability of word explaining the topic", x = "Word ranking") +
-    facet_wrap(.~ topic, scales = "free_y", ncol = 5) +
-    theme_light(base_size = 7) + 
-    theme(axis.text.x = element_text(size = 5))
-
-g_25
-```
-
+![Topics](figures/topics_50.png)
 Now let's see how topics change over time:
@@
 ```{r}
-df_documents <- tidy(topicNumber.TM[[3]], matrix = "gamma")
-
-papers_yr <- dat %>%
-    #filter(year > 2000, year < 2021) %>%
-    group_by(year) %>%
-    tally() %>% rename(total_papers = n)
-
-g_stack <-  df_documents %>% 
-    left_join(
-        dat %>% 
-            select(abstract, title, year) %>%
-            filter(!is.na(abstract)) %>%
-            unique(),
-        by = c("document" = "title") ) %>%
-    mutate(year = as.numeric(year)) %>% 
-    filter(year > 2000, year <2021) %>%
-    group_by(topic, year) %>% 
-    summarize(gamma_yr_tp = sum(gamma), .groups = "drop") %>% 
-    group_by(year) %>%
-    mutate(total = sum(gamma_yr_tp)) %>%
-    ungroup() %>% 
-    mutate(proportion = gamma_yr_tp / total,
-           topic = as.factor(topic)) %>%
-    ggplot(aes(x = year, y = proportion, group = topic)) +
-    geom_area(aes(fill = topic, color = topic), 
-              position = "stack", show.legend = FALSE, alpha = 0.5, size = 0.25) +
-    labs(y = "Relative proportion of topics", x = "Year") +
-    theme_classic(base_size = 10)
-
-g_stack
+# df_documents <- tidy(topicNumber.TM[[2]], matrix = "gamma")
+# 
+# papers_yr <- dat %>%
+#     #filter(year > 2000, year < 2021) %>%
+#     group_by(year) %>%
+#     tally() %>% rename(total_papers = n)
+# 
+# 
+# g_stack <-  df_documents %>% 
+#     left_join(
+#         dat %>% 
+#             select(abstract, title, year) %>%
+#             filter(!is.na(abstract)) %>%
+#             unique(),
+#         by = c("document" = "title") ) %>%
+#     mutate(year = as.numeric(year)) %>% 
+#     filter(year > 2000, year <2021) %>%
+#     group_by(topic, year) %>% 
+#     summarize(gamma_yr_tp = sum(gamma), .groups = "drop") %>% 
+#     group_by(year) %>%
+#     mutate(total = sum(gamma_yr_tp)) %>%
+#     ungroup() %>% 
+#     mutate(proportion = gamma_yr_tp / total,
+#            topic = as.factor(topic)) %>%
+#     ggplot(aes(x = year, y = proportion, group = topic)) +
+#     geom_area(aes(fill = topic, color = topic), 
+#               position = "stack", show.legend = FALSE, alpha = 0.5, size = 0.25) +
+#     labs(y = "Relative proportion of topics", x = "Year") +
+#     theme_classic(base_size = 15) 
+# 
+# g_stack
+# 
+# ggsave(
+#     filename = "topics_time.png",
+#     path = "figures/",
+#     plot = g_stack,
+#     device = "png",
+#     width = 4, height = 3, units = "in",
+#     dpi = 400,
+#     bg = "white"
+# )
 ```
-
+![Topics over time](figures/topics_time.png)
+There is no particular dominance of a topic over another, each topic has received relatively similar attention in academic productivity.
 If you're interested in learning more about the method or applications, you can learn more [here](https://arxiv.org/abs/2012.14312) or the references therein.
@@
 ## Co-author networks
+
+**This is based on the old analysis with 2000 papers, please ignore for now**
 
@@
 ```{r fig.width=8}
-library(network)
-authors <- dat %>% 
-    select(authors, doi) %>% 
-    mutate(authors = str_split(authors, pattern = ", ")) %>% 
-    unnest(cols = authors) %>% 
-    mutate(presence = 1)
-
-mat <- authors %>% unique() %>% 
-    pivot_wider(names_from = authors, values_from = presence, values_fill = 0) %>% 
-    select(-doi) %>%
-    as.matrix()
-
-mat <- t(mat) %*% (mat)
-
-df_authors <- as_tibble(mat, rownames = "author")
-
-df_authors <- df_authors %>% 
-    pivot_longer(2:last_col(), names_to = "co_author", values_to = "n_papers") %>% 
-    filter(author != co_author, n_papers > 0)
-
-net <- df_authors %>% 
-    # filtering reduces from 232k links and >4.5k authors to 708 links shared by 233 authors
-    filter(n_papers > 1) %>% 
-    network(., directed = FALSE, matrix.type = "edgelist")
-
-networkD3::simpleNetwork(
-    Data= df_authors %>% filter(n_papers >1) ,
-    Source = "author", Target = "co_author", 
-    zoom = TRUE, linkColour = "grey", nodeColour = "blue"
-)
-
-## free memory
-rm(mat, authors)
+# library(network)
+# authors <- dat %>% 
+#     select(authors, doi) %>% 
+#     mutate(authors = str_split(authors, pattern = ", ")) %>% 
+#     unnest(cols = authors) %>% 
+#     mutate(presence = 1)
+# 
+# mat <- authors %>% unique() %>% 
+#     pivot_wider(names_from = authors, values_from = presence, values_fill = 0) %>% 
+#     select(-doi) %>%
+#     as.matrix()
+# 
+# mat <- t(mat) %*% (mat)
+# 
+# 
+# df_authors <- as_tibble(mat, rownames = "author") 
+# 
+# df_authors <- df_authors %>% 
+#     pivot_longer(2:last_col(), names_to = "co_author", values_to = "n_papers") %>% 
+#     filter(author != co_author, n_papers > 0)
+# 
+# net <- df_authors %>% 
+#     # filtering reduces from 232k links and >4.5k authors to 708 links shared by 233 authors
+#     filter(n_papers > 1) %>% 
+#     network(., directed = FALSE, matrix.type = "edgelist")
+# 
+# networkD3::simpleNetwork(
+#     Data= df_authors %>% filter(n_papers >1) ,
+#     Source = "author", Target = "co_author", 
+#     zoom = TRUE, linkColour = "grey", nodeColour = "blue"
+# )
+# 
+# ## free memory
+# rm(mat, authors)
 

### 2021-10-27 missing_values.Rmd
updating gitignore

@@
 output:
+  html_notebook:
+    toc: yes
+    toc_float: yes
+    highlight: tango
+    code_folding: hide
+    df_print: paged
+    theme: 
+      bootswatch: cosmo
+      code_font:
+        google: Fira Code
   html_document:
+
     toc: yes
@@
     self_contained: true
-
-  html_notebook:
-    toc: yes
-    toc_float: yes
-    highlight: tango
-    code_folding: hide
-    df_print: paged
-    theme: 
-      bootswatch: cosmo
-      code_font:
-        google: Fira Code
 ---
@@
     ggplot(aes(year, gini_mean)) +
-    geom_line(aes(color = country), show.legend = FALSE, size = 0.5) 
+    geom_line(aes(color = country), show.legend = FALSE, size = 0.5) +
+  theme_light(base_size = 7)
 ggplotly(g, tooltip = "country") 

### 2021-10-27 fig_covid_paper.R

@@ -1,0 +1,51 @@
+library(tidyverse)
+library(patchwork)
+
+dat <- readxl::read_xlsx("data/international_vaccinations_0910.xlsx", sheet = 1) |> 
+    janitor::clean_names()
+
+inq <- readxl::read_xlsx(
+    path = "~/Documents/Projects/DATA/WIID_World_Income_Inequality_DB/WIID_06MAY2020.xlsx")
+
+
+dat <- dat |> left_join(
+    inq %>% 
+        filter(!is.na(gini_reported), year >= 1990) %>% 
+        select(country, isoa2 = c2, country_code = c3, year, gini_reported, bottom40) %>% 
+        group_by(country, isoa2, country_code) %>% 
+        summarize(gini_mean = mean(gini_reported, na.rm = TRUE),
+                  bottom40 = mean(bottom40, na.rm = TRUE)) 
+    
+)
+
+
+dat |> skimr::skim()
+
+g1 <- dat |> 
+    filter(!is.na(continent)) |> 
+    ggplot(aes(bottom40, vaccinated_per100)) +
+    geom_point(aes(color = continent, size = gdp_per_capita_int), alpha = 0.5) +
+    labs(tag = "A", y = "Fully vaccinated per 100 people", x = "Mean bottom 40% (1990-2018)") +
+    scale_color_brewer("Continent", palette = "Set1") +
+    scale_size("GDP per capita", range = c(0.5,3)) +
+    theme_light(base_size = 6) + theme(legend.key.size = unit(0.25, "cm"))
+
+g2 <- dat |> 
+    filter(!is.na(continent)) |> 
+    ggplot(aes(gini_mean, vaccinated_per100)) +
+    geom_point(aes(color = continent, size = gdp_per_capita_int), alpha = 0.5)+
+    labs(tag = "B", y = "Fully vaccinated per 100 people", x = "Mean GINI (1990-2018)") +
+    scale_color_brewer("Continent", palette = "Set1") +
+    scale_size("GDP per capita", range = c(0.5,3)) +
+    theme_light(base_size = 6) + theme(legend.key.size = unit(0.25, "cm"))
+
+
+g1 + g2 + plot_layout(guides = "collect") 
+
+
+ggsave(
+    filename = "figures/vaccination_ineq.png",
+    device = "png",
+    width = 6, height = 2.5,
+    dpi = 300, bg = "white"
+)

### 2021-10-27 lit_review.R

@@ -5,65 +5,101 @@ library (topicmodels)
 library(lda)
 library(tictoc)
 library(network)
+library(patchwork)
 
 # not working, need manual download
 # dat <- read_csv(file = "https://stockholmuniversity.box.com/s/uoii9orvhkoijk1hzg4ahqwdydwhuxlb")
 # "https://api.box.com/2.0/files/uoii9orvhkoijk1hzg4ahqwdydwhuxlb"
-dat <- read_csv(file = "data/April27_1-2000.csv") %>% 
-    janitor::clean_names()
+# dat <- read_csv(file = "data/April27_1-2000.csv") %>% 
+#     janitor::clean_names()
+
+fls <- list.files(path = "data/lit_review/")
+
+dat_xls <- paste0("data/lit_review/", str_subset(fls, ".xlsx")) %>% 
+    map(readxl::read_excel) %>% 
+    map(janitor::clean_names) %>% 
+    map(function(x) x %>% select(abstract, title, year, doi) %>% as_tibble()) %>% 
+    map(function(x) x %>% mutate(across(.cols = everything(), .fns = as.character))) %>% 
+    bind_rows()
+
+dat_csv  <- paste0("data/lit_review/", str_subset(fls, ".csv")) %>% # head() %>% 
+    map(read.csv) %>% 
+    map(janitor::clean_names) %>% 
+    map(function(x) x %>% select(abstract, title, year, doi) %>% as_tibble()) %>% 
+    map(function(x) x %>% mutate(across(.cols = everything(), .fns = as.character))) %>% 
+    bind_rows()
+
+dat <- bind_rows(dat_csv, dat_xls)
+rm(dat_csv, dat_xls)
 
 ## clean data: extra words
 too_words <- tibble(
-    word = c("paper", "study", "aim", "aims", "objective", "purpose")
+    word = c("john", "wiley", "sons", "springer", "verlag", "b.v", "abstract", "press", "reserved", "rights", "author", "taylor", "francis", "elsevier", "i.e.", "e.g.", "publisher", "publishers", "published", "publishing", "ii", "iv", "mdpi", "copyright", "journal", "auhors", "blackwell", "oxford", "cambridge", "publisher", "university", "book", "volume", "gmbh")
 )
 
 dtm <- dat %>% 
     select(abstract, title, year) %>%
     filter(!is.na(abstract)) %>%
-    unique() %>% # 2193
+    unique() %>% #
[…]

### 2022-02-08 Trilemma_clustering.Rmd
back up trilemma clustering analysis

@@
         google: Fira Code
-
+editor_options: 
+  chunk_output_type: inline
 ---
@@
 
-# theme_set(theme_light(base_size = 10))
+#theme_set(theme_light(base_size = 12))
 ```
@@
     janitor::clean_names()
+
+missingness <- dat1 %>% 
+    select(-last_col()) %>% 
+    filter(str_detect(indicator_name, "GNI")) %>% 
+    pivot_longer(cols = x1990:x2019, names_to = "year", values_to = "value") %>% 
+    mutate(year = str_remove(year, "x")) %>% 
+    mutate(year = as.numeric(year)) %>% 
+    mutate(ismissing = is.na(value)) %>% 
+    group_by(country_code, indicator_name, country_name) %>% 
+    summarize(missingness = sum(ismissing)/n())
+
+missingness%>% 
+    ggplot(aes(country_code, indicator_name)) +
+    geom_tile(aes(fill = missingness)) + 
+    scale_fill_viridis_c(option = "C") +
+    theme(axis.text.x = element_text(angle =90), legend.position = "top")
 ```
@@
     dat1 %>% 
-      select(-x33) %>% 
-      filter(indicator_name == "GNI per capita (current US$)") %>% 
-      pivot_longer(cols = x1990:x2017, names_to = "year", values_to = "value") %>%
+      select(-last_col()) %>% 
+      filter(indicator_name == "GNI per capita (US$)") %>% 
+      pivot_longer(cols = x1990:x2019, names_to = "year", values_to = "value") %>%
       mutate(year = str_remove(year, "x")) %>% 
@@
     # upper = list(continuous = GGally::wrap("cor", size = 2))
-    ) + theme_light(base_size = 16)
+    ) + theme_light(base_size = 20)
 ```
@@
 
-p1 + p2 + p3 + plot_layout(guides = "collect") & theme_light(base_size = 15)
+p1 + p2 + p3 + plot_layout(guides = "collect") & theme_light(base_size = 18)
 ```
@@
         mean_gni = mean(gni2),
-        mean_gini = mean(gini)
+        mean_gini = mean(gini),
+        var_ef = var(EFconsPerCap),
+        var_gni = var(gni2),
+        var_gini = var(gini)
     ) %>% 
@@
 ```{r}
-df_dat |> 
+df_dat <- df_dat |> 
+    # change scale of GNI to log so the magnitud is not so dominated by $
+    mutate(gni_log = log(gni2)) |> 
     mutate(
@@
         d_gini = slide_dbl(gini, diff, .before = 1, .after = 0, .complete = TRUE),
-        d_gni2 = slide_dbl(gni2, diff, .before = 1, .after = 0, .complete = TRUE)
+        d_gni = slide_dbl(gni_log, diff, .before = 1, .after = 0, .complete = TRUE)
     ) |> 
     mutate(
+        magni = sqrt(d_EF^2 + d_gini^2 + d_gni^2), 
         theta = atan(d_gini/d_EF),
-        gamma = atan(d_gini/d_gni2)
+        gamma = atan(d_gini/d_gni)
     ) 
+```
+
+```{r}
+df_dat |> 
+    as_tibble() |> 
+    ungroup() |> 
+    group_by(country) |> 
+    summarize(mgt = sum(magni, na.rm = TRUE), 
+              theta = sum(theta, na.rm = TRUE),
+              gamma = sum(gamma, na.rm = TRUE)) |> 
+    arrange(desc(mgt))
+```
+```{r}
+df_dat |> 
+    as_tibble() |> 
+    select(country, year, d_EF, d_gini, d_gni) |> 
+    pivot_longer(cols = starts_with("d_"), names_to = "vars", values_to = "vals") |> 
+    filter(!is.na(vals)) |> 
+    ggplot(aes(year, vals, group = country)) +
+    geom_path(size = 0.2, alpha = 0.5) +
+    facet_wrap(~vars, scales = "free") +
+    theme_light(base_size = 18)
+```
+```{r}
+df_dat |> 
+    ggplot(aes(EFconsPerCap, gini)) +
+    geom_density2d() +
+    geom_point(aes(color = gni_log)) +
+    geom_path(aes(group = country), size = 0.15) +
+    scico::scale_color_scico() +
+    theme_light(base_size = 18)
+```
+```{r}
+df_dat |> 
+    ggplot(aes(gni_log, gini)) +
+    geom_density2d() +
+    geom_point(aes(color = EFconsPerCap)) +
+    geom_path(aes(group = country), size = 0.15) +
+    scico::scale_color_scico(palette = "berlin") +
+    theme_light(base_size = 18)
+```
+```{r}
+df_dat |> 
+    ggplot(aes(gni_log, EFconsPerCap)) +
+    geom_density2d() +
+    geom_point(aes(color = gini)) +
+    geom_path(aes(group = country), size = 0.15) +
+    scico::scale_color_scico() +
+    theme_light(base_size = 18)
+```
+Who is the weird trajectory out there, world? No it's Luxemburg
+
+```{r}
+df_dat |> 
+    arrange(desc(EFconsPerCap))
 ```
@@
 ```{r}
+df_summary <- df_dat |> 
+    as_tibble() |> 
+    ungroup() |> 
+    group_by(country) |> 
+    summarize(m = sum(magni, na.rm = TRUE), 
+              t = sum(theta, na.rm = TRUE),
+              g = sum(gamma, na.rm = TRUE),
+              sd_mag = sd(magni, na.rm = TRUE),
+              sd_t = sd(theta, na.rm = TRUE),
+              sd_g = sd(gamma, na.rm = TRUE), 
+              ef = mean(EFconsPerCap, na.rm = TRUE),
+              inq = mean(gini, na.rm = TRUE),
+              gni = mean(gni_log, na.rm = TRUE),
+              #sd_ef = sd(ef, na.rm = TRUE),
+              sd_inq = sd(gini, na.rm = TRUE),
+              sd_gni = sd(gni_log, na.rm = TRUE))
+
 m <- "ward.D2" # minimize the total within-cluster variance
@@
 clust_num <- NbClust::NbClust( 
-    data = df_dat %>% as.data.frame() %>% 
-        select(where(is.numeric), -year),
+    data = df_summary %>% as.data.frame() %>% 
+        select(where(is.numeric)),
     #distance = "maximum",
-    min.nc = 2, max.nc = 10, 
+    min.nc = 2, max.nc = 7, 
     method = m, 
@@
 stab <- clValid::clValid(
-    obj = df_dat %>% as.data.frame() %>%
-        select(where(is.numeric), -year) %>%
+    obj = df_summary %>% as.data.frame() %>% 
+        select(where(is.numeric)) %>%
         as.matrix(),
-    nClust = c(2:9),
+    nClust = c(2:7),
     clMethods = c("hierarchical", "kmeans",  #"som", "model", "diana", "fanny",
@@
 The stability analysis also supports the choice of either 2 or 9 clusters. Partition around medioids does perform well with 9 clusters, but others are supporting the 2 clusters.
+
+```{r}
+stab@clusterObjs$pam$`2`$clustering
+```
+
+```{r fig.width=7, fig.height=8}
+df_summary |> 
+    mutate(clus = stab@clusterObjs$pam$`2`$clustering) |> 
+    arrange(desc(m)) |> 
+    mutate(country = as_factor(country)) |> 
+    mutate(country = fct_reorder2(country, m,g)) |> 
+    pivot_longer(cols = m:sd_gni, names_to = "vars", values_to = "vals") |> 
+    ggplot(aes(vals, country)) +
+    geom_col(aes(fill = as.factor(clus))) +
+    facet_grid(clus~vars, scales = "free") +
+    theme_light(base_size = 10)
+```
+
+```{r}
+df_dat |> 
+    as_tibble() |> 
+    group_by(country) |> 
+    summarize(
+        ef = sum(d_EF, na.rm = TRUE),
+        gini = sum(d_gini, na.rm = TRUE),
+        gni = sum(d_gni, na.rm = TRUE)
+    ) |> 
+    mutate(clus = stab@clusterObjs$pam$`2`$clustering) |> 
+    ggplot(aes(ef, gini)) +
+    geom_segment(
+        aes(x = 0, y = 0, xend = ef, yend = gini, color = as.factor(clus)), arrow = arrow(length = unit(0.5, "cm"))) +
+    geom_text(aes(label = country)) +
+    theme_light(base_size = 20)
+```
+
+```{r}
+df_dat |> 
+    as_tibble() |> 
+    group_by(country) |> 
+    summarize(
+        ef = sum(d_EF, na.rm = TRUE),
+        gini = sum(d_gini, na.rm = TRUE),
+        gni = sum(d_gni, na.rm = TRUE)
+    ) |> 
+    mutate(clus = stab@clusterObjs$pam$`2`$clustering) |> 
+    ggplot(aes(gni, gini)) +
+    geom_segment(aes(x = 0, y = 0, xend = gni, yend = gini, color = as.factor(clus)), arrow = arrow(length = unit(0.5, "cm"))) +
+    geom_text(aes(label = country)) +
+    theme_light(base_size = 20)
+```
+
+### Old stuff
 

### 2022-02-10 Trilemma_clustering.Rmd
updating figures and missing values analysis

@@
 
-The bottle neck is certaintly our proxies of inequality, no matter what we use. Below a mock up example with some arbitrary choices.
+The bottle neck is certaintly our proxies of inequality, no matter what we use. Based on the missing data explorations, inequality data improves from 1995 onward.
 
@@
   rename(EFconsPerCap = value, country = countryName) %>% 
-  filter(year >= 1990)
+  filter(year >= 1995)
 
@@
     inq %>% 
-      filter(!is.na(gini_reported), year >= 1990) %>% 
+      filter(!is.na(gini_reported), year >= 1995) %>% 
       select(country, isoa2 = c2, country_code = c3, year, gini_reported) %>% 
@@
       mutate(ismissing = is.na(value)) %>% 
-      filter(year >= 1990) %>% 
+      filter(year >= 1995) %>% 
       rename(country = country_name, gni = value)
@@
 key_countries <- missingness %>% 
-    filter(missingness < 0.3) %>% 
+    filter(missingness <= 0.3) %>% 
+    pull(country)
+
+droped_countries <- missingness %>% 
+    filter(missingness > 0.3) %>% 
     pull(country)
@@
 ```{r message=FALSE, warning=FALSE}
-df_dat %>% 
-    select(year, EFconsPerCap, gni2, gini) %>% 
+p <- df_dat %>% 
+    select(year, `Ecological footprint` = EFconsPerCap, GNI = gni2, Gini = gini) %>% 
     GGally::ggpairs(
-        columns = 3:5 
-    # lower = list(continuous = GGally::wrap("points", alpha = 0.5, size = 0.5)),
-    # diag = list(continuous = GGally::wrap("densityDiag", alpha = 0.2)),
+        columns = 3:5 ,
+    lower = list(continuous = GGally::wrap("points", alpha = 0.5, size = 0.5)),
+    diag = list(continuous = GGally::wrap("densityDiag", alpha = 0.2))
     # upper = list(continuous = GGally::wrap("cor", size = 2))
-    ) + theme_light(base_size = 20)
+    ) + theme_light(base_size = 6)
+
+# ggsave(
+#     plot = p,
+#     file = "correlogram.png", path = "figures/", device = "png",
+#     width = 4, height = 4, dpi = 300, bg = "white"
+# )
+p + theme_light(base_size = 20)
 ```
@@
     geom_path(size = 0.2, aes(color = year)) +
-  theme(legend.position = c(0.8, 0.8))
+    annotate(geom = "rect", xmin = 0, xmax = 1.68, ymin = 20, ymax = 30,
+             fill = "skyblue", alpha = 0.25) +
+    annotate(geom ="rect",xmin = 0, xmax = 1.68, ymin = 30, ymax = 62,
+             fill = "orange", alpha = 0.1) +
+    annotate(geom ="rect", xmin = 1.68, xmax = 18, ymin = 20, ymax = 30,
+             fill = "orange", alpha = 0.1) +
+    annotate(geom="rect", xmin = 1.68, xmax = 18, ymin = 30, ymax = 62,
+             fill = "red", alpha = 0.1) +
+    geom_hline(yintercept = 30, linetype = 2, color = "gray50") +
+    geom_vline(xintercept = 1.68, linetype = 2, color = "gray50") +
+    scico::scale_color_scico("Year", palette = "bilbao") +
+    labs(y = "Gini", x = "Ecological footprint") + theme_light() +
+    theme(legend.position = "top", panel.grid = element_blank(),
+          legend.key.width = unit(10, "mm"), legend.key.height = unit(3,"mm"))
 
@@
   geom_path(size = 0.2, aes(color = year)) +
-  scale_x_log10()
+    annotate(geom = "rect", xmin = 10, xmax = 1.68, ymin = 20, ymax = 30,
+             fill = "skyblue", alpha = 0.25) +
+    annotate(geom ="rect",xmin = 0, xmax = 1.68, ymin = 30, ymax = 62,
+             fill = "orange", alpha = 0.1) +
+    annotate(geom ="rect", xmin = 1.68, xmax = 18, ymin = 20, ymax = 30,
+             fill = "orange", alpha = 0.1) +
+    annotate(geom="rect", xmin = 1.68, xmax = 18, ymin = 30, ymax = 62,
+             fill = "red", alpha = 0.1) +
+    geom_hline(yintercept = 30, linetype = 2, color = "gray50") +
+    geom_vline(xintercept = 12746, linetype = 2, color = "gray50") +
+    scico::scale_color_scico("Year", palette = "bilbao") +
+      scale_x_log10() +
+    labs(y = "Gini", x = "Ecological footprint") + theme_light() +
+    theme(legend.position = "top", panel.grid = element_blank(),
+          legend.key.width = unit(10, "mm"), legend.key.height = unit(3,"mm"))
 
@@
 ```
-```{r}
-df_dat |> 
+```{r fig.width=6, fig.height=3}
+p <- df_dat |> 
     as_tibble() |> 
@@
     filter(!is.na(vals)) |> 
+    mutate(vars = case_when(
+        vars == "d_EF" ~ "Ecological Footprint",
+        vars == "d_gni" ~ "GNI [log]",
+        vars == "d_gini" ~ "Gini"
+    )) |> 
     ggplot(aes(year, vals, group = country)) +
@@
     facet_wrap(~vars, scales = "free") +
-    theme_light(base_size = 18)
+    labs(x = "Year", y = "First difference") +
+    theme_light(base_size = 7)
+
+# ggsave(
+#     plot = p, path = "figures/", file = "first_diff.png", device = "png",
+#     width = 6, height = 3, dpi = 300, bg = "white")
+p + theme_light(base_size = 20)
 ```
-```{r}
+```{r fig.width=3, fig.height=3}
 df_dat |> 
     ggplot(aes(EFconsPerCap, gini)) +
+    geom_hline(yintercept = c(30,40), linetype = 2, color = "gray50") +
+    geom_vline(xintercept = 1.68, linetype = 2, color = "gray50") +
     geom_density2d() +
@@
     geom_path(aes(group = country), size = 0.15) +
-    scico::scale_color_scico() +
-    theme_light(base_size = 18)
+    scico::scale_color_scico("GNI [log]") +
+    labs(y = "Gini", x = "Ecological footprint") +
+    theme_light(base_size = 18) +
+    theme(legend.position = "top", panel.grid = element_blank())
 ```

### 2022-02-10 inequality-SDGs.Rmd
updating figures and missing values analysis

---
title: "SDGs"
author: "Juan Rocha"
date: "Updated `r Sys.Date()`"
output:
  html_notebook:
    toc: yes
    toc_float: yes
    highlight: tango
    code_folding: hide
    df_print: paged
    theme: 
      bootswatch: cosmo
      code_font:
        google: Fira Code
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  cache = TRUE, warning = FALSE, message = FALSE, echo = TRUE,
  fig.width = 8, fig.height = 5
)

library(tidyverse)
library(naniar)
library(plotly)
library(patchwork)

# time series
library(tsibble)
library(fable)
library(imputeTS)
library(slider)

# # clustering
# library(vegan)
# library(NbClust)
# library(clValid)
# library(mclust)

#theme_set(theme_light(base_size = 12))
```

This notebooks explores synergies and trade-offs between inequality and other SDGs with the World Bank dataset on sustainable development goals at the national scale.

```{r}
dat <- read_csv(file = "~/Documents/Projects/DATA/WorldBank/SDG_csv/SDGData.csv") |> 
    janitor::clean_names()

dat
```
```{r}
key <- read_csv(
    file = "~/Documents/Projects/DATA/WorldBank/SDG_csv/SDGSeries.csv"
) |> janitor::clean_names()

key |> pull(topic) |> unique()
```

### 2022-02-10 missing_values.Rmd
updating figures and missing values analysis

@@
 ```{r}
-ecofp %>%
+g0 <- ecofp %>%
     filter(record == "EFConsPerCap") %>%
-    mutate(missing = is.na(value)) %>% 
+    select(countryName, year, value) %>% 
+    pivot_wider(id_col = countryName, names_from = year, values_from = value,
+                values_fill = NA) %>% 
+    pivot_longer(cols = 2:last_col(), names_to = "year", values_to = "value") %>% 
+    mutate(missing = is.na(value), year = as.numeric(year)) %>% 
     ggplot(aes(year, countryName)) +
     geom_tile(aes(fill = value)) +
-    scale_fill_viridis_c(option = "C") +
-    theme(axis.text.y = element_text(size = 4), legend.position = "top")
+    scale_fill_viridis_c(
+        "Ecological footprint",option = "D", na.value = "grey75")+
+    labs(x = "Year", y = "Countries (N=211)", tag = "A") +
+    scale_x_continuous(breaks = c( 1970, 1980, 1990, 2000, 2010)) +
+    theme_light(base_size = 7)  +
+    theme(axis.text.y = element_blank(), legend.position = "top")
+
+g0
 ```
 
-```{r fig.height=3, fig.width=8}
+```{r fig.height=5, fig.width=7}
 g1 <- ecofp %>%
@@
     ggplot(aes(year, n_nas)) + 
-    geom_col(aes(fill = (year>1995)), alpha = 0.5) + 
+    geom_col( alpha = 0.5) + 
+    geom_vline(xintercept = 1995, color = "red", linetype = 2) +
     coord_flip() +
     theme_light(base_size = 7) +
-    theme(legend.position = c(0.8,0.8)) + labs(title = "Missing values per year")
-
+    theme(legend.position = c(0.8,0.8)) + 
+    labs(y = "Number of missing values", x="Year", tag = "B")
+#g1
 g2 <- ecofp %>%
@@
     ggplot(aes(countryName, n_nas)) + 
-    geom_col(aes(fill = (n_nas < 11)), alpha = 0.5) + coord_flip() +
+    geom_col(alpha = 0.5) + 
+    geom_hline(yintercept = 10, color = "red", linetype = 2) +
+    coord_flip() +
     theme_light(base_size = 7) + 
-    labs(title = "Missing values per country after 1995") +
+    labs(y = "Number of missing values", x = "Country", tag = "C") +
     theme(legend.position = c(0.8, 0.2))
 
-g1 + g2
+g0 / (g1 + g2)
+
+# ggsave(
+#     filename = "ecofp_missing_values.png",
+#     plot = (g0 / (g1 + g2)), 
+#     device = "png", path = "figures/",
+#     width = 7, height = 5, bg = "white", dpi = 400
+# )
 ```
@@
 ```{r}
-inq %>% 
+i2 <- inq %>% 
     filter(year > 1950, resource_detailed == "Income, gross") %>% 
@@
     geom_miss_point(size = 1, alpha = 0.5) +
-    theme(legend.position = c(0.1,0.8))
+    labs(y = "Palma index", x = "Gini index", tag = "B") + 
+    scale_color_manual("Missing values", values = c("purple", "orange")) +
+    theme_light(base_size = 6) +
+    theme(legend.position = c(0.2,0.8))
+
+i2
 ```
@@
 ```{r}
-inq %>% 
+i1 <- inq %>% 
     filter(year > 1950, resource_detailed == "Income, gross") %>% 
@@
     geom_tile(aes(fill = gini)) +
-    scale_fill_viridis_c(option = "C") +
-    theme(axis.text.y = element_text(size = 4), legend.position = "top")
+    scale_fill_viridis_c(
+        "Mean Gini",option = "C", na.value = "grey75")+
+    labs(x = "Year", y = "Countries (N=200)", tag = "A") +
+    theme_light(base_size = 6) +
+    theme(axis.text.y = element_blank(), legend.position = "top")
+
+i1
 ```
@@
 ```{r}
-inq %>% 
-    filter(!is.na(gini_reported), year > 1990) %>% select(country, year, gini_reported) %>% 
+i1 <- inq %>% 
+    filter(!is.na(gini_reported), year > 1990) %>% 
+    select(country, year, gini_reported) %>% 
     group_by(country, year) %>% 
@@
     geom_tile(aes(fill = gini_mean)) +
-    scale_fill_viridis_c(option = "C") +
-    theme(axis.text.y = element_text(size = 4), legend.position = "top")
+    scale_fill_viridis_c(
+        "Mean Gini",option = "C", na.value = "grey75")+
+    labs(x = "Year", y = "Countries (N=196)", tag = "A") +
+    theme_light(base_size = 6) +
+    theme(axis.text.y = element_blank(), legend.position = "top", 
+          legend.key.height = unit(3, "mm"))
+i1
 ```
@@
 ```{r}
-inq %>% 
+i3 <- inq %>% 
     filter(!is.na(gini_reported), year > 1990) %>% select(country, year, gini_reported) %>% 
@@
     group_by(year) %>% 
-    summarise(n_nas = sum(missing), missingness = n_nas / 200)
+    summarise(n_nas = sum(missing), missingness = n_nas / 200) |> 
+    ggplot(aes(year, missingness)) +
+    geom_col(alpha = 0.5) +
+    theme_light(base_size = 6) +
+    theme(legend.position = c(0.8,0.8)) +
+    labs(y = "Proportion of countries with\n missing values", x="Year", tag = "C")
+
+i3
 ```
@@
 ```{r}
-inq %>% 
+i4 <- inq %>% 
     filter(!is.na(gini_reported), year > 1990) %>% select(country, year, gini_reported) %>% 
@@
     summarise(n_nas = sum(missing), missingness = n_nas / (2018-1990)) %>% 
-    arrange(missingness)
+    arrange(missingness) |> 
+    filter(missingness <=0.5) |> 
+    mutate(country = as_factor(country)) |> 
+    ggplot(aes(missingness, country)) +
+    geom_col(alpha = 0.5) +
+    geom_vline(xintercept = 0.3, color = "red", linetype = 2) +
+    theme_light(base_size = 6) + 
+    labs(x = "Proportion of missing values", y = "Country", tag = "D") +
+    theme(axis.text.y = element_text(size = 4))
+i4
 ```
@@
 
+```{r}
+(i1 / (i2+i3+i4))
+
+# ggsave(
+#     filename = "inq_missing_values.png",
+#     plot = (i1 / (i2+i3+i4)),
+#     device = "png", path = "figures/",
+#     width = 7, height = 6, bg = "white", dpi = 400
+# )
+```
+
 ## GNI: Prosperity
@@
 dat1 %>% 
-    select(-x33) %>% 
+    select(-x35) %>% 
     filter(str_detect(indicator_name, "GNI")) %>% 
-    pivot_longer(cols = x1990:x2017, names_to = "year", values_to = "value") %>% 
+    pivot_longer(cols = x1990:x2019, names_to = "year", values_to = "value") %>% 
     mutate(year = str_remove(year, "x")) %>% 
@@
 missingness <- dat1 %>% 
-    select(-x33) %>% 
+    select(-x35) %>% 
     filter(str_detect(indicator_name, "GNI")) %>% 
-    pivot_longer(cols = x1990:x2017, names_to = "year", values_to = "value") %>% 
+    pivot_longer(cols = x1990:x2019, names_to = "year", values_to = "value") %>% 
     mutate(year = str_remove(year, "x")) %>% 
@@
 
-missingness %>% 
+p1 <- missingness %>% 
     ggplot(aes(country_code, indicator_name)) +
     geom_tile(aes(fill = missingness)) + 
-    scale_fill_viridis_c(option = "C") +
-    theme(axis.text.x = element_text(angle =90), legend.position = "top")
+    scale_fill_viridis_c(
+        "Proportion of missing values", option = "D", na.value = "grey75")+
+    labs(y = "Prosperity metrics", x = "Countries (N=239)", tag = "A") +
+    theme_light(base_size = 6) +
+    theme(axis.text.x = element_blank(), legend.position = "top", 
+          legend.key.height = unit(3, "mm"))
+p1
+
 ```
@@
 ```{r}
-missingness %>% 
+p2 <- missingness %>% 
     ungroup() %>% 
@@
     summarize(x = mean(missingness)) %>% 
-    arrange(x)
+    arrange(desc(x)) |> 
+    mutate(indicator_name = as_factor(indicator_name)) |> 
+    ggplot(aes(x, indicator_name)) +
+    geom_col(alpha = 0.5) +
+    #geom_vline(xintercept = 0.3, color = "red", linetype = 2) +
+    theme_light(base_size = 6) + 
+    labs(x = "Proportion of missing values", y = "Prosperity metrics", tag = "B") +
+    theme(axis.text.y = element_text(size = 6))
+p2
 ```
@@
 ```{r}
-missingness %>% 
+p3 <- missingness %>% 
     ungroup() %>% 
@@
     summarize(x = mean(missingness)) %>% 
-    arrange(desc(x)) 
+    arrange(x) |> 
+    rownames_to_column("rank") |> 
+    mutate(rank = as.numeric(rank)) |> 
+    ggplot(aes(x, rank)) +
+    geom_point(alpha = 0.5) +
+    geom_vline(xintercept = 0.3, color = "red", linetype = 2) +
+    theme_light(base_size = 6) + 
+    labs(x = "Proportion of missing values", y = "Country rank", tag = "C") +
+    theme(axis.text.y = element_text(size = 4))
+p3
 ```
@@
 Perhaps 60% is tolerable?
+
+```{r}
+p1 / (p2+p3)
+
+ggsave(
+    filename = "prosperity_missing_values.png",
+    plot = p1 / (p2+p3),
+    device = "png", path = "figures/",
+    width = 7, height = 5, bg = "white", dpi = 400
+)
+```
 

### 2022-02-10 figures.R

@@ -1,0 +1,125 @@
+library(ggplot2)
+library(patchwork)
+
+#### Conceptual figure ####
+old_par <- par()
+
+quartz(width = 2.5, height = 2.5, pointsize = 7)
+par(mar = c(0,0,0,0))
+box3D(
+    x0 = -0.8, y0 = -0.8, z0 = -0.8, 
+    x1 = 0.8, y1 = 0.8, z1 = 0.8, 
+    border = "black", lwd = 1, 
+    col = gg.col(1, alpha = 0.25), 
+    xlab = "Inequality", ylab = "Environment",
+    zlab = "Prosperity", cex.lab = 1.2, mar = c(0,0,0,0))
+
+quartz.save(
+    file="figures/cube_concept.png", type = "png", dpi = 300,
+    width = 2.5, height = 2.5, pointsize = 7, bg = "white"
+)
+dev.off()
+par(old_par)
+
+
+p1 <- ggplot() +
+    geom_blank() + 
+    geom_rect(aes(xmin = 0, xmax = 0.5, ymin = 0, ymax = 0.5), 
+              fill = "skyblue", alpha = 0.5) +
+    geom_rect(aes(xmin = 0, xmax = 0.5, ymin = 0.5, ymax = 1), 
+              fill = "orange", alpha = 0.5) +
+    geom_rect(aes(xmin = 0.5, xmax = 1, ymin = 0, ymax = 0.5), 
+              fill = "orange", alpha = 0.5) +
+    geom_rect(aes(xmin = 0.5, xmax = 1, ymin = 0.5, ymax = 1), 
+              fill = "red", alpha = 0.5) +
+    geom_abline(intercept = 0, slope = 1, color = "blue") +
+    geom_abline(intercept = 1, slope = -1, color = "red") +
+    geom_hline(yintercept = 0.5, linetype = 2, color = "grey50", size = 0.5) +
+    geom_vline(xintercept = 0.5, linetype = 2, color = "grey50", size = 0.5) +
+    annotate("text", x = 0.15, y = 0.2, label = "Synergies", size = 2,
+             color = "blue", angle = 45) +
+    annotate("text", x = 0.15, y = 0.9, label = "Trade-off", size = 2,
+             color = "red", angle = 360-45) +
+    lims(x = c(0,1), y = c(0,1)) +
+    labs(x = "Inequality", y = "Environmental footprint", tag = "D") +
+    coord_equal() +
+    theme_light(base_size = 7) +
+    theme(axis.text = element_blank(), 
[…]

### 2022-02-10 lit_review.R

@@ -102,7 +102,7 @@ toc() #43955.608 or 12.2hrs
 save(tset.TM, topicNumber.TM, dtm, file = "data/211020_topic_models_gibbs.RData")
 
 
-
+load("data/211020_topic_models_gibbs.RData")
 
 
 #### visualizations ####
@@ -247,9 +247,10 @@ ggsave(
 df_documents |> 
     group_by(topic) |> 
     arrange(desc(gamma)) |> 
-    top_n(10) |> 
-    left_join(select(dat, document = title, year, doi)) |> 
-    write_csv("data/top_papers.csv")
+    #top_n(10) |> 
+    left_join(select(dat, document = title, year)) #|> 
+    #write_csv("data/topic_papers.csv")
+    #save(x, file = "data/topic_papers.RData")
 
 
 ### networks

### 2022-02-11 Trilemma_clustering.Rmd
improved figures for paper

@@
     df_print: paged
-    theme: 
-      bootswatch: cosmo
-      code_font:
-        google: Fira Code
 editor_options: 
@@
 # library(mclust)
+# library(ggbiplot)
 
@@
     annotate(geom = "rect", xmin = 0, xmax = 1.68, ymin = 20, ymax = 30,
-             fill = "skyblue", alpha = 0.25) +
+             fill = "skyblue", alpha = 0.2) +
     annotate(geom ="rect",xmin = 0, xmax = 1.68, ymin = 30, ymax = 62,
-             fill = "orange", alpha = 0.1) +
+             fill = "orange", alpha = 0.2) +
     annotate(geom ="rect", xmin = 1.68, xmax = 18, ymin = 20, ymax = 30,
-             fill = "orange", alpha = 0.1) +
+             fill = "orange", alpha = 0.2) +
     annotate(geom="rect", xmin = 1.68, xmax = 18, ymin = 30, ymax = 62,
-             fill = "red", alpha = 0.1) +
+             fill = "red", alpha = 0.2) +
     geom_hline(yintercept = 30, linetype = 2, color = "gray50") +
@@
     scico::scale_color_scico("Year", palette = "bilbao") +
-    labs(y = "Gini", x = "Ecological footprint") + theme_light() +
+    labs(y = "Gini", x = "Ecological footprint", tag = "A") + theme_light() +
     theme(legend.position = "top", panel.grid = element_blank(),
@@
   geom_path(size = 0.2, aes(color = year)) +
-    annotate(geom = "rect", xmin = 10, xmax = 1.68, ymin = 20, ymax = 30,
-             fill = "skyblue", alpha = 0.25) +
-    annotate(geom ="rect",xmin = 0, xmax = 1.68, ymin = 30, ymax = 62,
-             fill = "orange", alpha = 0.1) +
-    annotate(geom ="rect", xmin = 1.68, xmax = 18, ymin = 20, ymax = 30,
-             fill = "orange", alpha = 0.1) +
-    annotate(geom="rect", xmin = 1.68, xmax = 18, ymin = 30, ymax = 62,
-             fill = "red", alpha = 0.1) +
+    annotate(geom = "rect", xmin = 10, xmax = 12746, ymin = 20, ymax = 30,
+             fill = "orange", alpha = 0.2) +
+    annotate(geom ="rect",xmin = 10, xmax = 12746, ymin = 30, ymax = 62,
+             fill = "red", alpha = 0.2) +
+    annotate(geom ="rect", xmin = 12746, xmax = 105000, ymin = 20, ymax = 30,
+             fill = "skyblue", alpha = 0.2) +
+    annotate(geom="rect", xmin = 12746, xmax = 105000, ymin = 30, ymax = 62,
+             fill = "orange", alpha = 0.2) +
     geom_hline(yintercept = 30, linetype = 2, color = "gray50") +
@@
       scale_x_log10() +
-    labs(y = "Gini", x = "Ecological footprint") + theme_light() +
+    labs(y = "Gini", x = "GNI [log]", tag = "B") + theme_light() +
     theme(legend.position = "top", panel.grid = element_blank(),
@@
   geom_point(size = 0.5, show.legend = FALSE) +
-  geom_path(size = 0.2, aes(color = year)) +
-  scale_x_log10()
+  geom_path(size = 0.2, aes(color = year))+
+    annotate(geom = "rect", xmin = 10, xmax = 12746, ymin = 0, ymax = 1.68,
+             fill = "orange", alpha = 0.2) +
+    annotate(geom ="rect",xmin = 10, xmax = 12746, ymin = 1.68, ymax = 18,
+             fill = "red", alpha = 0.2) +
+    annotate(geom ="rect", xmin = 12746, xmax = 105000, ymin = 0, ymax = 1.68,
+             fill = "skyblue", alpha = 0.2) +
+    annotate(geom="rect", xmin = 12746, xmax = 105000, ymin = 1.68, ymax = 18,
+             fill = "orange", alpha = 0.2) +
+    geom_hline(yintercept = 1.68, linetype = 2, color = "gray50") +
+    geom_vline(xintercept = 12746, linetype = 2, color = "gray50") +
+    scico::scale_color_scico("Year", palette = "bilbao") +
+      scale_x_log10() +
+    labs(y = "Ecological footprint", x = "GNI [log]", tag = "C") + theme_light() +
+    theme(legend.position = "top", panel.grid = element_blank(),
+          legend.key.width = unit(10, "mm"), legend.key.height = unit(3,"mm"))
 
-p1 + p2 + p3 + plot_layout(guides = "collect") & theme_light(base_size = 18)
+# ggsave(
+#     plot = p1 + p2 + p3 + plot_layout(guides = "collect") & 
+#         theme_light(base_size = 7) & 
+#         theme(legend.position = "bottom", panel.grid = element_blank(),
+#           legend.key.width = unit(10, "mm"), legend.key.height = unit(3,"mm")),
+#     path = "figures/", file = "cube_time.png", device = "png", dpi = 300,
+#     width = 7, height = 3, bg = "white"
+# )
+
+p1 + p2 + p3 + 
+    plot_layout(guides = "collect") & theme_light(base_size = 18) & 
+    theme(legend.position = "bottom", panel.grid = element_blank(),
+          legend.key.width = unit(20, "mm"), legend.key.height = unit(3,"mm"))
 ```
@@
 ```{r fig.width=3, fig.height=3}
-df_dat |> 
+c1 <- df_dat |> 
     ggplot(aes(EFconsPerCap, gini)) +
-    geom_hline(yintercept = c(30,40), linetype = 2, color = "gray50") +
+    geom_hline(yintercept = 30, linetype = 2, color = "gray50") +
     geom_vline(xintercept = 1.68, linetype = 2, color = "gray50") +
@@
     labs(y = "Gini", x = "Ecological footprint") +
-    theme_light(base_size = 18) +
+    theme_light(base_size = 7) +
     theme(legend.position = "top", panel.grid = element_blank())
+
+c1 + theme_light(base_size = 18)
 ```
-```{r}
-df_dat |> 
-    ggplot(aes(gni_log, gini)) +
+```{r fig.width=3, fig.height=3}
+c2 <- df_dat |> 
+    ggplot(aes(gni2, gini)) +
     geom_density2d() +
@@
     geom_path(aes(group = country), size = 0.15) +
-    scico::scale_color_scico(palette = "berlin") +
-    theme_light(base_size = 18)
+    geom_hline(yintercept = 30, linetype = 2, color = "gray50") +
+    geom_vline(xintercept = (12546), linetype = 2, color = "gray50") +
+    scico::scale_color_scico("Ecological footprint",palette = "berlin") +
+    labs(x= "GNI [log]", y = "Gini") +
+    scale_x_log10() +
+    theme_light(base_size = 7) +
+    theme(legend.position = "top", panel.grid = element_blank())
+
+c2 + theme_light(base_size = 7)
 ```
-```{r}
-df_dat |> 
-    ggplot(aes(gni_log, EFconsPerCap)) +
+```{r fig.width=3, fig.height=3}
+c3 <- df_dat |> 
+    ggplot(aes(gni2, EFconsPerCap)) +
     geom_density2d() +
@@
     geom_path(aes(group = country), size = 0.15) +
-    scico::scale_color_scico() +
-    theme_light(base_size = 18)
+    scico::scale_color_scico("Gini") +
+    labs(x = "GNI", y = "Ecological footprint") +
+    scale_x_log10()+
+    geom_hline(yintercept = 1.68, linetype = 2, color = "gray50") +
+    geom_vline(xintercept = (12546), linetype = 2, color = "gray50") +
+    theme_light(base_size = 7) +
+    theme(legend.position = "top", panel.grid = element_blank())
+
+c3 + theme_light(base_size = 18)
 ```
@@
 
-```{r}
+```{r fig.width=6, fig.height=2}
 df_summary <- df_dat |> 
@@
     group_by(country) |> 
-    summarize(m = sum(magni, na.rm = TRUE), 
-              t = sum(theta, na.rm = TRUE),
-              g = sum(gamma, na.rm = TRUE),
-              sd_mag = sd(magni, na.rm = TRUE),
-              sd_t = sd(theta, na.rm = TRUE),
-              sd_g = sd(gamma, na.rm = TRUE), 
-              ef = mean(EFconsPerCap, na.rm = TRUE),
-              inq = mean(gini, na.rm = TRUE),
-              gni = mean(gni_log, na.rm = TRUE),
-              #sd_ef = sd(ef, na.rm = TRUE),
-              sd_inq = sd(gini, na.rm = TRUE),
-              sd_gni = sd(gni_log, na.rm = TRUE))
+    dplyr::summarize(sd_magnitude = sd(magni, na.rm = TRUE),
+              sd_theta = sd(theta, na.rm = TRUE),
+              sd_gamma = sd(gamma, na.rm = TRUE), 
+              EF = mean(EFconsPerCap, na.rm = TRUE),
+              Gini = mean(gini, na.rm = TRUE),
+              GNI = mean(gni_log, na.rm = TRUE),
+              sd_EF = sd(EFconsPerCap, na.rm = TRUE),
+              sd_Gini = sd(gini, na.rm = TRUE),
+              sd_GNI = sd(gni_log, na.rm = TRUE),
+              magnitude_m = mean(magni, na.rm = TRUE), 
+              theta_m = mean(theta, na.rm = TRUE),
+              gamma_m = mean(gamma, na.rm = TRUE),
+              magnitude_a = sum(magni, na.rm = TRUE), 
+              theta_a = sum(theta, na.rm = TRUE),
+              gamma_a = sum(gamma, na.rm = TRUE))
 
@@
 
-It is a bit contested because the numeric indices favours 2 clusters with 6 of them, but 3 and 4 clusters are favoured by 5 metrics each.
-The Hubert index (which is graphical) favours 4 indexes as well, tieing up with the 2 cluster option.
+The Hubert index (which is graphical) favours 4 groups while D index favours 3 groups.
+Thus in total 9 metrics proposed 2 groups, 9 metrics proposed 3 groups, and all other options are minorities.
+Let's continue the analysis with 3 to favour diversity.
 
@@
     #metric = "manhattan",
-    method = "ward",
+    method = "ward.D2",
     verbose = FALSE
@@
 ```
-The stability analysis also supports the choice of either 2 or 9 clusters. Partition around medioids does perform well with 9 clusters, but others are supporting the 2 clusters.
+The stability analysis also supports the choice of either 2 or 7 clusters. Partition around medioids does perform well with 9 clusters, but others are supporting the 2 clusters.
 
@@
     mutate(clus = stab@clusterObjs$pam$`2`$clustering) |> 
-    arrange(desc(m)) |> 
     mutate(country = as_factor(country)) |> 
-    mutate(country = fct_reorder2(country, m,g)) |> 
-    pivot_longer(cols = m:sd_gni, names_to = "vars", values_to = "vals") |> 
+    mutate(country = fct_reorder2(country, magnitude, Gini)) |> 
+    pivot_longer(cols = sd_magnitude:gamma, names_to = "vars", values_to = "vals") |> 
     ggplot(aes(vals, country)) +
@@
 ```{r}
-df_dat |> 
+df_summary |>
+    select(-starts_with("country")) |> 
+    GGally::ggpairs()
+```
+
+```{r fig.width=3, fig.height=3}
+pc <- prcomp(df_summary |> 
+                 select(-starts_with("country")),
+             center = TRUE, scale. = TRUE)
+## Fix short names for plotting
+df_summary <- df_summary |> 
+    left_join(
+        dat |> select(country, country_code) |> 
+            filter(!is.na(country_code)) |> unique())
+
+pc1 <- ggbiplot::ggbiplot(pc, groups = as.character(stab@clusterObjs$pam$`2`$clustering),
+         ellipse = TRUE, labels = df_summary$country_code, labels.size = 2.5,
+         varname.size = 4) +
+    scale_color_manual("Groups", values = c("orange", "purple")) +
+    theme_light(base_size = 18) +
+    theme(legend.position = c(0.2, 0.2), panel.grid = element_blank())
+
+pc1
+```
+
+Note to self: It does not make sense to plot the average distance traversed in the space per country.
+Take Luxemburg as example, it started and ended verly close by.
+The absolute distance is what I'm calculating with the euclidian distance on that plane, then the first-diff values are just increases or decreases with respect to initial conditions.
+Averages are then misleading, because on average Luxemburg moved a lot (in circles).
+
+Averages are worth calculating for the parameter space, the vector field of the cube.
+But not for individual countries, or particules moving within the cube.
+Be aware of that when presenting and interpreting the results.
+```{r fig.height=3,fig.width=3}
+a1 <- df_dat |> 
     as_tibble() |> 
     group_by(country) |> 
-    summarize(
-        ef = sum(d_EF, na.rm = TRUE),
-        gini = sum(d_gini, na.rm = TRUE),
-        gni = sum(d_gni, na.rm = TRUE)
+    dplyr::summarize(
+        EF = sum(d_EF, na.rm = TRUE),
+        Gini = sum(d_gini, na.rm = TRUE),
+        GNI = sum(d_gni, na.rm = TRUE)
     ) |> 
-    mutate(clus = stab@clusterObjs$pam$`2`$clustering) |> 
-    ggplot(aes(ef, gini)) +
+    mutate(Group = as.character(stab@clusterObjs$pam$`2`$clustering)) |> 
+    left_join(
+        dat |> select(country, country_code) |> 
+            filter(!is.na(country_code)) |> unique()) |> 
+    ggplot(aes(EF, Gini)) +
     geom_segment(
-        aes(x = 0, y = 0, xend = ef, yend = gini, color = as.factor(clus)), arrow = arrow(length = unit(0.5, "cm"))) +
-    geom_text(aes(label = country)) +
-    theme_light(base_size = 20)
+        aes(x = 0, y = 0, xend = EF, yend = Gini, color = as.factor(Group)),
+        arrow = arrow(length = unit(0.15, "cm")), size = 0.15) +
+    geom_text(
+        aes(x = EF *1.1, y = Gini * 1.1, label = country_code, colour = Group), size = 2) +
+    scale_color_manual("Groups", values = c("orange", "purple")) +
+    labs(x = "Ecological Footprint", y = "Gini") +
+    theme_light(base_size = 6) +
+    theme(panel.grid = element_blank(), legend.position = c(0.8,0.2))
+
+a1
+
+c1 + annotation_custom(
+        grob = ggplotGrob(a1),
+        xmin = 10, xmax = 18, ymin = 40, ymax = 60
+    ) 
 ```
 
-```{r}
-df_dat |> 
+```{r fig.width=2, fig.height=2}
+a2 <- df_dat |> 
     as_tibble() |> 
     group_by(country) |> 
-    summarize(
-        ef = sum(d_EF, na.rm = TRUE),
-        gini = sum(d_gini, na.rm = TRUE),
-        gni = sum(d_gni, na.rm = TRUE)
+    dplyr::summarize(
+        EF = sum(d_EF, na.rm = TRUE),
+        Gini = sum(d_gini, na.rm = TRUE),
+        GNI = sum(d_gni, na.rm = TRUE)
     ) |> 
-    mutate(clus = stab@clusterObjs$pam$`2`$clustering) |> 
-    ggplot(aes(gni, gini)) +
-    geom_segment(aes(x = 0, y = 0, xend = gni, yend = gini, color = as.factor(clus)), arrow = arrow(length = unit(0.5, "cm"))) +
-    geom_text(aes(label = country)) +
-    theme_light(base_size = 20)
+    mutate(Group = as.character(stab@clusterObjs$pam$`2`$clustering)) |> 
+    left_join(
+        dat |> select(country, country_code) |> 
+            filter(!is.na(country_code)) |> unique()) |> 
+    ggplot(aes(GNI, Gini)) +
+    geom_segment(
+        aes(x = 0, y = 0, xend = GNI, yend = Gini, 
+            color = as.factor(Group)), 
+        arrow = arrow(length = unit(0.15, "cm")), size = 0.15) +
+    geom_text(aes(x = GNI * 1.1, y = Gini *1.1, 
+                  label = country_code, colour = Group), size = 2) +
+    scale_color_manual("Groups", values = c("orange", "purple")) +
+    theme_light(base_size = 6) +
+    theme(panel.grid = element_blank(), legend.position = c(0.8,0.2))
+
+c2 + annotation_custom(
+        grob = ggplotGrob(a2),
+        xmin = 1e+01, xmax = 1e+03, ymin = 40, ymax = 60
+    )
+
+a2
 ```

### 2022-03-02 SDGs_UN.Rmd
data explorations for SDGs datasets

---
title: "SDGs UN data"
author: "Juan Rocha"
date: "Updated `r Sys.Date()`"
output:
  html_notebook:
    toc: yes
    toc_float: yes
    highlight: tango
    code_folding: hide
    df_print: paged
    theme: 
      bootswatch: cosmo
      code_font:
        google: Fira Code
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  cache = TRUE, warning = FALSE, message = FALSE, echo = TRUE,
  fig.width = 8, fig.height = 5
)

library(tidyverse)
library(naniar)
library(plotly)
library(patchwork)

# time series
library(tsibble)
library(fable)
library(imputeTS)
library(slider)

# # clustering
# library(vegan)
# library(NbClust)
# library(clValid)
# library(mclust)

#theme_set(theme_light(base_size = 12))
```

This notebooks explores synergies and trade-offs between inequality and other SDGs with the UN dataset on sustainable development goals at the national scale.

```{r}
dat <- read_csv(
    file = "~/Documents/Projects/DATA/SDGs_UNStats/20180817102044132_juan.rocha@su.se_data.csv") |> 
    janitor::clean_names()

```

Filter out non-countries:

```{r}
# dat |> pull(geo_area_name) |> unique()

no_country <- c(
    "World", "South America" ,"Oceania","Western Africa","Central America", "Eastern Africa", 
    "Northern Africa","Middle Africa", "Southern Africa","Northern America", "Caribbean" ,"Eastern Asia", 
    "Southern Asia" ,"South-Eastern Asia", "Southern Europe", "Australia and New Zealand",  "Melanesia", 
    "Polynesia" , "Central and Southern Asia" ,"Caucasus and Central Asia" , "Asia" ,"Central Asia" , 
    "Western Asia" , "Europe", "Eastern Europe", "Northern Europe" , "Western Europe",
    "Least Developed Countries (LDCs)","Sub-Saharan Africa" , "Latin America and the Caribbean" ,
    "Landlocked developing countries (LLDCs)", 
    "Western Asia (exc. Armenia, Azerbaijan, Cyprus, Israel and Georgia)", "Europe and Northern America",
    "Developed regions (Europe, Cyprus, Israel, Northern America, Japan, Australia & New Zealand)", 
    "Eastern Asia (excluding Japan)","Oceania (exc. Australia and New Zealand)", 
    "Micronesia (Federated States of)" , "Small island developing States (SIDS)" ,
    "Sub-Saharan Africa (inc. Sudan)", "Northern Africa (exc. Sudan)","Northern Africa and Western Asia", 
    "Eastern and South-Eastern Asia" , "French Polynesia" , "China, Macao Special Administrative Region"  , 
    "New Caledonia" ,"British Virgin Islands" ,"American Samoa", "Guam" ,"Northern Mariana Islands",
    "Tokelau", "Wallis and Futuna Islands", "Holy See",  "China, Hong Kong Special Administrative Region",
    "Africa", "Americas", "Micronesia", "Southern Asia (excluding India)" ,
    "Eastern Asia (excluding Japan and China)" , "Developing regions" , "Other non-specified areas",
    "Faroe Islands",  "Falkland Islands (Malvinas)","Greenland" , "Guadeloupe", "Martinique", "Montserrat",
    "Réunion" ,"United States Virgin Islands" , "International Centers (FAO)" , "Regional Centres (FAO)",
    "Mayotte" , "Cayman Islands" , "Netherlands Antilles","Curaçao" , "Sint Maarten (Dutch part)",
    "Bonaire, Sint Eustatius and Saba" , "Gibraltar" , "Saint Pierre and Miquelon" , "Channel Islands",
    "British Indian Ocean Territory" , "Christmas Island" , "Cocos (Keeling) Islands",
    "South Georgia and the South Sandwich Islands" , "Åland Islands" , "French Southern Territories",
    "Heard Island and McDonald Islands" , "Norfolk Island", "United States Minor Outlying Islands" ,
    "Pitcairn",  "Saint Barthélemy" , "Saint Helena", "Saint Martin (French Part)" , "Western Sahara",
    "Svalbard and Jan Mayen Islands","Guernsey", "Jersey" ,"European Union", "Bouvet Island", "Antarctica" ,
    "Iraq (Central Iraq)" , "Iraq (Kurdistan Region)" , "United Kingdom (England and Wales)" ,
    "European Union (EU) Institutions" 
)

# Included Northern Ireland and Scotland as countries although they are listed as part of the UK
## Dataset reduced to countries and SDGs 10, 14, 15
dat <- dat |> 
    filter(!geo_area_name %in% no_country) |> 
    filter(goal %in% c(10,14,15))
```

```{r}
n_ctrs <- dat$geo_area_name |> unique() |> length() # number of countries

p <- dat |> 
    mutate(value = as.numeric(value), is_na = is.na(value)) |> 
    group_by(series_description, time_period, goal) |> 
    summarize(nas = sum(is_na) / n_ctrs) |> 
    mutate(series_description = str_trunc(series_description, width = 100, side = "right")) |> 
    ggplot(aes(time_period, series_description)) +
    geom_tile(aes(fill = nas)) +
    scale_fill_viridis_c() +
    facet_wrap(~goal, scales = "free", nrow = 3) +
    theme_light(base_size = 6)

ggsave(
    plot = p, filename = "UN_SDG_10-14-15.png", path = "figures/", device = "png",
    width = 7, height = 5, dpi = 400, bg = "white"
)

p + theme_light(base_size = 12)
```

### 2022-03-02 Trilemma_clustering.Rmd
data explorations for SDGs datasets

@@
     geom_vline(xintercept = 1.68, linetype = 2, color = "gray50") +
-    geom_density2d() +
-    geom_point(aes(color = gni_log)) +
-    geom_path(aes(group = country), size = 0.15) +
+    geom_density2d(size = 0.25, color = "#446B9C") +
+    geom_point(aes(color = gni_log), size = 0.5) +
+    geom_path(aes(group = country), size = 0.1) +
     scico::scale_color_scico("GNI [log]") +
     labs(y = "Gini", x = "Ecological footprint") +
-    theme_light(base_size = 7) +
-    theme(legend.position = "top", panel.grid = element_blank())
+    theme_light(base_size = 6) +
+    theme(legend.position = "top", panel.grid = element_blank(),
+          legend.key.width = unit(7, "mm"), legend.key.height = unit(2,"mm"))
 
@@
     ggplot(aes(gni2, gini)) +
-    geom_density2d() +
-    geom_point(aes(color = EFconsPerCap)) +
-    geom_path(aes(group = country), size = 0.15) +
+    geom_density2d(size = 0.25, color = "#446B9C") +
+    geom_point(aes(color = EFconsPerCap), size = 0.5) +
+    geom_path(aes(group = country), size = 0.1) +
     geom_hline(yintercept = 30, linetype = 2, color = "gray50") +
@@
     scale_x_log10() +
-    theme_light(base_size = 7) +
-    theme(legend.position = "top", panel.grid = element_blank())
+    theme_light(base_size = 6) +
+    theme(legend.position = "top", panel.grid = element_blank(),
+          legend.key.width = unit(7, "mm"), legend.key.height = unit(3,"mm"))
 
@@
     ggplot(aes(gni2, EFconsPerCap)) +
-    geom_density2d() +
-    geom_point(aes(color = gini)) +
-    geom_path(aes(group = country), size = 0.15) +
+    geom_density2d(size = 0.25, color = "#446B9C") +
+    geom_point(aes(color = gini), size =0.5) +
+    geom_path(aes(group = country), size = 0.1) +
     scico::scale_color_scico("Gini") +
@@
     geom_vline(xintercept = (12546), linetype = 2, color = "gray50") +
-    theme_light(base_size = 7) +
-    theme(legend.position = "top", panel.grid = element_blank())
+    theme_light(base_size = 6) +
+    theme(legend.position = "top", panel.grid = element_blank(),
+          legend.key.width = unit(7, "mm"), legend.key.height = unit(3,"mm"))
 
@@
               sd_GNI = sd(gni_log, na.rm = TRUE),
-              magnitude_m = mean(magni, na.rm = TRUE), 
-              theta_m = mean(theta, na.rm = TRUE),
-              gamma_m = mean(gamma, na.rm = TRUE),
+              #magnitude_m = mean(magni, na.rm = TRUE),
+              #theta_m = mean(theta, na.rm = TRUE),
+              #gamma_m = mean(gamma, na.rm = TRUE),
               magnitude_a = sum(magni, na.rm = TRUE),
               theta_a = sum(theta, na.rm = TRUE),
-              gamma_a = sum(gamma, na.rm = TRUE))
+              gamma_a = sum(gamma, na.rm = TRUE)
+              )
 
@@
 clust_num <- NbClust::NbClust( 
-    data = df_summary %>% as.data.frame() %>% 
-        select(where(is.numeric)),
+    data = (df_summary %>% as.data.frame() %>% 
+        select(where(is.numeric)) ),
     #distance = "maximum",
@@
 
-```{r}
+```{r warning = FALSE, message=FALSE}
+
+library(clValid)
+library(mclust)
 ## Stability & Internal validation
@@
         as.matrix(),
-    nClust = c(2:7),
-    clMethods = c("hierarchical", "kmeans",  #"som", "model", "diana", "fanny",
-                 "sota", "pam", "clara", "agnes"),
+    nClust = 3,
+    clMethods = c(
+        "hierarchical", "kmeans",  "som", 
+        "model", "diana", "sota", "pam", "clara", "agnes"),
     validation = c('stability', "internal"),
     #metric = "manhattan",
-    method = "ward.D2",
+    method = "ward",
     verbose = FALSE
@@
 ```{r}
-stab@clusterObjs$pam$`2`$clustering
+# stab@clusterObjs$pam$`2`$clustering
+
+stab@clusterObjs$kmeans$`3`$cluster
+
+stab@clusterObjs$diana |> plot()
 ```
@@
 df_summary |> 
-    mutate(clus = stab@clusterObjs$pam$`2`$clustering) |> 
+    mutate(clus = stab@clusterObjs$kmeans$`3`$cluster ) |> 
     mutate(country = as_factor(country)) |> 
-    mutate(country = fct_reorder2(country, magnitude, Gini)) |> 
-    pivot_longer(cols = sd_magnitude:gamma, names_to = "vars", values_to = "vals") |> 
+    mutate(country = fct_reorder2(country, magnitude_a, Gini)) |> 
+    pivot_longer(
+        cols = sd_magnitude:gamma_a, names_to = "vars", 
+        values_to = "vals") |> 
     ggplot(aes(vals, country)) +
@@
 
-```{r}
+```{r message = FALSE, warning = FALSE, fig.width=5, fig.height=5}
 df_summary |>
@@
 
-pc1 <- ggbiplot::ggbiplot(pc, groups = as.character(stab@clusterObjs$pam$`2`$clustering),
+pc1 <- ggbiplot::ggbiplot(
+    pc, groups = as.character(stab@clusterObjs$kmeans$`3`$cluster),
     ellipse = TRUE, labels = df_summary$country_code, labels.size = 2.5,
-         varname.size = 4) +
-    scale_color_manual("Groups", values = c("orange", "purple")) +
-    theme_light(base_size = 18) +
-    theme(legend.position = c(0.2, 0.2), panel.grid = element_blank())
+    varname.size = 3) +
+    scale_color_manual("Groups", values = c("#73B3A3","#FEA621","#5BA4CA")) +
+    theme_light(base_size = 6) +
+    theme(legend.position = c(0.2, 0.2), panel.grid = element_blank(),
+          legend.text = element_text(size = 6), legend.title = element_text(size = 6),
+          legend.key.width = unit(3,"mm"), legend.key.height = unit(2, "mm"))
 
 pc1
+
+ggsave(
+    filename = "pca_trilema.png", path = "figures/", device = "png",
+    width = 4, height = 4, dpi = 300, bg = "white", 
+    plot = pc1
+)
+
 ```
@@
 Note to self: It does not make sense to plot the average distance traversed in the space per country.
-Take Luxemburg as example, it started and ended verly close by.
+Take Luxemburg as example, it started and ended very close by.
 The absolute distance is what I'm calculating with the euclidian distance on that plane, then the first-diff values are just increases or decreases with respect to initial conditions.
-Averages are then misleading, because on average Luxemburg moved a lot (in circles).
+Averages are then misleading, because on average Luxembourg moved a lot (in circles).
 
@@
     ) |> 
-    mutate(Group = as.character(stab@clusterObjs$pam$`2`$clustering)) |> 
+    mutate(Group = as.character(stab@clusterObjs$kmeans$`3`$cluster)) |> 
     left_join(
@@
         aes(x = 0, y = 0, xend = EF, yend = Gini, color = as.factor(Group)),
-        arrow = arrow(length = unit(0.15, "cm")), size = 0.15) +
+        arrow = arrow(length = unit(0.1, "cm")), show.legend = FALSE, size = 0.15) +
     geom_text(
-        aes(x = EF *1.1, y = Gini * 1.1, label = country_code, colour = Group), size = 2) +
-    scale_color_manual("Groups", values = c("orange", "purple")) +
+        aes(x = EF *1.1, y = Gini * 1.1, label = country_code, colour = Group), 
+        size = 1, show.legend = FALSE) +
+    scale_color_manual("Groups", values = c("#73B3A3","#FEA621","#5BA4CA")) +
     labs(x = "Ecological Footprint", y = "Gini") +
     theme_light(base_size = 6) +
-    theme(panel.grid = element_blank(), legend.position = c(0.8,0.2))
+    theme(panel.grid = element_blank(), legend.position = c(0.8,0.2), 
+          legend.text = element_text(size = 4), legend.title = element_text(size =4),
+          legend.key.width = unit(2,"mm"), legend.key.height = unit(2, "mm"))
 
@@
 
-c1 + annotation_custom(
+f1 <- c1 + annotation_custom(
         grob = ggplotGrob(a1),
@@
     )
+
+# ggsave(
+#     filename = "Gini_EF.png", path = "figures/", device = "png",
+#     width = 4, height = 4, dpi = 300, bg = "white", 
+#     plot = f1
+# )
+
+ ggplotly(c1, tooltip=c("country", "EFconsPerCap", "gini", "gni_log"))
+
 ```
 
-```{r fig.width=2, fig.height=2}
+```{r fig.width=3, fig.height=3}
 a2 <- df_dat |> 
@@
         aes(x = 0, y = 0, xend = GNI, yend = Gini, 
-            color = as.factor(Group)), 
-        arrow = arrow(length = unit(0.15, "cm")), size = 0.15) +
+            color = as.factor(Group)), show.legend = FALSE,
+        arrow = arrow(length = unit(0.1, "cm")), size = 0.15) +
     geom_text(aes(x = GNI * 1.1, y = Gini *1.1, 
-                  label = country_code, colour = Group), size = 2) +
-    scale_color_manual("Groups", values = c("orange", "purple")) +
-    theme_light(base_size = 6) +
-    theme(panel.grid = element_blank(), legend.position = c(0.8,0.2))
+                  label = country_code, colour = Group), 
+              size = 1, show.legend = FALSE) +
+    scale_color_manual("Groups", values = c("#73B3A3","#FEA621","#5BA4CA")) +
+    theme_light(base_size = 6) + labs(x = "GNI [log]") +
+    theme(panel.grid = element_blank(), legend.position = c(0.8,0.2),
+          legend.text = element_text(size = 4), legend.title = element_text(size = 4),
+          legend.key.width = unit(2,"mm"), legend.key.height = unit(2, "mm"))
 
-c2 + annotation_custom(
+f2 <- c2 + annotation_custom(
         grob = ggplotGrob(a2),
-        xmin = 1e+01, xmax = 1e+03, ymin = 40, ymax = 60
+        xmin = 0.8, xmax = 2.5, ymin = 40, ymax = 62
     )
 
-a2
+# ggsave(
+#     filename = "Gini_GNI.png", path = "figures/", device = "png",
+#     width = 4, height = 4, dpi = 300, bg = "white", 
+#     plot = f2
+# )
+
+```
+
+```{r fig.height=3, fig.width=3}
+a3 <- df_dat |> 
+    as_tibble() |> 
+    group_by(country) |> 
+    dplyr::summarize(
+        EF = sum(d_EF, na.rm = TRUE),
+        Gini = sum(d_gini, na.rm = TRUE),
+        GNI = sum(d_gni, na.rm = TRUE)
+    ) |> 
+    mutate(Group = as.character(stab@clusterObjs$kmeans$`3`$cluster)) |> 
+    left_join(
+        dat |> select(country, country_code) |> 
+            filter(!is.na(country_code)) |> unique()) |> 
+    ggplot(aes(GNI, EF)) +
+    geom_segment(
+        aes(x = 0, y = 0, xend = GNI, yend = EF, 
+            color = as.factor(Group)), show.legend = FALSE, 
+        arrow = arrow(length = unit(0.1, "cm")), size = 0.15) +
+    geom_text(aes(x = GNI * 1.1, y = EF *1.1, 
+                  label = country_code, colour = Group), 
+              size = 1, show.legend = FALSE) +
+    scale_color_manual("Groups", values = c("#73B3A3","#FEA621","#5BA4CA")) +
+    theme_light(base_size = 6) + labs(y = "Ecological footprint") +
+    theme(panel.grid = element_blank(), legend.position = c(0.8,0.8),
+          legend.text = element_text(size = 4), legend.title = element_text(size = 4),
+          legend.key.width = unit(2,"mm"), legend.key.height = unit(2, "mm"))
+
+f3 <- c3 + annotation_custom(
+        grob = ggplotGrob(a3),
+        xmin = 0.8, xmax = 2.8, ymin = 8, ymax = 18
+    )
+
+# ggsave(
+#     filename = "EF_GNI.png", path = "figures/", device = "png",
+#     width = 4, height = 4, dpi = 300, bg = "white", 
+#     plot = f3
+# )
+
+```
+
+```{r}
+
+# ggsave(
+#     filename = "trilema_ensamble.png", device = "png", path = "figures/",
+#     width = 7.5, height = 3, bg = "white", dpi = 300, 
+#     plot = f1 + labs(tag = "A") + f2 + labs(tag = "B") +  f3 + labs(tag = "C") 
+# )
 ```

### 2022-03-02 inequality-SDGs.Rmd
data explorations for SDGs datasets

@@
 key |> pull(topic) |> unique()
+
+countries <- read_csv(
+    file = "~/Documents/Projects/DATA/WorldBank/SDG_csv/SDGCountry.csv") |> 
+    janitor::clean_names()
+
+countries_list <- countries |> filter(!is.na(currency_unit)) |> pull(country_code)
+
 ```
+
+```{r}
+dat$indicator_name |> unique() |> length()
+```
+Extract environmental variables only
+
+```{r}
+env_topics <- key$topic |> str_subset("Environment")
+
+env_vars <- key |> 
+    filter(topic %in% env_topics) |> 
+    pull(indicator_name) |> 
+    unique()
+```
+
+```{r}
+n_countries <- dat |> filter(country_code %in% countries_list) |> 
+    pull(country_code) |> unique() |> length()
+
+p <- dat |>
+    filter(indicator_name %in% env_vars, country_code %in% countries_list) |> 
+    select(-x35) |> 
+    pivot_longer(cols = x1990:last_col(), values_to = "value", names_to = "year") |> 
+    mutate(year = str_remove(year, "x"), year = as.numeric(year)) |> 
+    group_by(indicator_name, year) |> 
+    mutate(nas = is.na(value)) |> 
+    summarise(missing = sum(nas) / n_countries) |> 
+    ggplot(aes(year, indicator_name)) +
+    geom_tile(aes(fill = missing)) +
+    scale_fill_viridis_c() +
+    theme_light(base_size = 6)
+
+ggsave(
+    plot = p, filename = "WB_environment_SDGs.png", path = "figures/", device = "png",
+    width = 6, height = 4, dpi = 400, bg = "white"
+)
+
+p + theme_light(base_size = 8)
+```
+And the inequality variables
+
+```{r}
+ineq_topics <- c("Financial Sector: Access", )
+
+key$topic |> unique()
+```
+
+## To-dos
+
+Wait until Maike and Tong revise the selection of variables.
+Take the analysis from there.
+Alternatively, select a couple of topics as example, just to explore how to deal with missing values and what is doable.
+Once one take missing values into account, the number of variables and countries get reduced quickly.

### 2022-04-14 Trilemma_clustering.Rmd
new figures for paper

@@
               #gamma_m = mean(gamma, na.rm = TRUE),
-              magnitude_a = sum(magni, na.rm = TRUE),
-              theta_a = sum(theta, na.rm = TRUE),
-              gamma_a = sum(gamma, na.rm = TRUE)
+              magnitude = sum(magni, na.rm = TRUE),
+              theta = sum(theta, na.rm = TRUE),
+              gamma = sum(gamma, na.rm = TRUE)
               )
@@
     pc, groups = as.character(stab@clusterObjs$kmeans$`3`$cluster),
-    ellipse = TRUE, labels = df_summary$country_code, labels.size = 2.5,
-    varname.size = 3) +
+    ellipse = TRUE, labels = df_summary$country_code, labels.size = 1.5,
+    varname.size = 2) +
     scale_color_manual("Groups", values = c("#73B3A3","#FEA621","#5BA4CA")) +
+    labs(tag = "A") +
     theme_light(base_size = 6) +
@@
     filename = "pca_trilema.png", path = "figures/", device = "png",
-    width = 4, height = 4, dpi = 300, bg = "white", 
+    width = 3, height = 3, dpi = 300, bg = "white",
     plot = pc1
@@
 Be aware of that when presenting and interpreting the results.
+
 ```{r fig.height=3,fig.width=3}
@@
 
- ggplotly(c1, tooltip=c("country", "EFconsPerCap", "gini", "gni_log"))
-
+# ggplotly(c1, tooltip=c("country", "EFconsPerCap", "gini", "gni_log"))
+f1
 ```
@@
 
-### Old stuff
-
-Here is an attempt with another package `kml` that take into consideration the fact that our data is longitudinal.
-However, looking into the documentation, it seems that it only handles a variable at the time.
-For the sake of testing:
+## Static
 
 ```{r}
-# library(kml)
+## ecological footprint
+ef <- ecofp %>%
+  filter(record == "EFConsPerCap") %>%
+  select(countryName, isoa2, year, value) %>% 
+  dplyr::rename(EFconsPerCap = value, country = countryName) |> 
+    group_by(country, isoa2) |> 
+    dplyr::summarize(
+        m_ef = mean(EFconsPerCap, na.rm = TRUE), 
+        sd_ef = sd(EFconsPerCap, na.rm = TRUE))
 
-test_dat <- df_dat %>% 
-  select(country, gini, year) %>% 
-  as_tibble() %>% 
-  pivot_wider(names_from = year, values_from = gini, names_prefix = "yr_") %>% 
-  as_data_frame() %>% 
-  kml::cld(idAll = "country", timeInData = 2:29)
+gini <- inq %>% 
+    filter(!is.na(gini_reported)) %>% 
+    select(country, isoa2 = c2, country_code = c3, year, gini_reported) %>% 
+    group_by(country, year, isoa2, country_code) %>% 
+    dplyr::summarize(gini_mean = mean(gini_reported, na.rm = TRUE)) |> 
+    ungroup() |> group_by(country, isoa2, country_code) |> 
+    dplyr::summarize(
+        m_gini = mean(gini_mean, na.rm = TRUE),
+        sd_gini = sd(gini_mean, na.rm = TRUE)
+    )
+
+gni <- dat1 %>% 
+      select(-last_col()) %>% 
+      filter(indicator_name == "GNI per capita (US$)") %>% 
+      pivot_longer(cols = x1990:x2019, names_to = "year", values_to = "value") %>%
+      mutate(year = str_remove(year, "x")) %>% 
+      mutate(year = as.numeric(year)) %>% 
+      mutate(ismissing = is.na(value)) %>% 
+      dplyr::rename(country = country_name, gni = value) |> 
+    group_by(country, country_code) |> 
+    dplyr::summarize(
+        m_gni = mean(gni, na.rm = TRUE),
+        sd_gni = sd(gni, na.rm = TRUE)
+    )
+
+df_all <- gini |> left_join(ef) |> left_join(gni)
+
+df_all <- df_all |> ungroup() |> 
+    filter(!is.na(sd_gini), !is.na(m_ef), !is.na(sd_ef), !is.na(m_gni), !is.na(sd_gni)) 
+```
+
+```{r}
+m <- "ward.D2" # minimize the total within-cluster variance
+
+clust_num <- NbClust::NbClust( 
+    data = (df_all %>% as.data.frame() %>% 
+        select(where(is.numeric)) ),
+    #distance = "maximum",
+    min.nc = 2, max.nc = 7, 
+    method = m, 
+    index = 'all')
+```
+
+```{r warning = FALSE, message=FALSE}
+
+library(clValid)
+library(mclust)
+## Stability & Internal validation
+stab <- clValid::clValid(
+    obj = df_all %>% as.data.frame() %>% 
+        select(where(is.numeric)) %>%
+        as.matrix(),
+    nClust = 2,
+    clMethods = c(
+        "hierarchical", "kmeans",  "som", 
+        "model", "diana", "sota", "pam", "clara", "agnes"),
+    validation = c('stability', "internal"),
+    #metric = "manhattan",
+    method = "ward",
+    verbose = FALSE
+) ## Hierarchical is the optimal method
+
+clValid::optimalScores(stab)
+```
+
+```{r}
+stab@clusterObjs$pam$`2`$clustering
 
 ```
-An error I don't understand.
+
+```{r}
+df_all |> 
+    mutate(clus = stab@clusterObjs$pam$`2`$clustering ) 
+```
+
+```{r}
+pc_all <- prcomp(df_all |> 
+                 select(where(is.numeric)) ,
+             center = TRUE, scale. = TRUE)
+
+pc2 <- ggbiplot::ggbiplot(
+    pc_all, groups = as.character(stab@clusterObjs$pam$`2`$clustering ),
+    ellipse = TRUE, labels = df_all$country_code, labels.size = 1.5,
+    varname.size = 2) +
+    scale_color_manual("Groups", values = c("#73B3A3","#FEA621")) +
+    labs(tag = "B") + 
+    theme_light(base_size = 6) +
+    theme(legend.position = c(0.2, 0.2), panel.grid = element_blank(),
+          legend.text = element_text(size = 6), legend.title = element_text(size = 6),
+          legend.key.width = unit(3,"mm"), legend.key.height = unit(2, "mm"))
+
+pc1 + pc2 + plot_layout(widths = 3, heights = 3)
+
+ggsave(
+    filename = "pca_trilemma.png", path = "figures/", device = "png",
+    width = 6, height = 3, dpi = 300, bg = "white",
+    plot = pc1+pc2
+)
+```
+
+### Maps
+
+```{r}
+library(sf)
+library(spData)
+```
+
+```{r}
+
+data(world)
+
+ef_map <-  world |>
+    filter(!is.na(pop)) |> 
+    left_join(ef, by = c("iso_a2" = "isoa2")) |> 
+    ggplot() +
+    geom_sf(aes(fill = m_ef), size = 0.05, color = "white") +
+    scico::scale_fill_scico(
+        "Mean Ecological Footprint",palette = "romaO",
+        guide = guide_colorbar(title.position = "top")) + 
+    labs(tag = "A") +
+    theme_void(base_size = 6) +
+    theme(legend.position = "bottom", legend.key.height = unit(2, "mm"), 
+          legend.key.width = unit(8, "mm"))
+
+gini_map <-  world |>
+    filter(!is.na(pop)) |> 
+    left_join(gini, by = c("iso_a2" = "isoa2")) |> 
+    ggplot() +
+    geom_sf(aes(fill = m_gini), size = 0.05, color = "white") +
+    scico::scale_fill_scico(
+        "Mean Gini",palette = "berlin",
+        guide = guide_colorbar(title.position = "top")) + 
+    labs(tag = "B") +
+    theme_void(base_size = 6) +
+    theme(legend.position = "bottom", legend.key.height = unit(2, "mm"), 
+          legend.key.width = unit(8, "mm"))
+
+gni_map <-  world |>
+    left_join(gini |> dplyr::select(isoa2, country_code),  
+              by = c("iso_a2" = "isoa2") ) |> 
+    left_join(gni |> ungroup() |> select(-country)) |> 
+    filter(!is.na(pop)) |>
+    ggplot() +
+    geom_sf(aes(fill = m_gni), size = 0.05, color = "white") +
+    scico::scale_fill_scico(
+        "Mean Gross National Income",palette = "batlowW",
+        guide = guide_colorbar(title.position = "top")) + 
+    labs(tag = "C") +
+    theme_void(base_size = 6) +
+    theme(legend.position = "bottom", legend.key.height = grid::unit(2, "mm"), 
+          legend.key.width = grid::unit(8, "mm"))
+
+ef_map + gini_map + gni_map
+
+ggsave(
+    filename = "maps_trilemma.png", path = "figures/", device = "png",
+    width = 6, height = 1.5, dpi = 400, bg = "white",
+    plot = ef_map + gini_map + gni_map
+)
+
+```

### 2022-09-27 Trilemma_clustering.Rmd
subnational analysis of SDG + back ups

@@
 fls <- list.files("data/ecofootprint/")
-ecofp <- map(fls, function(x) {
+ecofp <- purrr::map(fls, function(x) {
     jsonlite::read_json(
@@
     geom_path(aes(group = country), size = 0.1) +
-    scico::scale_color_scico("GNI [log]") +
-    labs(y = "Gini", x = "Ecological footprint") +
+    scico::scale_color_scico("Prosperity (GNI [log])") +
+    labs(y = "Inequality (Gini)", x = "Ecological footprint") +
     theme_light(base_size = 6) +
@@
     scico::scale_color_scico("Ecological footprint",palette = "berlin") +
-    labs(x= "GNI [log]", y = "Gini") +
+    labs(x= "Prosperity (GNI [log])", y = "Inequality (Gini)") +
     scale_x_log10() +
@@
     geom_path(aes(group = country), size = 0.1) +
-    scico::scale_color_scico("Gini") +
-    labs(x = "GNI", y = "Ecological footprint") +
+    scico::scale_color_scico("Inequality (Gini)") +
+    labs(x = "Prosperity (GNI)", y = "Ecological footprint") +
     scale_x_log10()+
@@
     pc, groups = as.character(stab@clusterObjs$kmeans$`3`$cluster),
-    ellipse = TRUE, labels = df_summary$country_code, labels.size = 1.5,
-    varname.size = 2) +
+    ellipse = TRUE, labels = df_summary$country_code, labels.size = 1,
+    varname.size = 1.5) +
     scale_color_manual("Groups", values = c("#73B3A3","#FEA621","#5BA4CA")) +
     labs(tag = "A") +
-    theme_light(base_size = 6) +
+    theme_light(base_size = 5) +
     theme(legend.position = c(0.2, 0.2), panel.grid = element_blank(),
-          legend.text = element_text(size = 6), legend.title = element_text(size = 6),
+          legend.text = element_text(size = 5), legend.title = element_text(size = 5),
           legend.key.width = unit(3,"mm"), legend.key.height = unit(2, "mm"))
@@
 
-ggsave(
-    filename = "pca_trilema.png", path = "figures/", device = "png",
-    width = 3, height = 3, dpi = 300, bg = "white",
-    plot = pc1
-)
+# ggsave(
+#     filename = "pca_trilema.png", path = "figures/", device = "png",
+#     width = 3, height = 3, dpi = 300, bg = "white",
+#     plot = pc1
+# )
 
@@
     scale_color_manual("Groups", values = c("#73B3A3","#FEA621","#5BA4CA")) +
-    labs(x = "Ecological Footprint", y = "Gini") +
+    labs(x = "Ecological Footprint", y = "Inequality (Gini)") +
     theme_light(base_size = 6) +
@@
     scale_color_manual("Groups", values = c("#73B3A3","#FEA621","#5BA4CA")) +
-    theme_light(base_size = 6) + labs(x = "GNI [log]") +
+    theme_light(base_size = 6) + labs(x = "Prosperity (GNI [log])", y = "Inequality (Gini)") +
     theme(panel.grid = element_blank(), legend.position = c(0.8,0.2),
@@
     scale_color_manual("Groups", values = c("#73B3A3","#FEA621","#5BA4CA")) +
-    theme_light(base_size = 6) + labs(y = "Ecological footprint") +
+    theme_light(base_size = 6) + labs(y = "Ecological footprint", x = "Prosperity (GNI)") +
     theme(panel.grid = element_blank(), legend.position = c(0.8,0.8),
@@
 
-df_all <- gini |> left_join(ef) |> left_join(gni)
+df_all2 <- gini |> left_join(ef) |> left_join(gni)
 
-df_all <- df_all |> ungroup() |> 
+df_all2 <- df_all |> ungroup() |> 
     filter(!is.na(sd_gini), !is.na(m_ef), !is.na(sd_ef), !is.na(m_gni), !is.na(sd_gni)) 
@@
 
-clust_num <- NbClust::NbClust( 
-    data = (df_all %>% as.data.frame() %>% 
+clust_num2 <- NbClust::NbClust( 
+    data = (df_all2 %>% as.data.frame() %>% 
         select(where(is.numeric)) ),
@@
 ## Stability & Internal validation
-stab <- clValid::clValid(
-    obj = df_all %>% as.data.frame() %>% 
+stab2 <- clValid::clValid(
+    obj = df_all2 %>% as.data.frame() %>% 
         select(where(is.numeric)) %>%
@@
 ```{r}
-stab@clusterObjs$pam$`2`$clustering
+stab2@clusterObjs$pam$`2`$clustering
 
@@
 ```{r}
-df_all |> 
-    mutate(clus = stab@clusterObjs$pam$`2`$clustering ) 
+df_all2 |> 
+    mutate(clus = stab2@clusterObjs$pam$`2`$clustering ) 
 ```
@@
 ```{r}
-pc_all <- prcomp(df_all |> 
+pc_all <- prcomp(df_all2 |> 
                  select(where(is.numeric)) ,
@@
 pc2 <- ggbiplot::ggbiplot(
-    pc_all, groups = as.character(stab@clusterObjs$pam$`2`$clustering ),
-    ellipse = TRUE, labels = df_all$country_code, labels.size = 1.5,
-    varname.size = 2) +
+    pc_all, groups = as.character(stab2@clusterObjs$pam$`2`$clustering ),
+    ellipse = TRUE, labels = df_all2$country_code, labels.size = 1,
+    varname.size = 1.5) +
     scale_color_manual("Groups", values = c("#73B3A3","#FEA621")) +
     labs(tag = "B") + 
-    theme_light(base_size = 6) +
+    theme_light(base_size = 5) +
     theme(legend.position = c(0.2, 0.2), panel.grid = element_blank(),
-          legend.text = element_text(size = 6), legend.title = element_text(size = 6),
+          legend.text = element_text(size = 5), legend.title = element_text(size = 5),
           legend.key.width = unit(3,"mm"), legend.key.height = unit(2, "mm"))
@@
 
-ggsave(
-    filename = "pca_trilemma.png", path = "figures/", device = "png",
-    width = 6, height = 3, dpi = 300, bg = "white",
-    plot = pc1+pc2
-)
+# ggsave(
+#     filename = "pca_trilemma.png", path = "figures/", device = "png",
+#     width = 6, height = 3, dpi = 300, bg = "white",
+#     plot = pc1+pc2
+# )
 ```
@@
         guide = guide_colorbar(title.position = "top")) + 
-    labs(tag = "A") +
-    theme_void(base_size = 6) +
-    theme(legend.position = "bottom", legend.key.height = unit(2, "mm"), 
-          legend.key.width = unit(8, "mm"))
+    labs(tag = "C") +
+    theme_void(base_size = 5) +
+    theme(legend.position = "bottom", legend.key.height = unit(1, "mm"), 
+          legend.key.width = unit(5, "mm"))
 
@@
         guide = guide_colorbar(title.position = "top")) + 
-    labs(tag = "B") +
-    theme_void(base_size = 6) +
-    theme(legend.position = "bottom", legend.key.height = unit(2, "mm"), 
-          legend.key.width = unit(8, "mm"))
+    labs(tag = "D") +
+    theme_void(base_size = 5) +
+    theme(legend.position = "bottom", legend.key.height = unit(1, "mm"), 
+          legend.key.width = unit(5, "mm"))
 
@@
         guide = guide_colorbar(title.position = "top")) + 
-    labs(tag = "C") +
-    theme_void(base_size = 6) +
-    theme(legend.position = "bottom", legend.key.height = grid::unit(2, "mm"), 
-          legend.key.width = grid::unit(8, "mm"))
+    labs(tag = "E") +
+    theme_void(base_size = 5) +
+    theme(legend.position = "bottom", legend.key.height = grid::unit(1, "mm"), 
+          legend.key.width = grid::unit(5, "mm"))
 
 ef_map + gini_map + gni_map
+# 
+# ggsave(
+#     filename = "maps_trilemma.png", path = "figures/", device = "png",
+#     width = 6, height = 1.5, dpi = 400, bg = "white",
+#     plot = ef_map + gini_map + gni_map
+# )
+
+```
+
+```{r}
 
 ggsave(
-    filename = "maps_trilemma.png", path = "figures/", device = "png",
-    width = 6, height = 1.5, dpi = 400, bg = "white",
-    plot = ef_map + gini_map + gni_map
+    filename = "pca_maps.png", path = "figures/", device = "png",
+    width = 5, height = 4, dpi = 400, bg = "white",
+    plot = (pc1 / pc2 ) | (ef_map / gini_map / gni_map) + 
+    plot_annotation(tag_levels = "A") + 
+    plot_layout(widths = c(1,4), tag_level = "new")
 )

### 2022-09-27 inequality-SDGs.Rmd
subnational analysis of SDG + back ups

--- 2022-02-10 inequality-SDGs.Rmd
@@
 ```{r}
-key <- read_csv(
-    file = "~/Documents/Projects/DATA/WorldBank/SDG_csv/SDGSeries.csv"
-) |> janitor::clean_names()
+# key <- read_csv(
+#     file = "~/Documents/Projects/DATA/WorldBank/SDG_csv/SDGSeries.csv"
+# ) |> janitor::clean_names()
+key <- googlesheets4::read_sheet(
+    "https://docs.google.com/spreadsheets/d/1T6rZ5T1qL4BPDL5oatMf3y8lLzhnIDvxIvkbJeBkdUc/edit?usp=sharing", sheet = 1
+)
 
+key <- key |> janitor::clean_names()
 key |> pull(topic) |> unique()
+
+countries <- read_csv(
+    file = "~/Documents/Projects/DATA/WorldBank/SDG_csv/SDGCountry.csv") |> 
+    janitor::clean_names()
+
+countries_list <- countries |> filter(!is.na(currency_unit)) |> pull(country_code)
+
 ```
+
+```{r}
+dat$indicator_name |> unique() |> length()
+```
+Extract environmental variables only
+
+```{r}
+# env_topics <- key$topic |> str_subset("Environment")
+# 
+# env_vars <- key |> 
+#     filter(topic %in% env_topics) |> 
+#     pull(indicator_name) |> 
+#     unique()
+maike_vars <- key |> 
+    filter(!is.na(maike_selection)) |> 
+    pull(indicator_name) |> 
+    unique()
+```
+
+```{r}
+n_countries <- dat |> filter(country_code %in% countries_list) |> 
+    pull(country_code) |> unique() |> length()
+
+p <- dat |>
+    filter(indicator_name %in% maike_vars, country_code %in% countries_list) |> 
+    select(-x35) |> 
+    pivot_longer(cols = x1990:last_col(), values_to = "value", names_to = "year") |> 
+    mutate(year = str_remove(year, "x"), year = as.numeric(year)) |> 
+    group_by(indicator_name, year) |> 
+    mutate(nas = is.na(value)) |> 
+    summarise(missing = sum(nas) / n_countries) |> 
+    ggplot(aes(year, indicator_name)) +
+    geom_tile(aes(fill = missing)) +
+    scale_fill_viridis_c() +
+    theme_light(base_size = 6)
+
+# ggsave(
+#     plot = p, filename = "WB_environment_SDGs.png", path = "figures/", device = "png",
+#     width = 6, height = 4, dpi = 400, bg = "white"
+# )
+
+p + theme_light(base_size = 8)
+```
+
+```{r}
+other_vars <- key |> 
+    filter(!is.na(include)) |> 
+    pull(indicator_name) |> 
+    unique()
+
+p <- dat |>
+    filter(indicator_name %in% other_vars, country_code %in% countries_list) |> 
+    select(-x35) |> 
+    pivot_longer(cols = x1990:last_col(), values_to = "value", names_to = "year") |> 
+    mutate(year = str_remove(year, "x"), year = as.numeric(year)) |> 
+    group_by(indicator_name, year) |> 
+    mutate(nas = is.na(value)) |> 
+    summarise(missing = sum(nas) / n_countries) |> 
+    ggplot(aes(year, indicator_name)) +
+    geom_tile(aes(fill = missing)) +
+    scale_fill_viridis_c() +
+    theme_light(base_size = 6)
+
+p + theme_light(base_size = 8)
+```
+
+And the inequality variables
+
+```{r}
+# ineq_topics <- c("Financial Sector: Access", )
+
+key$topic |> unique()
+```
+
+## To-dos
+
+Wait until Maike and Tong revise the selection of variables.
+Take the analysis from there.
+Alternatively, select a couple of topics as example, just to explore how to deal with missing values and what is doable.
+Once one take missing values into account, the number of variables and countries get reduced quickly.
+
+Combining both options:
+
+```{r}
+other_vars <- key |> 
+    filter(!is.na(include) | !is.na(maike_selection)) |> 
+    pull(indicator_name) |> 
+    unique()
+
+p <- dat |>
+    filter(indicator_name %in% other_vars, country_code %in% countries_list) |> 
+    select(-x35) |> 
+    pivot_longer(cols = x1990:last_col(), values_to = "value", names_to = "year") |> 
+    mutate(year = str_remove(year, "x"), year = as.numeric(year)) |> 
+    group_by(indicator_name, year) |> 
+    mutate(nas = is.na(value)) |> 
+    summarise(missing = sum(nas) / n_countries) |> 
+    ungroup() |> group_by(indicator_name) |> 
+    summarize(mean_missing = mean(missing)) |> 
+    filter(mean_missing < 0.3) |> 
+    ggplot(aes(mean_missing, indicator_name)) +
+    geom_point() +
+    #scale_fill_viridis_c() +
+    theme_light(base_size = 6)
+
+p + theme_light(base_size = 8)
+
+vars <- p$data$indicator_name
+
+```
+### Correlations
+
+Probably need to log-transform cereal yield and forest area.
+
+```{r}
+dat |>
+    filter(indicator_name %in% vars, country_code %in% countries_list) |> 
+    select(-x35) |> 
+    pivot_longer(cols = x1990:last_col(), values_to = "value", names_to = "year") |> 
+    mutate(year = str_remove(year, "x"), year = as.numeric(year)) |> 
+    select(-indicator_code) |> 
+    pivot_wider(values_from = value, names_from = indicator_name) |> 
+    GGally::ggpairs(
+        columns = 4:12 ,
+    lower = list(continuous = GGally::wrap("points", alpha = 0.5, size = 0.5)),
+    diag = list(continuous = GGally::wrap("densityDiag", alpha = 0.2))
+    # upper = list(continuous = GGally::wrap("cor", size = 2))
+    ) + theme_light(base_size = 6)
+```
+Interpolate missing values
+
+```{r}
+df_dat <- dat |>
+    filter(indicator_name %in% vars, country_code %in% countries_list) |> 
+    select(-x35) |> 
+    pivot_longer(cols = x1990:last_col(), values_to = "value", names_to = "year") |> 
+    mutate(year = str_remove(year, "x"), year = as.numeric(year)) |> 
+    select(-indicator_code) |> 
+    pivot_wider(values_from = value, names_from = indicator_name) %>% 
+    group_by(country_code) %>% 
+    mutate(across(.cols = vars, .fns = na_interpolation, option = "spline"))
+
+df_dat
+```
+
+## UN SDGs
+
+```{r}
+un_countries <- read_csv2(file = "~/Documents/Projects/DATA/SDGs_UNStats/UNSD — Methodology.csv")
+sdg <- read_csv(file = "~/Documents/Projects/DATA/SDGs_UNStats/20180817102044132_juan.rocha@su.se_data.csv") |> janitor::clean_names()
+
+sdg <- sdg |> 
+    left_join(
+        select(un_countries, country = `Country or Area`, starts_with("ISO")),
+        by = c("geo_area_name" = "country")) |> 
+    filter(value != "-")|> 
+    mutate(value = as.numeric(value))
+
+sdg 
+```
+
+Why are there more than one obs per country per year per indicator???
+```{r}
+sdg |> 
+    filter(`ISO-alpha3 Code` %in% countries_list) |> 
+    group_by(geo_area_code, indicator, time_period) |> 
+    summarize(n = n()) |> 
+    filter(n>19) |>  # 19 is the max number of years
+    arrange(desc(n))
+
+sdg |> 
+    filter(geo_area_code == 40, indicator == "5.4.1", time_period == 2009)
+
+```
+Answer: there are indicators that are calculated per age group, sex and rural / urban divide. They need to be processed to a single number per year. To discuss options:
+
+- What dimension to use?: gender, age, rural / urban / all
+- What summary statistic to use: mean, var, gini?
+
+```{r}
+n_countries2 <- sdg |> 
+    filter(`ISO-alpha3 Code` %in% countries_list) |>
+    pull(geo_area_name) |> unique() |> length() # number of countries, but here are also non-countries categories like continents, etc.
+
+sdg |> 
+    rename(indicator_name = series_description, year = time_period, iso3 =`ISO-alpha3 Code` ) |> 
+    filter(iso3 %in% countries_list, !is.na(value)) |> # use only countries
+    filter(goal == 10) |> 
+    group_by(indicator, indicator_name, year, iso3) |> 
+    summarize(mean_val = mean(value, na.rm = TRUE)) |> 
+    pivot_wider(names_from = year, values_from = mean_val) |> 
+    pivot_longer(cols = starts_with("2"), names_to = "year", values_to = "value",
+                 names_transform = list(year = as.numeric, value = as.numeric)) |> 
+    mutate(nas = is.na(value)) |> 
+    group_by(indicator_name, year) |> 
+    summarise(missing = sum(nas) / n_countries2) |> 
+    ggplot(aes(year, indicator_name)) +
+    geom_tile(aes(fill = missing)) +
+    scale_fill_viridis_c() +
+    labs(title = "SDG10: Inequality indicators") +
+    theme_light(base_size = 8)
+
+# sdg |> 
+#     filter(series_description == "Proportion of voting rights of developing countries in international organizations, by organization (%)")
+
+```
+
+```{r}
+sdg |> 
+    rename(indicator_name = series_description, year = time_period, iso3 =`ISO-alpha3 Code` ) |> 
+    filter(iso3 %in% countries_list, !is.na(value)) |> # use only countries
+    filter(goal == 14) |> 
+    group_by(indicator, indicator_name, year, iso3) |> 
+    summarize(mean_val = mean(value, na.rm = TRUE)) |> 
+    pivot_wider(names_from = year, values_from = mean_val) |> 
+    pivot_longer(cols = starts_with("2"), names_to = "year", values_to = "value",
+                 names_transform = list(year = as.numeric, value = as.numeric)) |> 
+    mutate(nas = is.na(value)) |> 
+    group_by(indicator_name, year) |> 
+    summarise(missing = sum(nas) / n_countries2) |> 
+    ggplot(aes(year, indicator_name)) +
+    geom_tile(aes(fill = missing)) +
+    scale_fill_viridis_c() +
+    labs(title = "SDG14: Life under water") +
+    theme_light(base_size = 8)
+```
+
+```{r}
+sdg |> 
+    rename(indicator_name = series_description, year = time_period, iso3 =`ISO-alpha3 Code` ) |> 
+    filter(iso3 %in% countries_list, !is.na(value)) |> # use only countries
+    filter(goal == 15) |> 
+    group_by(indicator, indicator_name, year, iso3) |>
+    summarize(mean_val = mean(value, na.rm = TRUE)) |>
+    pivot_wider(names_from = year, values_from = mean_val) |> 
+    pivot_longer(cols = starts_with("2"), names_to = "year", values_to = "value",
+                 names_transform = list(year = as.numeric, value = as.numeric)) |> 
+    mutate(nas = is.na(value)) |> 
+    group_by(indicator_name, year) |> 
+    summarise(missing = sum(nas) / n_countries2) |> 
+     mutate(indicator_name = str_trunc(indicator_name, width = 30, side = "right")) |> 
+    ggplot(aes(year, indicator_name)) +
+    geom_tile(aes(fill = missing)) +
+    scale_fill_viridis_c() +
+    labs(title = "SDG15: Life on land") +
+    theme_light(base_size = 8)
+
+sdg|> 
+    rename(indicator_name = series_description, year = time_period, iso3 =`ISO-alpha3 Code` ) |> 
+    filter(iso3 %in% countries_list, !is.na(value)) |> # use only countries
+    filter(goal == 15) |> 
+    pull(indicator_name) |> unique()
+
+```
+
+## Notes:
+
+Select the variables for which you have time series. Repeat the trajectories analysis. 
+Do static analysis with all SDGs
+
+```{r}
+ctr <- sdg |> pull(geo_area_name) |> unique() |> length()
+sdg <- sdg |> select(-c(age:type_of_speed))
+
+sdg <- sdg |> 
+    mutate(value = as.numeric(value)) |> 
+    rename(year = time_period)
+yrs <- sdg$year |> range() |> diff()
+sdg <- sdg |> unique()
+
+dat2 <- sdg |> 
+    filter(goal == 10 | goal == 14 | goal == 15 ) |> 
+    group_by(indicator, geo_area_code, year, geo_area_name, goal, target, series_code, series_description) |> 
+    summarize(value = mean(value))
+
+## If I add series description it duplicate entries, possibly due to misspellings
+#test
+dat2 |>
+    unique() |>
+    add_count() |>
+    filter(n>1)
+
+dat2 |> 
+    add_count() |> 
+    ungroup() |> 
+    group_by(geo_area_name,  series_code) |> 
+    summarize(yrs_cov = sum(n)) |> 
+    filter(yrs_cov > 20)
+
+sdg |> filter(series_code == "DC_ODA_BDVDL")
+```

### 2022-09-27 01-load_data.R

@@ -1,0 +1,96 @@
+library(tidyverse)
+library(naniar)
+library(plotly)
+library(patchwork)
+
+# time series
+library(tsibble)
+library(fable)
+library(imputeTS)
+library(slider)
+
+
+## Load datasets
+ecofp <- list()
+fls <- list.files("data/ecofootprint/")
+ecofp <- map(fls, function(x) {
+    jsonlite::read_json(
+        path = paste0("data/ecofootprint/", x),
+        simplifyVector = TRUE)
+}) %>% 
+    bind_rows()
+
+inq <- readxl::read_xlsx(
+    path = "~/Documents/Projects/DATA/WIID_World_Income_Inequality_DB/WIID_06MAY2020.xlsx")
+
+dat1 <- read_csv(
+    "~/Documents/Projects/DATA/WorldBank/SDG_csv/SDGData.csv") %>% 
+    janitor::clean_names()
+
+dat <- ecofp %>%
+    filter(record == "EFConsPerCap") %>%
+    select(countryName, isoa2, year, value) %>% 
+    rename(EFconsPerCap = value, country = countryName) %>% 
+    filter(year >= 1995)
+
+dat <- dat %>% 
+    left_join(
+        .,
+        inq %>% 
+            filter(!is.na(gini_reported), year >= 1995) %>% 
+            select(country, isoa2 = c2, country_code = c3, year, gini_reported) %>% 
+            group_by(country, year, isoa2, country_code) %>% 
+            summarize(gini_mean = mean(gini_reported, na.rm = TRUE)) 
+        
+    )
+
+dat <- dat %>% 
+    left_join(
+        ., 
+        dat1 %>% 
+            select(-last_col()) %>% 
+            filter(indicator_name == "GNI per capita (US$)") %>% 
+            pivot_longer(cols = x1990:x2019, names_to = "year", values_to = "value") %>%
+            mutate(year = str_remove(year, "x")) %>% 
+            mutate(year = as.numeric(year)) %>% 
+            mutate(ismissing = is.na(value)) %>% 
+            filter(year >= 1995) %>% 
+            rename(country = country_name, gni = value)
+    )
+
+## compute missing values
+missingness <- dat %>% 
+    mutate(ismissing = is.na(gini_mean)) %>% 
+    group_by(country) %>% 
+    summarize(missingness = sum(ismissing)/n()) %>% 
+    arrange(missingness)
+
+
[…]

### 2022-09-27 figures.R

@@ -110,16 +110,91 @@ p0 <- ggplot() +
         xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf
     ) +
     labs(tag = "A")
+  
+p4 <- df_dat %>% 
+    ggplot(aes(EFconsPerCap, gini, group = country)) +
+    geom_point(size = 0.5) +
+    geom_path(size = 0.2, aes(color = year)) +
+    annotate(geom = "rect", xmin = 0, xmax = 1.68, ymin = 20, ymax = 30,
+             fill = "skyblue", alpha = 0.2) +
+    annotate(geom ="rect",xmin = 0, xmax = 1.68, ymin = 30, ymax = 62,
+             fill = "orange", alpha = 0.2) +
+    annotate(geom ="rect", xmin = 1.68, xmax = 18, ymin = 20, ymax = 30,
+             fill = "orange", alpha = 0.2) +
+    annotate(geom="rect", xmin = 1.68, xmax = 18, ymin = 30, ymax = 62,
+             fill = "red", alpha = 0.2) +
+    geom_hline(yintercept = 30, linetype = 2, color = "gray50") +
+    geom_vline(xintercept = 1.68, linetype = 2, color = "gray50") +
+    scico::scale_color_scico("Year", palette = "bilbao") +
+    labs(y = "Inequality (Gini)", x = "Ecological footprint", tag = "E") + theme_light() +
+    theme(legend.position = "top", panel.grid = element_blank(),
+          legend.key.width = unit(10, "mm"), legend.key.height = unit(3,"mm"))
 
-p0 + p3 + p2 + p1 + plot_layout(ncol = 2)
+p5<- df_dat %>% 
+    ggplot(aes(gni2, gini, group = country)) +
+    geom_point(size = 0.5, show.legend = FALSE) +
+    geom_path(size = 0.2, aes(color = year)) +
+    annotate(geom = "rect", xmin = 10, xmax = 12746, ymin = 20, ymax = 30,
+             fill = "orange", alpha = 0.2) +
+    annotate(geom ="rect",xmin = 10, xmax = 12746, ymin = 30, ymax = 62,
+             fill = "red", alpha = 0.2) +
+    annotate(geom ="rect", xmin = 12746, xmax = 105000, ymin = 20, ymax = 30,
+             fill = "skyblue", alpha = 0.2) +
+    annotate(geom="rect", xmin = 12746, xmax = 105000, ymin = 30, ymax = 62,
+             fill = "orange", alpha = 0.2) +
[…]

### 2022-09-27 lit_review.R

@@ -6,6 +6,7 @@ library(lda)
 library(tictoc)
 library(network)
 library(patchwork)
+library(fs)
 
 # not working, need manual download
 # dat <- read_csv(file = "https://stockholmuniversity.box.com/s/uoii9orvhkoijk1hzg4ahqwdydwhuxlb")
@@ -13,24 +14,41 @@ library(patchwork)
 # dat <- read_csv(file = "data/April27_1-2000.csv") %>% 
 #     janitor::clean_names()
 
-fls <- list.files(path = "data/lit_review/")
+fls <- dir_ls("data/lit_review_20220624/")
 
-dat_xls <- paste0("data/lit_review/", str_subset(fls, ".xlsx")) %>% 
-    map(readxl::read_excel) %>% 
+## working with a 2022 dataset. thre is a strange behaviour, reading the data 
+## results in more observations that hits reported in Scopus. Using read.csv I get 87k
+## observations, but using read_csv I get 74k (the expected amount). From now on, only use 
+## read_csv; the other is not reliable.
+
+# dat_xls <- paste0("data/lit_review/", str_subset(fls, ".xlsx")) %>% 
+#     map(readxl::read_excel) %>% 
+#     map(janitor::clean_names) %>% 
+#     map(function(x) x %>% select(abstract, title, year, doi) %>% as_tibble()) %>% 
+#     map(function(x) x %>% mutate(across(.cols = everything(), .fns = as.character))) %>% 
+#     bind_rows()
+
+dat  <- str_subset(fls, ".csv") %>% # head() %>% 
+    map(read_csv) %>% 
     map(janitor::clean_names) %>% 
     map(function(x) x %>% select(abstract, title, year, doi) %>% as_tibble()) %>% 
     map(function(x) x %>% mutate(across(.cols = everything(), .fns = as.character))) %>% 
     bind_rows()
 
-dat_csv  <- paste0("data/lit_review/", str_subset(fls, ".csv")) %>% # head() %>% 
-    map(read.csv) %>% 
-    map(janitor::clean_names) %>% 
-    map(function(x) x %>% select(abstract, title, year, doi) %>% as_tibble()) %>% 
-    map(function(x) x %>% mutate(across(.cols = everything(), .fns = as.character))) %>% 
-    bind_rows()
 
[…]

### 2022-09-27 sdg_api.R

@@ -1,0 +1,3 @@
+
+
+"https://unstats.un.org/SDGAPI/swagger/../swagger/v1/swagger.json"

### 2022-12-02 inequality-SDGs.Rmd
progress on inequality SDG analysis at national scale + matching problem

@@
 This notebooks explores synergies and trade-offs between inequality and other SDGs with the World Bank dataset on sustainable development goals at the national scale.
+
+## World Bank SDGs
 
@@
 
+q <- dat |>
+    filter(indicator_name %in%  vars) |>
+    select(-x35) |>
+    pivot_longer(cols = x1990:last_col(), values_to = "value", names_to = "year") |>
+    mutate(year = str_remove(year, "x"), year = as.numeric(year)) |>
+    mutate(nas = is.na(value)) |>
+    filter(year >= 2000, year < 2019) |>  # restricting time improves
+    group_by(country_code, indicator_name) |>
+    summarise(missing = sum(nas) / diff(range(year))) |>
+    #filter(country_code %in% complete_countries) |> # this line only if next cell already run
+    # ungroup() |> group_by(country_code) |>
+    # summarize(mean_missing = mean(missing)) |>
+    #filter(mean_missing < 0.3) |>
+    ggplot(aes(country_code, indicator_name)) +
+    geom_tile(aes(fill = missing)) +
+    scale_fill_viridis_c() +
+    theme_light(base_size = 6) +
+    theme(axis.text.x = element_text(angle = 90))
+
+q
+
 ```
+
+```{r}
+complete_countries <-  dat |>
+    filter(indicator_name %in%  vars) |>
+    select(-x35) |>
+    pivot_longer(cols = x1990:last_col(), values_to = "value", names_to = "year") |>
+    mutate(year = str_remove(year, "x"), year = as.numeric(year)) |>
+    mutate(nas = is.na(value)) |>
+    filter(year >= 2000, year < 2019) |>  # restricting time improves
+    group_by(country_code, indicator_name) |>
+    summarise(missing = sum(nas) / diff(range(year))) |> 
+    # filter out any country with missing values higher than 70%
+    filter(!any(missing > 0.3)) |> 
+    pull(country_code) |> unique()
+
+complete_countries
+```
+
+Using `tsibble`:
+
+```{r}
+
+dat |>
+    filter(indicator_name %in% vars, country_code %in% complete_countries) |> 
+    select(-x35) |> 
+    pivot_longer(cols = x1990:last_col(), values_to = "value", names_to = "year") |> 
+    mutate(year = str_remove(year, "x")) |>  ## Formating for year-month-day
+    #mutate(year = lubridate::parse_date_time(year, orders = "Y", select_formats = "%Y"))  |> 
+    #mutate(year = strptime(year, format = "%Y")) |>
+    mutate(year = lubridate::make_date(year = year)) |> 
+    as_tsibble(key = c(country_code, indicator_name), index=year, regular = TRUE) 
+    # count_gaps()
+
+```
+
 ### Correlations
@@
 dat |>
-    filter(indicator_name %in% vars, country_code %in% countries_list) |> 
+    filter(indicator_name %in% vars, country_code %in% complete_countries) |> 
     select(-x35) |> 
@@
     mutate(year = str_remove(year, "x"), year = as.numeric(year)) |> 
+    filter(year >= 2000, year < 2019) |>  # restricting time improves
     select(-indicator_code) |> 
@@
 df_dat <- dat |>
-    filter(indicator_name %in% vars, country_code %in% countries_list) |> 
+    filter(indicator_name %in% vars, country_code %in% complete_countries) |> 
     select(-x35) |> 
@@
 
-df_dat
+df_dat |> ungroup() |> skimr::skim() # no missing values
 ```
@@
     filter(n>19) |>  # 19 is the max number of years
-    arrange(desc(n))
+    arrange(desc(n)) #|> pull(geo_area_code) |> unique() # 191 countries (all of them)
 
@@
     mutate(nas = is.na(value)) |> 
-    group_by(indicator_name, year) |> 
+    group_by(indicator, indicator_name, year) |> 
     summarise(missing = sum(nas) / n_countries2) |> 
     mutate(indicator_name = str_trunc(indicator_name, width = 30, side = "right")) |>
-    ggplot(aes(year, indicator_name)) +
+    ggplot(aes(year, indicator)) +
     geom_tile(aes(fill = missing)) +
@@
 
+## Very long names:
 sdg|> 
@@
     filter(goal == 15) |> 
-    pull(indicator_name) |> unique()
+    select(indicator, indicator_name) |> 
+    unique()
 
@@
 
+Usable variables from UN Dataset:
+
+```{r}
+p2 <- sdg |> 
+    rename(series_name = series_description, year = time_period, iso3 =`ISO-alpha3 Code` ) |> 
+    filter(iso3 %in% complete_countries) |> # use only countries
+    filter(goal %in% c(10, 14, 15)) |> 
+    group_by(series_code, year, iso3) |>
+    # this gets rid of multiple values per year
+    summarize(mean_val = mean(value, na.rm = TRUE)) |> 
+    # add_count() |> 
+    pivot_wider(names_from = year, values_from = mean_val) |> 
+    pivot_longer(cols = starts_with("2"), names_to = "year", values_to = "value",
+                 names_transform = list(year = as.numeric, value = as.numeric)) |> 
+    mutate(nas = is.na(value)) |> 
+    ungroup() |> group_by(series_code, iso3) |> 
+   # group_by(indicator, indicator_name, iso3) |> 
+    summarise(missing = sum(nas) / diff(range(year)) ) |>  ## dividing by number of years (per country)
+    #pull(missing) |> range() # should be between 0-1
+    # summarize(mean_missing = mean(missing)) |> 
+    filter(missing < 0.3) |> 
+    ungroup() |> 
+    pivot_wider(names_from = series_code, values_from = missing) |> 
+    pivot_longer(cols = AG_LND_FRSTCERT:last_col(), names_to = "series_code",
+                 values_to = "missing") |> 
+    group_by(series_code) |> 
+    #filter(!any(is.na(missing))) |> 
+    # ggplot(aes(iso3, series_code)) +
+    # geom_tile(aes(fill = missing)) +
+    # theme_light()
+    summarize(missing_countries = sum(is.na(missing))) |> 
+    filter(missing_countries < 45) |> 
+    ggplot(aes(missing_countries, series_code))+
+    geom_point()
+
+p2 + theme_light(base_size = 12)
+
+vars2 <- p2$data$series_code
+```
+
+```{r}
+q2 <- sdg |> 
+    rename(series_name = series_description, year = time_period, iso3 =`ISO-alpha3 Code` ) |> 
+    filter(iso3 %in% complete_countries) |> # use only countries
+    filter(series_code %in% vars2) |> 
+    ## this line is to get rid of multiple values per year
+    group_by(series_code, year, iso3) |>
+    summarize(mean_val = mean(value, na.rm = TRUE)) |> 
+    ungroup() |> #group_by(indicator_name, year, iso3) |> 
+    # add_count() |> 
+    pivot_wider(names_from = year, values_from = mean_val) |> 
+    pivot_longer(cols = starts_with("2"), names_to = "year", values_to = "value",
+                 names_transform = list(year = as.numeric, value = as.numeric)) |> 
+    mutate(nas = is.na(value)) |> 
+    group_by(series_code, iso3) |> 
+    summarise(missing = sum(nas) / diff(range(year))) |>  ## dividing by number of years (per country)
+    # summarize(mean_missing = mean(missing)) |> 
+    ungroup() |> 
+    filter(missing < 0.3) |> 
+    ## add missing values again for combination of indicator country:
+    ungroup() |> group_by(series_code) |>
+    pivot_wider(names_from = iso3, values_from = missing) |> 
+    pivot_longer(cols = AFG:last_col(), names_to = "iso3", values_to = "missing") |>
+    ## completely empty time series
+    #summarize(absent_countries = sum(is.na(missing))) |> 
+    ungroup() |> group_by(iso3 ) |> 
+    #filter(sum(is.na(missing)) < 45) |> # remove absent_countries > 45
+    filter(!any(is.na(missing))) |> 
+    ggplot(aes(iso3, series_code)) +
+    #geom_point()
+    geom_tile(aes(fill = missing)) +
+    scale_fill_viridis_c() +
+    theme_light() +
+    theme(axis.text.x = element_text(angle = 90))
+
+q2
+
+complete_countries_sdg <- q2$data$iso3 |> unique()
+```
+
+Reduce the dataset to variables and countries with almost complete data. Then impute missing values.
+
+```{r}
+df_un_reduced <- sdg |> 
+    rename(series_name = series_description, year = time_period, iso3 =`ISO-alpha3 Code` ) |> 
+    filter(iso3 %in% complete_countries_sdg) |> # use only countries
+    filter(series_code %in% vars2) |> 
+    group_by(series_code, year, iso3) |>
+    summarize(mean_val = mean(value, na.rm = TRUE)) |> 
+    ungroup() |> 
+    pivot_wider(names_from = year, values_from = mean_val) |> 
+    pivot_longer(cols = starts_with("2"), names_to = "year", values_to = "value",
+                 names_transform = list(year = as.numeric, value = as.numeric)) |> 
+    group_by(iso3, series_code) |> 
+    mutate(value = na_interpolation(value, option = "spline"))
+
+df_un_reduced |> ungroup() |> skimr::skim() # zero missing values
+```
+
+Now there is two working datasets cleaned and complete: `df_dat` is the data from World Bank, and `df_un_reduced` is the data from UN SDGs.
+
 ## Notes:
 
-Select the variables for which you have time series. Repeat the trajectories analysis. 
-Do static analysis with all SDGs
+1. Select the variables for which you have time series. Repeat the trajectories analysis. 
+2. Do static analysis with all SDGs
+
+Clean up a bit:
 
 ```{r}
-ctr <- sdg |> pull(geo_area_name) |> unique() |> length()
-sdg <- sdg |> select(-c(age:type_of_speed))
+rm(df_sdg, df_un, df_wb, maike_vars, other_vars)
+# key files
+# sdg raw data from UN SDG
+# dat raw data from WorldBank
+```
 
-sdg <- sdg |> 
-    mutate(value = as.numeric(value)) |> 
-    rename(year = time_period)
-yrs <- sdg$year |> range() |> diff()
-sdg <- sdg |> unique()
+## Ordination
 
-dat2 <- sdg |> 
-    filter(goal == 10 | goal == 14 | goal == 15 ) |> 
-    group_by(indicator, geo_area_code, year, geo_area_name, goal, target, series_code, series_description) |> 
-    summarize(value = mean(value))
+```{r}
+pca1 <- df_dat |> 
+    select(-country_name) |> 
+    pivot_longer(cols = `Access to electricity (% of population)`:last_col(),
+                 values_to = "value", names_to = "variable") |> 
+    unite(col = "var", year, variable, sep = "_") |> 
+    pivot_wider(names_from = var, values_from = value) |> 
+    ungroup() |> 
+    select(-country_code) |> 
+    prcomp(center = TRUE, scale. = TRUE)
+```
 
-## If I add series description it duplicate entries, possibly due to misspellings
-#test
-dat2 |>
-    unique() |>
-    add_count() |>
-    filter(n>1)
+PCA
+```{r}
+plot(pca1)
+```
 
-dat2 |> 
-    add_count() |> 
+```{r}
+biplot(pca1)
+```
+
+```{r}
+summary(pca1)
+```
+
+```{r}
+pca2 <- df_un_reduced |> 
+    unite(col = "var", series_code, year, sep = "_") |> 
+    pivot_wider(values_from = value, names_from = var) |> 
     ungroup() |> 
-    group_by(geo_area_name,  series_code) |> 
-    summarize(yrs_cov = sum(n)) |> 
-    filter(yrs_cov > 20)
+    select(-iso3) |> 
+    as.matrix(dimnames = list(df_un_reduced |> pull(iso3) |> unique())) |> 
+    prcomp(center = TRUE, scale. = TRUE)
+```
 
-sdg |> filter(series_code == "DC_ODA_BDVDL")
+```{r}
+biplot(pca2)
 ```
+### Multiple factor analysis
+
+```{r}
+mfa1 <- df_un_reduced |> 
+    unite(col = "var", series_code, year, sep = "_") |> 
+    pivot_wider(values_from = value, names_from = var) |> 
+    ungroup() |> 
+    select(-iso3) |> 
+    as.matrix(dimnames = list(df_un_reduced |> pull(iso3) |> unique())) |> 
+    FactoMineR::MFA(group = rep(19, 6), type = rep("s", 6)) # 18 years times 6 variables
+```
+
+```{r}
+mfa1
+```
+
+```{r}
+mfa1$eig[1:10,2] |> barplot(names = 1:10)
+
+```
+```{r}
+mfa1
+```
+
+## To dos:
+
+- one PCA per year, see how PCA1 change over time.
+
+## Leftovers
+
+Old code, most of the steps outlined are already implemented
+
+```{r}
+### World Bank dataset:
+# vars is the list of variables with <30% missing values over time and space
+# vars2 is the same list for UN dataset
+# yrs <- sdg$time_period|> range(na.rm = TRUE) 
+# 
+# df_wb <- dat |>
+#     filter(indicator_name %in% vars, country_code %in% complete_countries) |> 
+#     select(-x35) |> 
+#     pivot_longer(cols = x1990:last_col(), values_to = "value", names_to = "year") |> 
+#     mutate(year = str_remove(year, "x"), year = as.numeric(year)) |> 
+#     filter(year >= yrs[1], year <= yrs[2]) |> # use only years available in both datsets
+#     rename(iso3 = country_code) |> select(-indicator_code) |> 
+#     left_join(
+#         select(un_countries, iso3 = `ISO-alpha3 Code`, country_name = `Country or Area`)
+#     )
+# 
+# df_un <- sdg |> 
+#     rename(indicator_name = series_description, year = time_period, iso3 =`ISO-alpha3 Code` ) |> 
+#     filter(iso3 %in% countries_list, !is.na(value), indicator_name %in% vars2) |>
+#     filter(goal %in% c(10, 14, 15)) |> 
+#     group_by(indicator_name, year, iso3) |>
+#     summarize(mean_val = mean(value, na.rm = TRUE)) |> 
+#     ungroup() |> #group_by(indicator_name, year, iso3) |> 
+#     # add_count() |> 
+#     pivot_wider(names_from = year, values_from = mean_val) |> 
+#     pivot_longer(cols = starts_with("2"), names_to = "year", values_to = "value",
+#                  names_transform = list(year = as.numeric, value = as.numeric))|> 
+#     left_join(
+#         select(un_countries, iso3 = `ISO-alpha3 Code`, country_name = `Country or Area`)
+#     )
+#     
+# df_sdg <- full_join(df_un, df_wb) |> 
+#     rename(indicator = indicator_name)
+# 
+# df_sdg |> skimr::skim()
+```
+
+```{r}
+# df_sdg |> 
+#     ggplot(aes(year, value)) +
+#     geom_line(aes(group = iso3), show.legend = FALSE) +
+#     facet_wrap(~indicator, scales = "free_y")
+```
+remove countries with too few observations:
+
+```{r}
+# tot_vars <- df_sdg |> pull(indicator) |> unique() |> length()
+# tot_yrs <- df_sdg |> pull(year) |> unique() |> length()
+# tot_cnt <- df_sdg |> pull(iso3) |> unique() |> length()
+# 
+# df_sdg |> 
+#     pivot_wider(names_from = indicator, values_from = value) |> 
+#     pivot_longer(cols = 4:last_col(),names_to = "indicator", values_to = "value") |> 
+#     mutate(nas = is.na(value)) |>
+#     group_by(iso3, indicator) |> 
+#     dplyr::summarize(missing = sum(nas)/ (tot_yrs)) |> 
+#     summarize(suitable = missing > 0.3) |> 
+#     ungroup() |> group_by(iso3) |> 
+#     summarize(varsok = sum(suitable)) |> 
+#     filter(varsok == tot_vars)
+# 
+# key_cntrs
+```
+
+Impute missing values:
+
+```{r}
+# df_sdg <- df_sdg |> 
+#     filter(iso3 %in% key_cntrs) |> 
+#     pivot_wider(names_from = indicator, values_from = value) |> 
+#     pivot_longer(cols = 4:last_col(),names_to = "indicator", values_to = "value") |> 
+#     group_by(indicator, iso3) |> 
+#     mutate(value = na_interpolation(value, option = "stine"))
+```
+
+```{r}
+# df_sdg |> 
+#     ggplot(aes(year, value)) +
+#     geom_line(aes(group = iso3), alpha = 0.5, show.legend = FALSE) +
+#     facet_wrap(~indicator, scales = "free_y")
+```
+Check for correlations:
+
+```{r}
+# df_sdg |> 
+#     pivot_wider(names_from = indicator, values_from = value) |> 
+#     GGally::ggpairs(
+#         columns = 4:21 ,
+#     lower = list(continuous = GGally::wrap("points", alpha = 0.5, size = 0.5)),
+#     diag = list(continuous = GGally::wrap("densityDiag", alpha = 0.2))
+#     # upper = list(continuous = GGally::wrap("cor", size = 2))
+#     ) + theme_light(base_size = 6)
+```
+
+```{r}
+# df_sdg <- df_sdg |> 
+#     ungroup() |> 
+#     pivot_wider(names_from = indicator, values_from = value) |> 
+#     mutate(
+#         across(.cols = c("Cereal yield (kg per hectare)" ,"Forest area (sq. km)" ,
+#             "Forest area certified under an independently verified certification scheme (thousands of hectares)" , "Total assistance for development, by donor countries (millions of current United States dollars)"  , "Total assistance for development, by recipient countries (millions of current United States dollars)"
+#         ), .fns = log1p)
+#     ) 
+```
+
+Paper title: the blind spots of sustainable development
+
+- underscore the fact that a lot of variables are not available
+- the kind of variables available are not appropriate in many cases
+- sub-national conversation: countries interpret and collect data differently [carla and emmy work]
+- what can be done to address the gaps? use of other datasets
+- open questions for other SDGs dimensions...

### 2022-12-02 99-merging_problem.R

@@ -1,0 +1,104 @@
+library(tidyverse)
+library(tidytext)
+library(tm)
+library (topicmodels)
+library(lda)
+library(tictoc)
+library(fs)
+library(fuzzyjoin)
+
+prob <- read_csv("data/MaxGamma2_DOI_Included_Nov10_2022.csv", locale = locale("sv") )
+
+#current fit
+load("data/220625_topic_models_gibbs.RData")
+
+
+
+new_fit <- topicNumber.TM[[3]] |> 
+    slot("gamma") |> # take the gamma matrix
+    as_tibble() |> # convert to data frame
+    # add a column with paper titles
+    add_column(docs = topicNumber.TM[[3]] |> slot("documents")) |> 
+    # make all gamma values into one column for filtering later
+    pivot_longer(cols = starts_with("V"), names_to = "topic", values_to = "gamma" ) |> 
+    # select only topic 1
+    # filter(topic == "V1") |> 
+    group_by(docs) |> 
+    filter(gamma == max(gamma))
+# arrange by descending gamma value (the prob of a document belonging to a topic)
+# arrange(desc(gamma)) |> 
+# # print only the top 10
+# top_n(10)
+# 
+
+
+#old fit: 
+load("data/211020_topic_models_gibbs.RData")
+old_fit <- topicNumber.TM[[3]] |> 
+    slot("gamma") |> # take the gamma matrix
+    as_tibble() |> # convert to data frame
+    # add a column with paper titles
+    add_column(docs = topicNumber.TM[[3]] |> slot("documents")) |> 
+    # make all gamma values into one column for filtering later
+    pivot_longer(cols = starts_with("V"), names_to = "topic", values_to = "gamma" ) |> 
+    # select only topic 1
+    # filter(topic == "V1") |> 
+    group_by(docs) |> 
+    filter(gamma == max(gamma))
+
+## Comparison with All documents in prob file
+new_fit$docs[!new_fit$docs %in% prob$Document_OLD_ALL] |> length() # 5261
+prob$Document_OLD_ALL[!prob$Document_OLD_ALL %in% new_fit$docs] |> length() # 6108
+
+## Comparison with coded documents in prob file (different file)
+new_fit$docs[!new_fit$docs %in% prob$Document_OLD_INCL] |> length() # 56725
+prob$Document_OLD_INCL[!prob$Document_OLD_INCL %in% new_fit$docs] |> length() # 1912
+
[…]

### 2022-12-02 Max Gamma Calculations_Tom.R

@@ -1,0 +1,28 @@
+
+ls()
+rm(list = ls())
+
+#setwd("C:/Users/twbc201/OneDrive - University of Exeter")
+setwd("/Users/emilielindqvist/Dropbox/BYS2/R Inequlity and the Biospere/from Tom/")
+
+library(tidyverse)     
+
+x <- read.csv("topic_papers_DOI_fromJuan.csv")
+pause
+names(x)
+
+summary(x)
+str(x)
+names(x)
+x$topic <- as.factor(x$topic)
+str(x)
+
+
+# this works 
+test <- (
+x %>%
+  group_by(document) %>%
+  filter(gamma == max(gamma))
+)
+head(test)
+write.csv(test, "maxgamma_DOI_fromJuan.csv")

### 2022-12-02 lit_review.R

@@ -120,7 +120,7 @@ toc() #43876.676 sec elapsed or 12.2hrs
 save(tset.TM, topicNumber.TM, dtm, file = "data/220625_topic_models_gibbs.RData")
 
 
-load("data/211020_topic_models_gibbs.RData")
+load("data/220625_topic_models_gibbs.RData")
 
 
 #### visualizations ####
@@ -315,3 +315,50 @@ networkD3::simpleNetwork(
 
 ## free memory
 rm(mat, authors)
+
+
+new_fit <- topicNumber.TM[[3]] |> 
+    slot("gamma") |> # take the gamma matrix
+    as_tibble() |> # convert to data frame
+    # add a column with paper titles
+    add_column(docs = topicNumber.TM[[3]] |> slot("documents")) |> 
+    # make all gamma values into one column for filtering later
+    pivot_longer(cols = starts_with("V"), names_to = "topic", values_to = "gamma" ) |> 
+    # select only topic 1
+    # filter(topic == "V1") |> 
+    group_by(docs) |> 
+    filter(gamma == max(gamma))
+    # arrange by descending gamma value (the prob of a document belonging to a topic)
+    # arrange(desc(gamma)) |> 
+    # # print only the top 10
+    # top_n(10)
+
+
+#old fit: 
+load("data/211020_topic_models_gibbs.RData")
+
+old_fit <- topicNumber.TM[[3]] |> 
+    slot("gamma") |> # take the gamma matrix
+    as_tibble() |> # convert to data frame
+    # add a column with paper titles
+    add_column(docs = topicNumber.TM[[3]] |> slot("documents")) |> 
+    # make all gamma values into one column for filtering later
+    pivot_longer(cols = starts_with("V"), names_to = "topic", values_to = "gamma" ) |> 
+    # select only topic 1
+    # filter(topic == "V1") |> 
+    group_by(docs) |> 
+    filter(gamma == max(gamma))
+
+
+new_fit |> ungroup() |> 
+    left_join(old_fit |> ungroup() |> 
+                  rename(topic_old = topic, gamma_old = gamma)) |> 
+    filter(is.na(old))
+
+new_fit$docs[!new_fit$docs %in% old_fit$docs] |> length() #112 docs
+old_fit$docs[!old_fit$docs %in% new_fit$docs] |> length() #11
+
+    #write_csv(file = "data/old_and_new_max_gammas.csv")
+
+
+

### 2022-12-02 99-merging_problem.R
fuzzy matching problem

@@ -59,10 +59,10 @@ prob$Document_OLD_INCL[!prob$Document_OLD_INCL %in% new_fit$docs] |> length() # 
 
 ## Extract the 1912 problematic titles
 probs <- prob[!prob$Document_OLD_INCL %in% new_fit$docs,] |> 
-    select(Document_OLD_INCL, Topic_OLD_INCL) |> 
+    #select(Document_OLD_INCL, Topic_OLD_INCL) |> 
     filter(!is.na(Topic_OLD_INCL)) |> 
-    rename(docs = Document_OLD_INCL) |> 
-    select(-Topic_OLD_INCL)
+    rename(docs = Document_OLD_INCL) #|> 
+    #select(-Topic_OLD_INCL)
 
 ## The problems are reduced to 1852, I removed all documents that do not have old topic
 ## since they are not matched by name, they do not have new topic either. There are a lot of 
@@ -101,4 +101,17 @@ fuzzy_match |>
     arrange(desc(n))
 
 
-save(fuzzy_match, file = "data/fuzzy_match.RData")
+# save(fuzzy_match, file = "data/fuzzy_match.RData")
+
+
+### Reconstruct the file with the non-problematic cases
+
+result <- prob[prob$Document_OLD_INCL %in% new_fit$docs,] |> 
+    left_join(new_fit, by = c("Document_OLD_INCL" = "docs"))
+
+
+result <- fuzzy_match |> 
+    rename(Document_OLD_INCL = docs.x) |> 
+    bind_rows(result)
+
+write_csv(result, file = "data/matching_problem.csv")

### 2023-02-20 Trilemma_clustering.Rmd
first paper draft + older files

@@
     arrange(desc(EFconsPerCap))
+
+#save(df_dat, file = "data/cleaned_trilemma.RData")
 ```
@@
     mutate(country = as_factor(country)) |> 
-    mutate(country = fct_reorder2(country, magnitude_a, Gini)) |> 
+    mutate(country = fct_reorder2(country, magnitude, Gini)) |> 
     pivot_longer(
-        cols = sd_magnitude:gamma_a, names_to = "vars", 
+        cols = sd_magnitude:gamma, names_to = "vars", 
         values_to = "vals") |> 

### 2023-02-20 inequality-SDGs.Rmd
first paper draft + older files

@@
 p + theme_light(base_size = 8)
+
+# ggsave(
+#     plot = p, filename = "WB_environment_SDGs_othervars.png", path = "figures/", device = "png",
+#     width = 6, height = 4, dpi = 400, bg = "white"
+# )
 ```
@@
 
-```{r}
+```{r warnings=FALSE, message=FALSE}
 dat |>
@@
 
+```{r}
+#save(df_dat, df_un_reduced, file = "data/cleaned_sdg.RData")
+```
+
 ## Notes:
@@
 ```{r}
-summary(pca1)
+# summary(pca1)
 ```
@@
 ```{r}
-mfa1
+mfa1$group
+```
+
+```{r}
+mfa1 |> summary()
+```
+
+```{r}
+ggbiplot::ggbiplot(
+    pca1,
+    labels = df_dat$country_code |> unique(),
+    circle = TRUE,
+    varname.abbrev = TRUE)
 ```
@@
 - one PCA per year, see how PCA1 change over time.
+
+```{r}
+pca3 <- df_dat |> 
+    ungroup() |> 
+    mutate(year = as_factor(year)) |> 
+    split(~year) |>
+    map(function(x) select(x, where(is.numeric))) |> 
+    map(function(x) princomp(x, center = TRUE, scale. = TRUE))
+```
+
+```{r}
+pca3 |> map(biplot)
+```
+
+## Paper outline
+
+Title: Trade-offs between biosphere goals and inequality at the national scale
+
+Introduction
+
+- Previous studies proposed the hypothesis of the trilemma
+- Do the trade-offs hold at national scale with SDG data?
+
+Methods:
+
+- SDG datasets at national scale collected by World Bank and United Nations
+- Selection of variables
+- Multiple Factor Analysis (PCA over time)
+- Clustering
+
+Results:
+
+-
 

### 2023-02-20 paper/manuscript.Rmd
first paper draft + older files

---
title: "Trade-offs between biosphere goals and inequality at the national scale"
output:
  pdf_document:
    dev: pdf
    keep_tex: yes
    toc: no
    latex_engine: xelatex
  rtf_document:
    toc: no
  word_document:
csl: /Users/juanrocha/Documents/styles/nature.csl
bibliography: refs.bib
citecolor: blue
documentclass: article
font: Arial
fontsize: 10pt
linkcolor: blue
urlcolor: blue
header-includes:
- \usepackage{dcolumn, rotating, longtable, lineno, float, array, tabularx, inputenc}
- \setmainfont{Arial}
- \renewcommand{\abstractname}{}
- \graphicspath{{figures/}}
- \usepackage[margin=2.5cm]{geometry}
- \linenumbers
- \usepackage{setspace}
---

# Outline

Target: EcolEcon?
/ People and Nature?
/ PNAS / ERL [4000w]

Authors:

Introduction {800w}

-   Previous studies proposed the hypothesis of the trilemma

-   Do the trade-offs hold at national scale with SDG data?

Methods {800w}:

-   SDG datasets at national scale collected by World Bank and United Nations
-   Selection of variables
-   Multiple Factor Analysis (PCA over time)
-   Clustering

Results {1000w}:

-

Discussion {1000w}

Conclusion {400w}

SM

-   Figure with missing values from WB dataset

-   Figure with missing alues from UN dataset

-   Correlograms

-

# Introduction

Countries around the world have committed to achieve 17 sustainable development goals.
The ambitious agenda is materialized in 169 targets and indicators, yet not all targets are monitored and not all countries report them.
An open question in sustainability science is if these targets are achievable simultaneously?
Previous work on synergies and trade-offs between sustainable development goals (SDGs) suggest that both options are possible, yet this work is limited to expert elicitation and correlational studies.
It is unclear if there are structural constraints that limits a country capacity to achieve particular SDGs, or if there are milestones that need to be unlocked before it can achieve the desired performance in a particular dimension of sustainability.
In other words, these correlations do not account for the path dependency inherent of the process of development.

Here we address the question of synergies and trade-offs by focusing on goals related to biosphere and inequality (SDGs 10, 14 and 15).
The motivation is twofold.
First, the Convention for Biological Diversity is currently negotiating and agreeing on the next set of goals and ambitions to bend the curve of biodiversity loss.
Understanding which countries are well positioned to achieve biodiversity goals or not will be necessary to design strategies that support countries with low capabilities, that address the dimensions of justice, and design targets with plausibility in mind.
Second, within country inequality has been rising in the last decades even in high income countries, and as consequence of the Covid pandemic, in between country inequality raised for first time in a generation.
These statistics set back progress on the inequality SDG by at least a decade, and undermines the mantra of leaving no one behind.

Recent work shows that a trilemma exist where countries struggle to achieve high prosperity, high environmental standards while reducing inequality.
Using data from environmental footprint, the gross national product, and the Gini coefficient time series, 53 countries were clustered and a typology of trajectories identified.
No country achieved the three goals simultenously, and Latin American countries seem to have dynamics characterized by an inequality trap.
Countries trajectories on the trilemma space, however, suggest the hypothesis that social progress can be achieved without compromising the biosphere.
It remains an open questions if these patterns hold across different datasets and scales.
Here we address such gap by studying synergies and trade-offs between SDGs related to inequality and the biosphere.

# Methods

```{r echo = FALSE}
dat <- read_csv(file = "~/Documents/Projects/DATA/WorldBank/SDG_csv/SDGData.csv") |> 
    janitor::clean_names()

un_countries <- read_csv2(file = "~/Documents/Projects/DATA/SDGs_UNStats/UNSD — Methodology.csv")
sdg <- read_csv(file = "~/Documents/Projects/DATA/SDGs_UNStats/20180817102044132_juan.rocha@su.se_data.csv") |> janitor::clean_names()

sdg <- sdg |> 
    left_join(
        select(un_countries, country = `Country or Area`, starts_with("ISO")),
        by = c("geo_area_name" = "country")) |> 
    filter(value != "-")|> 
    mutate(value = as.numeric(value))

```

*Datasets:* To that end, we used the SDGs datasets made openly available by the World Bank (WB) and United Nations (UN).
The WB dataset offers `r dat$indicator_name |> unique() |> length()` indicators, with time series from 1990 to 2019 for `r dat$country_name |> unique() |> length()` countries or administrative areas (N = 2'013.791 observations).
The UN dataset offers time series from 2000 to 2019 for 17 SDGs, 112 targets, and 356 indicators, in 310 administrative areas (N = 1'072.731 observations).
Despite their coverage, both datasets contain high proportion of missing values and the sample is unbalanced across countries and time, meaning some countries have better temporal coverage than others.
We focused subsequent analysis only on country level indicators for SDGs 10, 14 and 15, related to inequality and the biosphere respectively.

```{r echo = FALSE}
load(file = "data/cleaned_sdg.RData")
# `df_dat` is the data from World Bank, and `df_un_reduced` is the data from UN SDGs
df_dat <- df_dat |> ungroup()
df_un_reduced <- df_un_reduced |> ungroup()
```

*Variable selection:* We computed the proportion of missing values for all time series related to our initial selection of indicators (Figs S1, S2, S3).
We discarded indicators for which time series contained more than 30% of missing values, or less than 45 countries.
Missing values were then imputed using a cubic spline, leaving us with `r df_dat |> pull(country_code) |> unique() |> length()` countries, `r df_dat |> pull(year) |> unique() |> length()` years of data across 9 indicators for the WB dataset; and `r df_un_reduced |> pull(series_code) |> unique() |> length()` indicators, `r df_un_reduced |> pull(iso3) |> unique() |> length()` countries over `r df_un_reduced |> pull(year) |> unique() |> length()` years for the UN dataset.
Table X summarizes our variable selection, their units and available ranges.

*Ordination:* We used multiple factor analysis and principal component analysis over time to ...

Notes:

-   Add Gini to the datasets!
    repeat the analysis

-   Add the SM figures, they are ready

-   Paper figures:

    -   Ordination plot for each dataset (nicer arrows) using iso codes instead of numbers.
        Show clustering groups in colors.

    -   Plot with PCA selection (2-3 components?) and then change over time of the components.
        Do clustering on trayectories as well?
        -- and then add a map with the clusters for each dataset.

\# References {.unnumbered}

::: {#refs}
:::

\\pagebreak

# Supplementary Material {#sec:SM}

### 2023-02-20 99-merging_problem.R

@@ -115,3 +115,5 @@ result <- fuzzy_match |>
     bind_rows(result)
 
 write_csv(result, file = "data/matching_problem.csv")
+
+result <- read_csv(file = "data/matching_problem.csv")

### 2023-02-20 figures.R

@@ -1,5 +1,10 @@
 library(ggplot2)
 library(patchwork)
+library(plot3D)
+library(magrittr)
+
+
+load("data/cleaned_trilemma.RData")
 
 #### Conceptual figure ####
 old_par <- par()
@@ -41,7 +46,7 @@ p1 <- ggplot() +
     annotate("text", x = 0.15, y = 0.9, label = "Trade-off", size = 2,
              color = "red", angle = 360-45) +
     lims(x = c(0,1), y = c(0,1)) +
-    labs(x = "Inequality", y = "Environmental footprint", tag = "D") +
+    labs(x = "Inequality", y = "Ecological footprint", tag = "D") +
     coord_equal() +
     theme_light(base_size = 7) +
     theme(axis.text = element_blank(), 
@@ -94,7 +99,7 @@ p3 <- ggplot() +
     annotate("text", x = 0.15, y = 0.9, label = "Synergies", size = 2,
              color = "blue", angle = 360-45) +
     lims(x = c(0,1), y = c(0,1)) +
-    labs(x = "Environmental footprint", y = "Prosperity", tag = "B") +
+    labs(x = "Ecological footprint", y = "Prosperity", tag = "B") +
     coord_equal() +
     theme_light(base_size = 7) +
     theme(axis.text = element_blank(), 
@@ -111,63 +116,66 @@ p0 <- ggplot() +
     ) +
     labs(tag = "A")
   
+p6 <- df_dat %>% 
+    ggplot(aes( gini,EFconsPerCap , group = country)) +
+    geom_point(size = 0.15,aes(color = year), show.legend = FALSE) +
+    geom_path(size = 0.1, aes(color = year)) +
+    annotate(geom = "rect", ymin = 0, ymax = 1.68, xmin = 20, xmax = 30,
+             fill = "skyblue", alpha = 0.2) +
+    annotate(geom ="rect",ymin = 0, ymax = 1.68, xmin = 30, xmax = 62,
+             fill = "orange", alpha = 0.2) +
+    annotate(geom ="rect", ymin = 1.68, ymax = 18, xmin = 20, xmax = 30,
+             fill = "orange", alpha = 0.2) +
+    annotate(geom="rect", ymin = 1.68, ymax = 18, xmin = 30, xmax = 62,
+             fill = "red", alpha = 0.2) +
+    geom_hline(yintercept = 1.68, linetype = 2, color = "gray50", size = 0.1) +
[…]

### 2023-02-28 inequality-SDGs.Rmd
updating pca visualizations

@@
     geom_tile(aes(fill = missing)) +
-    scale_fill_viridis_c() +
+    scale_fill_viridis_c(
+        guide = guide_colorbar(barwidth = unit(2,"mm"), barheight = unit(20, "mm"))) + 
+    labs(x = "Year", y = "Indicator name") +
     theme_light(base_size = 6)
-
+# 
 # ggsave(
@@
     geom_tile(aes(fill = missing)) +
-    scale_fill_viridis_c() +
+    scale_fill_viridis_c(
+        guide = guide_colorbar(barwidth = unit(2,"mm"), barheight = unit(20, "mm"))) + 
+    labs(x = "Year", y = "Indicator name")
     theme_light(base_size = 6)
@@
 
-Wait until Maike and Tong revise the selection of variables.
-Take the analysis from there.
+**Wait until Maike and Tong revise the selection of variables.** Take the analysis from there.
 Alternatively, select a couple of topics as example, just to explore how to deal with missing values and what is doable.
@@
     geom_point() +
-    #scale_fill_viridis_c() +
+    #scale_fill_viridis_c(
+    #    guide = guide_colorbar(barwidth = unit(2,"mm"), barheight = unit(20, "mm"))) + 
+    labs(x = "Mean proportion of missing values across countries", y = "Indicator name", tag = "A")+
     theme_light(base_size = 6)
 
-p + theme_light(base_size = 8)
+#p + theme_light(base_size = 8)
 
@@
     geom_tile(aes(fill = missing)) +
-    scale_fill_viridis_c() +
+    scale_fill_viridis_c(
+        "Proportion of missing values",
+        guide = guide_colorbar(barwidth = unit(2,"cm"), barheight = unit(2,"mm"))) + 
+    labs(x = "Country code", y = "Indicator name", tag = "B") +
     theme_light(base_size = 6) +
-    theme(axis.text.x = element_text(angle = 90))
+    theme(axis.text.x = element_text(angle = 90, size = 3), 
+          legend.position = "bottom")
 
-q
+p/q
+# ggsave(
+#     plot = p/q, filename = "WB_selected_vars.png", path = "figures/", device = "png",
+#     width = 7, height = 7, dpi = 400, bg = "white"
+# )
 
@@
 
+**J230226:** If I add Gini for inequality from the inequality database, our available countries will drop down to 50+, which is what we did in the trilemma paper.
+It is risky talking about inequality wihtout Gini because the variables selected are not necessarily interpretable as inequality, but let's see what the group supports.
+
 ### Correlations
@@
 ```{r warnings=FALSE, message=FALSE}
-dat |>
+crg <- dat |>
     filter(indicator_name %in% vars, country_code %in% complete_countries) |> 
@@
     # upper = list(continuous = GGally::wrap("cor", size = 2))
-    ) + theme_light(base_size = 6)
+    ) + theme_light(base_size = 4)
+
+crg
+
+# ggsave(
+#     plot = crg, filename = "WB_correlogram.png", path = "figures/", device = "png",
+#     width = 7, height = 7, dpi = 400, bg = "white"
+# )
 ```
@@
 
-sdg |> 
+a <- sdg |> 
     rename(indicator_name = series_description, year = time_period, iso3 =`ISO-alpha3 Code` ) |> 
@@
     geom_tile(aes(fill = missing)) +
-    scale_fill_viridis_c() +
-    labs(title = "SDG10: Inequality indicators") +
-    theme_light(base_size = 8)
-
+    scale_fill_viridis_c(
+        "Missing",
+        guide = guide_colorbar(barheight = unit(2,"cm"), barwidth = unit(2,"mm"))) + 
+    labs(title = "SDG10: Inequality indicators", 
+         x = "Year", y = "Indicator name", tag = "A") +
+    theme_light(base_size = 6)
+a
 # sdg |> 
@@
 ```{r}
-sdg |> 
+b <- sdg |> 
     rename(indicator_name = series_description, year = time_period, iso3 =`ISO-alpha3 Code` ) |> 
@@
     geom_tile(aes(fill = missing)) +
-    scale_fill_viridis_c() +
-    labs(title = "SDG14: Life under water") +
-    theme_light(base_size = 8)
+     scale_fill_viridis_c(
+        "Missing",
+        guide = guide_colorbar(barheight = unit(2,"cm"), barwidth = unit(2,"mm"))) +
+    labs(title = "SDG14: Life under water", x = "Year", y = "Indicator name", tag = "B") +
+    theme_light(base_size = 6)
 ```
@@
 ```{r}
-sdg |> 
+c <- sdg |> 
     rename(indicator_name = series_description, year = time_period, iso3 =`ISO-alpha3 Code` ) |> 
@@
     summarise(missing = sum(nas) / n_countries2) |> 
-    mutate(indicator_name = str_trunc(indicator_name, width = 30, side = "right")) |>
-    ggplot(aes(year, indicator)) +
+    mutate(indicator_name = str_trunc(indicator_name, width = 100, side = "right")) |>
+    ggplot(aes(year, indicator_name)) +
     geom_tile(aes(fill = missing)) +
-    scale_fill_viridis_c() +
-    labs(title = "SDG15: Life on land") +
-    theme_light(base_size = 8)
+    scale_fill_viridis_c("Missing",
+        guide = guide_colorbar(barheight = unit(2,"cm"), barwidth = unit(2,"mm"))) +
+    labs(title = "SDG15: Life on land", x = "Year", y = "Indicator name", tag = "C") +
+    theme_light(base_size = 6)
+c
 
@@
     unique()
+
+# ggsave(
+#     plot = a/b/c, filename = "UN_selected_vars.png", path = "figures/", device = "png",
+#     width = 7, height = 7, dpi = 400, bg = "white"
+# )
 
@@
     filter(goal %in% c(10, 14, 15)) |> 
-    group_by(series_code, year, iso3) |>
+    group_by(series_name, year, iso3) |>
     # this gets rid of multiple values per year
@@
     mutate(nas = is.na(value)) |> 
-    ungroup() |> group_by(series_code, iso3) |> 
+    ungroup() |> group_by(series_name, iso3) |> 
    # group_by(indicator, indicator_name, iso3) |> 
@@
     ungroup() |> 
-    pivot_wider(names_from = series_code, values_from = missing) |> 
-    pivot_longer(cols = AG_LND_FRSTCERT:last_col(), names_to = "series_code",
+    pivot_wider(names_from = series_name, values_from = missing) |> 
+    pivot_longer(cols = `Average proportion of Freshwater Key Biodiversity Areas (KBAs) covered by protected areas (%)`:last_col(), names_to = "series_name",
                  values_to = "missing") |> 
-    group_by(series_code) |> 
+    group_by(series_name) |> 
     #filter(!any(is.na(missing))) |> 
-    # ggplot(aes(iso3, series_code)) +
+    # ggplot(aes(iso3, series_name)) +
     # geom_tile(aes(fill = missing)) +
@@
     filter(missing_countries < 45) |> 
-    ggplot(aes(missing_countries, series_code))+
-    geom_point()
-
-p2 + theme_light(base_size = 12)
-
-vars2 <- p2$data$series_code
+    ggplot(aes(missing_countries, series_name))+
+    geom_point() +
+    labs(tag = "A", y = "Series name", x = "Proportion of missing countries") +
+    theme_light(base_size = 6)
+p2
+vars2 <- p2$data$series_name
 ```
@@
     filter(iso3 %in% complete_countries) |> # use only countries
-    filter(series_code %in% vars2) |> 
+    filter(series_name %in% vars2) |> 
     ## this line is to get rid of multiple values per year
-    group_by(series_code, year, iso3) |>
+    group_by(series_name, year, iso3) |>
     summarize(mean_val = mean(value, na.rm = TRUE)) |> 
@@
     mutate(nas = is.na(value)) |> 
-    group_by(series_code, iso3) |> 
+    group_by(series_name, iso3) |> 
     summarise(missing = sum(nas) / diff(range(year))) |>  ## dividing by number of years (per country)
@@
     ## add missing values again for combination of indicator country:
-    ungroup() |> group_by(series_code) |>
+    ungroup() |> group_by(series_name) |>
     pivot_wider(names_from = iso3, values_from = missing) |> 
@@
     filter(!any(is.na(missing))) |> 
-    ggplot(aes(iso3, series_code)) +
+    ggplot(aes(iso3, series_name)) +
     #geom_point()
     geom_tile(aes(fill = missing)) +
-    scale_fill_viridis_c() +
-    theme_light() +
+    scale_fill_viridis_c("Missing",
+        guide = guide_colorbar(barwidth = unit(2,"mm"), barheight = unit(2,"cm"))) +
+    labs(tag = "B", y = "Series name", x = "Countries") +
+    theme_light(base_size = 6) +
     theme(axis.text.x = element_text(angle = 90))
@@
 complete_countries_sdg <- q2$data$iso3 |> unique()
+
+# ggsave(
+#     plot = p2/q2, filename = "UN_reduced_vars.png", path = "figures/", device = "png",
+#     width = 7, height = 7, dpi = 400, bg = "white"
+# )
+
 ```
@@
     filter(iso3 %in% complete_countries_sdg) |> # use only countries
-    filter(series_code %in% vars2) |> 
-    group_by(series_code, year, iso3) |>
+    filter(series_name %in% vars2) |> 
+    group_by(series_name, year, iso3) |>
     summarize(mean_val = mean(value, na.rm = TRUE)) |> 
@@
                  names_transform = list(year = as.numeric, value = as.numeric)) |> 
-    group_by(iso3, series_code) |> 
+    group_by(iso3, series_name) |> 
     mutate(value = na_interpolation(value, option = "spline"))
@@
 df_un_reduced |> ungroup() |> skimr::skim() # zero missing values
+```
+
+```{r}
+crg <- df_un_reduced |> 
+    mutate(series_name = str_replace(series_name, "Biodiversity", "Biodiversity\n"),
+           series_name = str_replace(series_name, "verified", "verified\n")) |> 
+    pivot_wider(values_from = value, names_from = series_name) |> 
+    GGally::ggpairs(
+        columns = 3:8,
+    lower = list(continuous = GGally::wrap("points", alpha = 0.5, size = 0.5)),
+    diag = list(continuous = GGally::wrap("densityDiag", alpha = 0.2))
+    # upper = list(continuous = GGally::wrap("cor", size = 2))
+    ) + theme_light(base_size = 4)
+
+crg
+
+# ggsave(
+#     plot = crg, filename = "UN_correlogram.png", path = "figures/", device = "png",
+#     width = 7, height = 7, dpi = 400, bg = "white"
+# )
 ```
@@
 # dat raw data from WorldBank
+
 ```
@@
 ## Ordination
+
+Following Ingo's advise, here I run a PCA with each variable_year as variable.
 
@@
 ```{r}
-biplot(pca1)
+# some notes on the pca object
+pca1$rotation |> dim() # 171 variables*years, 160 PCs because there was 160 countries to start with.
+pca1$x |> dim() # The loadings of 160 countries in each of the 160 components.
+pca1$center |> length() # 171 variable centers
+pca1$scale |> length() # 171 variable scales.
 ```
@@
 ```{r}
-# summary(pca1)
+pca1$rotation |> 
+    as_tibble(rownames = "vars") |> 
+    separate(vars, sep = "_", into = c("year", "variable")) |> 
+    mutate(year = as.numeric(year)) |> 
+    pivot_longer(cols = starts_with("PC"), names_to = "PC", values_to = "load") |> 
+    mutate(PC = str_remove(PC, "PC"), PC = as.numeric(PC)) |> 
+    filter(PC < 4) |> 
+    ggplot(aes(year, load)) +
+    geom_line(aes(color = variable)) +
+    facet_wrap(~PC) +
+    theme_light(base_size = 12)
+
+```
+
+```{r}
+biplot(pca1, var.axes = FALSE)
+```
+
+```{r}
+x <- summary(pca1, loadings = F, cutoff = 0.1)
+x <- x$importance |> as_tibble(rownames = "var") |> 
+    pivot_longer(cols = starts_with("PC"), names_to = "comp") |> 
+    mutate(comp = str_remove(comp, "PC") |> as.numeric()) |> 
+    dplyr::filter( comp <= 10 , var != "Standard deviation")
+
+ggplot(data = x |> filter(var == "Proportion of Variance"), aes(comp, value)) +
+    geom_col(color = "grey25") +
+    # geom_point(data = x |> filter(var == "Cumulative Proportion"), aes(comp, value)) +
+    # geom_line(data = x |> filter(var == "Cumulative Proportion"), aes(comp, value)) +
+    labs(x = "Principal components", y = "Explained variance") +
+    scale_x_continuous(n.breaks = 10) +
+    scale_y_continuous(labels = scales::label_percent()) +
+    theme_light(base_size = 6)
+
+pca1$x |> as_tibble() |> 
+    select(PC1:PC2) |> 
+    mutate(iso = df_dat$country_code |> unique()) |> 
+    ggplot(aes(PC1, PC2)) +
+    geom_text(aes(label = iso), size = 2) +
+    geom_segment(
+        data = pca1$rotation |>
+            as_tibble(rownames = "factor") |>
+            separate(factor, into = c("year", "factor")),
+        aes(xend = PC1, yend = PC2, x= 0, y = 0, color = factor, alpha = year)) +
+    coord_equal()
+
 ```
@@
 pca2 <- df_un_reduced |> 
-    unite(col = "var", series_code, year, sep = "_") |> 
+    unite(col = "var", series_name, year, sep = "_") |> 
     pivot_wider(values_from = value, names_from = var) |> 

### 2023-06-14 Trilemma_clustering.Rmd
paper revisions (graphs)

@@
 ```{r}
+## there are different country name spellings, so work only with iso codes and recover names later if you need them.
+countries <- read_csv(
+    file = "~/Documents/Projects/DATA/WorldBank/SDG_csv/SDGCountry.csv") |> 
+    janitor::clean_names()
 
@@
   rename(EFconsPerCap = value, country = countryName) %>% 
+    filter(!is.na(isoa2)) |> 
   filter(year >= 1995)
+
+## correct Sudan:
+dat$isoa2[dat$country == "Sudan (former)"] <- "SD"
+## get rid of country names
+dat <- dat |> select(-country)
+
+## add iso3 code
+dat <- dat |> 
+    left_join(countries |> select(iso3 = country_code, isoa2 = x2_alpha_code))
+
+dat$iso3[dat$isoa2 == "NA"] <- "NAM" # correct Namibia
 
@@
       filter(!is.na(gini_reported), year >= 1995) %>% 
-      select(country, isoa2 = c2, country_code = c3, year, gini_reported) %>% 
-      group_by(country, year, isoa2, country_code) %>% 
-      summarize(gini_mean = mean(gini_reported, na.rm = TRUE))
+      select(country, isoa2 = c2, iso3 = c3, year, gini_reported) %>% 
+      group_by(country, year, isoa2, iso3) %>% 
+      summarize(gini_mean = mean(gini_reported, na.rm = TRUE)) |> 
+        ungroup() |> select(-country) # remove country names
 
@@
       filter(year >= 1995) %>% 
-      rename(country = country_name, gni = value)
+      rename(country = country_name, gni = value, iso3 = country_code) |> 
+        ungroup() |> select(-country)
   )
@@
     mutate(ismissing = is.na(gini_mean)) %>% 
-    group_by(country) %>% 
+    group_by(iso3) %>% 
     summarize(missingness = sum(ismissing)/n()) %>% 
@@
     filter(missingness <= 0.3) %>% 
-    pull(country)
+    pull(iso3)
 
@@
     filter(missingness > 0.3) %>% 
-    pull(country)
+    pull(iso3)
 ```
@@
     select(-starts_with("indicator")) %>% 
-    filter(country %in% key_countries) %>% 
-    as_tsibble(key = country, index = year)
+    filter(iso3 %in% key_countries) %>% 
+    as_tsibble(key = iso3, index = year)
 
@@
 df_dat <- df_dat %>% 
-    group_by(country) %>% 
+    group_by(iso3) %>% 
     mutate(
@@
 
+Use this graph for trajectories of all countries later
+
 ```{r}
@@
     ggplot(aes(year, value)) +
-    geom_line(aes(group = country)) +
+    geom_line(aes(group = iso3)) +
     facet_wrap(~variable, ncol = 1, scales = "free") +
@@
 p1 <- df_dat %>% 
-    ggplot(aes(EFconsPerCap, gini, group = country)) +
-    geom_point(size = 0.5) +
+    ggplot(aes(EFconsPerCap, gini, group = iso3)) +
+    geom_point(size = 0.5, aes(color = year)) +
     geom_path(size = 0.2, aes(color = year)) +
@@
 p2<- df_dat %>% 
-  ggplot(aes(gni2, gini, group = country)) +
-  geom_point(size = 0.5, show.legend = FALSE) +
+  ggplot(aes(gni2, gini, group = iso3)) +
+  geom_point(size = 0.5, show.legend = FALSE, aes(color = year)) +
   geom_path(size = 0.2, aes(color = year)) +
@@
 p3<- df_dat %>% 
-  ggplot(aes(gni2, EFconsPerCap, group = country)) +
-  geom_point(size = 0.5, show.legend = FALSE) +
+  ggplot(aes(gni2, EFconsPerCap, group = iso3)) +
+  geom_point(size = 0.5, show.legend = FALSE, aes(color = year)) +
   geom_path(size = 0.2, aes(color = year))+
@@
 p1 + p2 + p3 + 
-    plot_layout(guides = "collect") & theme_light(base_size = 18) & 
+    plot_layout(guides = "collect") & theme_light(base_size = 12) & 
     theme(legend.position = "bottom", panel.grid = element_blank(),
@@
     as_tibble() %>% 
-    group_by(country) %>% 
+    group_by(iso3) %>% 
     summarize(
@@
     ungroup() |> 
-    group_by(country) |> 
+    group_by(iso3) |> 
     summarize(mgt = sum(magni, na.rm = TRUE), 
@@
     as_tibble() |> 
-    select(country, year, d_EF, d_gini, d_gni) |> 
+    select(iso3, year, d_EF, d_gini, d_gni) |> 
     pivot_longer(cols = starts_with("d_"), names_to = "vars", values_to = "vals") |> 
@@
     )) |> 
-    ggplot(aes(year, vals, group = country)) +
+    ggplot(aes(year, vals, group = iso3)) +
     geom_path(size = 0.2, alpha = 0.5) +
@@
     geom_point(aes(color = gni_log), size = 0.5) +
-    geom_path(aes(group = country), size = 0.1) +
+    geom_path(aes(group = iso3), size = 0.1) +
     scico::scale_color_scico("Prosperity (GNI [log])") +
@@
 
-c1 + theme_light(base_size = 18)
+c1 
 ```
@@
     geom_point(aes(color = EFconsPerCap), size = 0.5) +
-    geom_path(aes(group = country), size = 0.1) +
+    geom_path(aes(group = iso3), size = 0.1) +
     geom_hline(yintercept = 30, linetype = 2, color = "gray50") +
@@
     geom_point(aes(color = gini), size =0.5) +
-    geom_path(aes(group = country), size = 0.1) +
+    geom_path(aes(group = iso3), size = 0.1) +
     scico::scale_color_scico("Inequality (Gini)") +
@@
 
-c3 + theme_light(base_size = 18)
+c3 
 ```
@@
 
-#save(df_dat, file = "data/cleaned_trilemma.RData")
+# save(df_dat, file = "data/cleaned_trilemma-20230614.RData")
 ```
@@
 
-```{r fig.width=6, fig.height=2}
+```{r fig.width=6, fig.height=4}
 df_summary <- df_dat |> 
@@
     ungroup() |> 
-    group_by(country) |> 
+    group_by(iso3) |> 
     dplyr::summarize(sd_magnitude = sd(magni, na.rm = TRUE),
@@
         as.matrix(),
-    nClust = 3,
+    nClust = 2:9,
     clMethods = c(
@@
     mutate(clus = stab@clusterObjs$kmeans$`3`$cluster ) |> 
-    mutate(country = as_factor(country)) |> 
-    mutate(country = fct_reorder2(country, magnitude, Gini)) |> 
+    mutate(iso3 = as_factor(iso3)) |> 
+    mutate(iso3 = fct_reorder2(iso3, magnitude, Gini)) |> 
     pivot_longer(
@@
         values_to = "vals") |> 
-    ggplot(aes(vals, country)) +
+    ggplot(aes(vals, iso3)) +
     geom_col(aes(fill = as.factor(clus))) +
@@
 df_summary |>
-    select(-starts_with("country")) |> 
+    select(-starts_with("iso3")) |> 
     GGally::ggpairs()

### 2023-06-14 figures.R

@@ -4,7 +4,7 @@ library(plot3D)
 library(magrittr)
 
 
-load("data/cleaned_trilemma.RData")
+load("data/cleaned_trilemma-20230614.RData")
 
 #### Conceptual figure ####
 old_par <- par()
@@ -117,7 +117,7 @@ p0 <- ggplot() +
     labs(tag = "A")
   
 p6 <- df_dat %>% 
-    ggplot(aes( gini,EFconsPerCap , group = country)) +
+    ggplot(aes( gini,EFconsPerCap , group = iso3)) +
     geom_point(size = 0.15,aes(color = year), show.legend = FALSE) +
     geom_path(size = 0.1, aes(color = year)) +
     annotate(geom = "rect", ymin = 0, ymax = 1.68, xmin = 20, xmax = 30,
@@ -137,7 +137,7 @@ p6 <- df_dat %>%
           legend.key.width = unit(10, "mm"), legend.key.height = unit(3,"mm"))
 
 p5 <- df_dat %>% 
-    ggplot(aes( gini, gni2, group = country)) +
+    ggplot(aes( gini, gni2, group = iso3)) +
     geom_point(size = 0.15, show.legend = FALSE, aes(color = year)) +
     geom_path(size = 0.1, aes(color = year)) +
     annotate(geom = "rect", ymin = 10, ymax = 12746, xmin = 20, xmax = 30,
@@ -159,7 +159,7 @@ p5 <- df_dat %>%
 
 
 p4 <- df_dat %>% 
-    ggplot(aes( EFconsPerCap, gni2, group = country)) +
+    ggplot(aes( EFconsPerCap, gni2, group = iso3)) +
     geom_point(size = 0.15, show.legend = FALSE, aes(color = year)) +
     geom_path(size = 0.1, aes(color = year))+
     annotate(geom = "rect", ymin = 10, ymax = 12746, xmin = 0, xmax = 1.68,
@@ -193,7 +193,7 @@ p4 <- df_dat %>%
 (p4 + p5 + p6 + plot_layout(guides = "collect") & theme_light(base_size = 7) & 
     theme(legend.position = "bottom", panel.grid = element_blank(),
           legend.key.width = unit(10, "mm"), legend.key.height = unit(2,"mm")))
-`
+
 
 ggsave(
     plot = ((p0 + p3 + p2 + p1 + plot_layout(ncol = 4)) /
@@ -202,7 +202,7 @@ ggsave(
[…]

### 2023-06-14 inequality-SDGs.Rmd
national level analysis + wii new dataset

@@
 countries_list <- countries |> filter(!is.na(currency_unit)) |> pull(country_code)
+## correct Namibia, it is not an NA in x2_alpha_code
+
+countries$x2_alpha_code[countries$country_code == "NAM"] <- "NA"
 
@@
         guide = guide_colorbar(barwidth = unit(2,"mm"), barheight = unit(20, "mm"))) + 
-    labs(x = "Year", y = "Indicator name")
+    labs(x = "Year", y = "Indicator name") +
     theme_light(base_size = 6)
@@
 
+**J230323** Team recommended only to use inequality on UN Data which is available for ~50+ countries. Check newer dataset from WIID
+
+```{r}
+load("data/wii_small-data.RData")
+```
+
+```{r}
+ineq_dat
+```
+`m` is total, `a` is average. All other codes are documented here: [https://wid.world/codes-dictionary/#packages](https://wid.world/codes-dictionary/#packages).
+
 ### Correlations
@@
 df_dat |> ungroup() |> skimr::skim() # no missing values
+```
+
+```{r}
+## both datasets are wider format, so makes sense to merge as is
+df_dat <- df_dat |>
+    left_join(countries |> select(country_code, country = x2_alpha_code)) |> 
+    left_join(ineq_dat |> 
+                  select(-age, -pop) |> 
+                  unite(col = "var", c("variable", "percentile")) |> 
+                  pivot_wider(names_from = var, values_from = value)
+              )
 ```
@@
     theme_light(base_size = 6)
+b
 ```
@@
 ## Very long names:
-sdg|> 
+sdg |> 
     rename(indicator_name = series_description, year = time_period, iso3 =`ISO-alpha3 Code` ) |> 
@@
 
+Merge with WID inequality data:
+
+```{r}
+
+df_un_reduced <- df_un_reduced |> ungroup() |> #group_by(iso3)
+    left_join(
+        countries |> select(iso3 = country_code, country = x2_alpha_code) ) |> 
+    # filter(is.na(country)) |> pull(iso3) |> unique()
+    unite(col = "var", year, series_name, sep = "_") |> 
+    pivot_wider(values_from = value, names_from = var) |> 
+    left_join(
+        ineq_dat |> 
+            select(-age, -pop) |> 
+            filter(year >=  2000, year < 2019) |> # max year 2018 in UN dataset
+            unite(col = "series_name", c("variable", "percentile")) |> 
+            right_join(
+                countries |> 
+                    select(iso3 = country_code, country = x2_alpha_code) |> 
+                    filter(iso3 != "CHI") # filter out channel islands
+            ) |> 
+            unite(col = "var", year, series_name, sep = "_") |> 
+            pivot_wider(values_from = value, names_from = var) 
+        )
+```
+
 Now there is two working datasets cleaned and complete: `df_dat` is the data from World Bank, and `df_un_reduced` is the data from UN SDGs.
@@
 ```{r}
-rm(df_sdg, df_un, df_wb, maike_vars, other_vars)
+rm(df_sdg, df_un, df_wb, maike_vars, other_vars, ineq_dat)
 # key files
@@
 # dat raw data from WorldBank
-
+load("data/cleaned_sdg.RData")
 ```
@@
 pca1 <- df_dat |> 
-    select(-country_name) |> 
+    select(-country_name, -country, -sptinc992j_p0p100) |> 
+    filter(!is.na(rptinc992j_p0p100)) |> 
     pivot_longer(cols = `Access to electricity (% of population)`:last_col(),
@@
 ```{r}
-pca1$rotation |> 
+p <- pca1$rotation |> 
     as_tibble(rownames = "vars") |> 
@@
     mutate(PC = str_remove(PC, "PC"), PC = as.numeric(PC)) |> 
-    filter(PC < 4) |> 
+    filter(PC < 5) |> 
     ggplot(aes(year, load)) +
@@
     facet_wrap(~PC) +
-    theme_light(base_size = 12)
+    theme_light(base_size = 8)
 
+# ggsave(
+#     plot = p, file = "pca-wb-time.png", path = "figures/", device = "png",
+#     width = 7, height = 4, bg = "white", dpi = 400
+# )
+p + scale_color_brewer(palette = "Set3")
 ```
@@
     geom_col(color = "grey25") +
-    # geom_point(data = x |> filter(var == "Cumulative Proportion"), aes(comp, value)) +
-    # geom_line(data = x |> filter(var == "Cumulative Proportion"), aes(comp, value)) +
+    geom_point(data = x |> filter(var == "Cumulative Proportion"), aes(comp, value)) +
+    geom_line(data = x |> filter(var == "Cumulative Proportion"), aes(comp, value)) +
     labs(x = "Principal components", y = "Explained variance") +
@@
 
-pca1$x |> as_tibble() |> 
+a <- pca1$x |> as_tibble() |> 
     select(PC1:PC2) |> 
-    mutate(iso = df_dat$country_code |> unique()) |> 
+    mutate(iso = df_dat |> 
+               filter(!is.na(rptinc992j_p0p100)) |>
+               pull(country_code) |> unique()) |> 
     ggplot(aes(PC1, PC2)) +
     geom_text(aes(label = iso), size = 2) +
+    coord_equal() +
+    theme_light(base_size = 6)
+
+b <- ggplot() +
     geom_segment(
@@
             as_tibble(rownames = "factor") |> 
-            separate(factor, into = c("year", "factor")),
-        aes(xend = PC1, yend = PC2, x= 0, y = 0, color = factor, alpha = year)) +
-    coord_equal()
+            mutate(year = str_sub(factor, start = 1L, end = 4L),
+                   factor = str_remove(factor, pattern = "\\d{4}_"))  |> 
+            select(year, factor, PC1:PC2),
+        aes(xend = PC1, yend = PC2, color = factor, alpha = as.numeric(year)),
+        x = 0, y = 0,
+        arrow = arrow(length = unit(0.1,"cm")), linewidth = 0.3) +
+    scale_color_brewer("Variable",palette = "Set3",
+        guide = guide_legend(ncol = 2, title.position = "top")) +
+    scale_alpha("Year", guide = guide_legend(ncol = 1, title.position = "top")) +
+    coord_equal()+
+    theme_light(base_size = 6)
+
+a + b + plot_layout(guides = "collect") & 
+    theme_light(base_size = 10) &
+    theme(legend.position = "bottom")
+
+# ggsave(
+#     filename = "pca_wb-data.png", path = "figures/", dpi = 400, bg = "white",
+#     device = "png", width = 7, height = 4,
+#     plot = (a + b + plot_layout(guides = "collect") & 
+#     theme_light(base_size = 6) &
+#     theme(legend.position = "bottom") )
+# )
 
@@
 
+- 95% of the variability is explained by the first 10 PCs
+
+For the UN dataset, the area based nature proxies have zero variance (SD), which means that despite completness, it is not informative.
+
 ```{r}
+##
+zero_variance_vars <- df_un_reduced |>
+    pivot_longer(3:last_col(), names_to = "variable", values_to = "value") |> 
+    mutate(
+        year = str_sub(variable, 1L, 4L),
+        variable = str_sub(variable, 6L, -1L)
+    ) |> # pull(variable) |> unique()
+    filter(variable != "sptinc992j_p0p100") |>
+    group_by(iso3, variable) |> 
+    mutate(variance = var(value)) |>
+    filter(variance == 0) |>
+    pull(variable) |> unique()
+
 pca2 <- df_un_reduced |>
-    unite(col = "var", series_name, year, sep = "_") |> 
-    pivot_wider(values_from = value, names_from = var) |> 
     ungroup() |> 
-    select(-iso3) |> 
-    as.matrix(dimnames = list(df_un_reduced |> pull(iso3) |> unique())) |> 
+    select(-country, -iso3, -NA_NA) |> #skimr::skim()
+    select(-contains("sptinc"), -contains("gptinc"), -contains("rptinc")) |> 
     prcomp(center = TRUE, scale. = TRUE)
+
 ```
+**J230529:** It doesn't work, there is no enough variability in the data.
+It seems I have several variables / countries with zero variance or constants despite no missing values.
+However, it only screws up when adding the inequality variables.
+Without it does still run, why?
+Check with Maike.
+It might have to do with number of observations (countries) vs number of variables.
+Most likely too wide of a dataset can be problematic.
 
@@
 mfa1 <- df_un_reduced |> 
-    unite(col = "var", series_code, year, sep = "_") |> 
-    pivot_wider(values_from = value, names_from = var) |> 
-    ungroup() |> 
-    select(-iso3) |> 
+    select(-iso3, -country, -NA_NA) |> 
     as.matrix(dimnames = list(df_un_reduced |> pull(iso3) |> unique())) |> 
-    FactoMineR::MFA(group = rep(19, 6), type = rep("s", 6)) # 18 years times 6 variables
+    FactoMineR::MFA(group = rep(19, 10), type = rep("s", 10)) # 18 years times 6 variables
 ```
@@
     pca1,
-    labels = df_dat$country_code |> unique(),
+    labels = df_dat |>  select(-country_name, -country, -sptinc992j_p0p100) |> 
+    filter(!is.na(rptinc992j_p0p100)) |> pull(country_code) |> unique(),
     circle = TRUE,

### 2023-06-14 wii-db.Rmd
national level analysis + wii new dataset

---
title: "WIID database"
author: "Juan Rocha"
date: "Updated `r Sys.Date()`"
output:
  html_notebook:
    toc: yes
    toc_float: yes
    highlight: tango
    code_folding: hide
    df_print: paged
    theme: 
      bootswatch: cosmo
      code_font:
        google: Fira Code
editor_options: 
  chunk_output_type: inline
---

The World Income Inequality database has recently released new data for their
calculations of wealth and income inequality. Two resources are important:

- The [Code Dictionary](https://wid.world/codes-dictionary/): describes the 
rationale for new variables and how they are constructed. Currently there is over
1000 variables in the database
- The DINA guidelines is a more technical report (pdf) that explains the history
and legacy of the database, and describes the technical rationale for calculating
the different indicators. It is a good resource to understand limitations and 
caveats of each variable.

Both of these documents advise against using single number indicators (such as 
Gini) to describe inequality, and provide examples of when it fails. That is why
one needs the micro data, and why they make it available.

If we convert to PPP we need to use the last year of their convertion factors

> The WID contains two types of currency conversion factors (PPPs and market exchange rates) for three currencies (US dollars, euros and yuans). We provide long-term series for both of these indicators, but note that to convert between currencies in our benchmark approach, you only need to use the last value of each series: see Thomas Blanchet, Prices and currency conversions in WID.world, WID.world Technical Note Series n°2017/3 for details.

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  cache = TRUE, warning = FALSE, message = FALSE, echo = TRUE,
  fig.width = 8, fig.height = 5
)

library(fs)
library(tidyverse)
library(naniar)
library(plotly)
library(patchwork)

# time series
library(tsibble)
library(fable)
library(imputeTS)
library(slider)
library(tictoc)
# # clustering
# library(vegan)
# library(NbClust)
# library(clValid)
# library(mclust)

#theme_set(theme_light(base_size = 12))
```

Import data:

```{r}

fls <- dir_ls("~/Documents/Projects/DATA/WIID_World_Income_Inequality_DB/wid_all_data-20230323/")
# subseting countries only but there is subnational info for some nations (e.g. US)
fls <- fls |> str_subset(pattern = "WID_data_[:upper:]{2}\\.csv")

inq <- read_csv2(fls[44]) # 250 = US, the country with most data, 1093vas, 500k obs
inq                        # 44 Colombia, 215 Sweden, 211 South Africa

indicators <- map(fls, read_csv2) |> 
    map(function(x) x |> select(country,variable) |> unique()) |> 
    bind_rows()

indicators |> group_by(country) |> summarize(n=n()) |> arrange(desc(n)) # 277 countries
indicators |> group_by(variable) |> summarize(n = n() / 277) |> arrange(desc(n)) # 1377 variables

indicators |> 
    filter(str_detect(variable, "^g")) # g = gini, ptnic = pre-tax national income
```

Key variables:
- `ptninc` = pre-tax national income. It can be `g` for gini, `a` average, `b` inverted pareto-lorenz coefficient, `s` share, `r` ratio top 10 / bottom 50.
- `hweal` = net wealth, with same options as above.

```{r}
key_vars <- inq |> 
    pull(variable) |> 
    unique() |> 
    str_subset("(r|g|s)ptinc")

inq |> 
    filter(variable %in% key_vars) |># pull(percentile) |> unique()
    filter(percentile %in% c("p0p100", "p99p100")) |> 
    mutate(value = as.numeric(value)) |> 
    ggplot(aes(x = year, y = value)) +
    geom_line(aes(color = percentile), show.legend = FALSE) + 
    facet_wrap(~variable, scales = "free_y")
```

The ratio and gini are almost the same, at least for Colombia. Extract these variables for all files, all countries, all years:

```{r}
tic()
ineq_dat <- map(fls, read_csv2) |> 
    map(function(x) {
        x |> 
            filter(variable %in% key_vars) |># pull(percentile) |> unique()
            filter(percentile %in% c("p0p100", "p99p100")) |> 
            mutate(value = as.numeric(value), age = as.numeric(age))
        }) |> 
    bind_rows()
toc()

ineq_dat |> 
    filter(variable != "sptinc992j") |> 
    ggplot(aes(x = year, y = value)) +
    geom_line(aes(color = country), show.legend = FALSE) + 
    facet_wrap(~variable, scales = "free_y")
```

```{r}
ineq_dat |> 
    filter(variable != "sptinc992j") |> 
    pivot_wider(names_from = variable, values_from = value) |> 
    ggplot(aes(gptinc992j, rptinc992j)) +
    geom_point(aes(color = country), show.legend = FALSE, alpha = 0.5)

```

```{r}
library(naniar)
ineq_dat |> 
    filter(year >= 1980) |> 
    filter(variable != "sptinc992j") |> 
    ggplot(aes(year, value)) +
    geom_miss_point(aes(color = country), show.legend = FALSE, size = 0.1, alpha = 0.3) +
    geom_line(aes(color = country), show.legend = FALSE, linewidth = 0.1) +
    facet_wrap(~variable, scales = "free_y")
```

Ha! no missing values!!Add wealth variables and export for analysis.

```{r}
save(ineq_dat, file = "data/wii_small-data.RData")
```

### 2023-10-09 Trilemma_clustering.Rmd
fixed PCA, initial explroations with MFA

@@
 pc <- prcomp(df_summary |> 
-                 select(-starts_with("country")),
+                 select(-starts_with("iso3")),
              center = TRUE, scale. = TRUE)
 ## Fix short names for plotting
-df_summary <- df_summary |> 
-    left_join(
-        dat |> select(country, country_code) |> 
-            filter(!is.na(country_code)) |> unique())
+# df_summary <- df_summary |> 
+#     left_join(
+#         dat |> select(country, country_code) |> 
+#             filter(!is.na(country_code)) |> unique())
 
@@
     pc, groups = as.character(stab@clusterObjs$kmeans$`3`$cluster),
-    ellipse = TRUE, labels = df_summary$country_code, labels.size = 1,
+    ellipse = TRUE, labels = df_summary$iso3, labels.size = 1,
     varname.size = 1.5) +
@@
     as_tibble() |> 
-    group_by(country) |> 
+    group_by(iso3) |> 
     dplyr::summarize(
@@
     mutate(Group = as.character(stab@clusterObjs$kmeans$`3`$cluster)) |> 
-    left_join(
-        dat |> select(country, country_code) |> 
-            filter(!is.na(country_code)) |> unique()) |> 
+    # left_join(
+    #     dat |> select(country, country_code) |> 
+    #         filter(!is.na(country_code)) |> unique()) |> 
     ggplot(aes(EF, Gini)) +
@@
     geom_text(
-        aes(x = EF *1.1, y = Gini * 1.1, label = country_code, colour = Group), 
+        aes(x = EF *1.1, y = Gini * 1.1, label = iso3, colour = Group), 
         size = 1, show.legend = FALSE) +
@@
     mutate(Group = as.character(stab@clusterObjs$kmeans$`3`$cluster)) |> 
-    left_join(
-        dat |> select(country, country_code) |> 
-            filter(!is.na(country_code)) |> unique()) |> 
+    # left_join(
+    #     dat |> select(country, country_code) |> 
+    #         filter(!is.na(country_code)) |> unique()) |> 
     ggplot(aes(GNI, Gini)) +
@@
     geom_text(aes(x = GNI * 1.1, y = Gini *1.1, 
-                  label = country_code, colour = Group), 
+                  label = iso3, colour = Group), 
               size = 1, show.legend = FALSE) +
@@
 
-df_all2 <- df_all |> ungroup() |> 
+df_all2 <- df_all2 |> ungroup() |> 
     filter(!is.na(sd_gini), !is.na(m_ef), !is.na(sd_ef), !is.na(m_gni), !is.na(sd_gni))
+
+save(df_all2, file = "data/trilemma_static.Rda")
 ```
@@
     scico::scale_fill_scico(
-        "Mean Ecological Footprint",palette = "romaO",
+        "Mean Ecological Footprint",palette = "romaO", direction = -1,
         guide = guide_colorbar(title.position = "top")) + 
@@
     filename = "pca_maps.png", path = "figures/", device = "png",
-    width = 5, height = 4, dpi = 400, bg = "white",
+    width = 5, height = 4, dpi = 600, bg = "white",
     plot = (pc1 / pc2 ) | (ef_map / gini_map / gni_map) + 

### 2023-10-09 inequality-SDGs.Rmd
fixed PCA, initial explroations with MFA

@@
 ineq_dat
+
+## correct Namibia is not a NA, but the country code is NA
+ineq_dat <- ineq_dat |> 
+    mutate(country = case_when(is.na(country) ~ "NA", TRUE ~ country)) #|> filter(country == "NA")
 ```
@@
               )
+
+df_dat
 ```
@@
         countries |> select(iso3 = country_code, country = x2_alpha_code) ) |> 
-    # filter(is.na(country)) |> pull(iso3) |> unique()
+    filter(iso3 != "FJI") |> # drop out Fiji, it does not exist in ineq data and creates NAs
+    #filter(is.na(country)) |> pull(iso3) |> unique()
     unite(col = "var", year, series_name, sep = "_") |> 
@@
             unite(col = "series_name", c("variable", "percentile")) |>
+            filter(series_name != "sptinc992j_p0p100") |> # all ones, no variability
             right_join(
@@
             unite(col = "var", year, series_name, sep = "_") |> 
-            pivot_wider(values_from = value, names_from = var) 
+            pivot_wider(values_from = value, names_from = var) |> 
+            select(-NA_NA)
         )
+
+### No missing value in any variable!
+df_un_reduced |>
+    pivot_longer(3:last_col(), names_to = "variable", values_to = "value") |> 
+    mutate(
+        year = str_sub(variable, 1L, 4L),
+        variable = str_sub(variable, 6L, -1L)
+    ) |> 
+    filter(is.na(value))
+
 ```
@@
 ```{r}
-#save(df_dat, df_un_reduced, file = "data/cleaned_sdg.RData")
+# save(df_dat, df_un_reduced, file = "data/cleaned_sdg.RData")
 ```
@@
 # some notes on the pca object
-pca1$rotation |> dim() # 171 variables*years, 160 PCs because there was 160 countries to start with.
-pca1$x |> dim() # The loadings of 160 countries in each of the 160 components.
-pca1$center |> length() # 171 variable centers
-pca1$scale |> length() # 171 variable scales.
+pca1$rotation |> dim() # 228 variables*years, 150 PCs because there was 150 countries after filtering missing values in rptinc992j_p0p100.
+pca1$x |> dim() # The loadings of 150 countries in each of the 150 components.
+pca1$center |> length() # 228 variable centers
+pca1$scale |> length() # 228 variable scales.
 ```
@@
 pca2 <- df_un_reduced |>
-    ungroup() |> 
-    select(-country, -iso3, -NA_NA) |> #skimr::skim()
-    select(-contains("sptinc"), -contains("gptinc"), -contains("rptinc")) |> 
+    select(-country, -iso3) |>
     prcomp(center = TRUE, scale. = TRUE)
-
 ```
@@
 
+**J231009** Now it's working, the problem was not the zero or low variance, the problem was NA induced by two countries that didn't match on the joints, Namibia whose iso2 is NA and got interpreted as missing value, and Fiji that does not exist on the WII database.
+So only 50 countries at the end.
+
 ```{r}
@@
 ```
+
+```{r}
+x <- summary(pca2, loadings = F, cutoff = 0.1)
+x <- x$importance |> as_tibble(rownames = "var") |> 
+    pivot_longer(cols = starts_with("PC"), names_to = "comp") |> 
+    mutate(comp = str_remove(comp, "PC") |> as.numeric()) |> 
+    dplyr::filter( comp <= 10 , var != "Standard deviation")
+
+ggplot(data = x |> filter(var == "Proportion of Variance"), aes(comp, value)) +
+    geom_col(color = "grey25") +
+    geom_point(data = x |> filter(var == "Cumulative Proportion"), aes(comp, value)) +
+    geom_line(data = x |> filter(var == "Cumulative Proportion"), aes(comp, value)) +
+    labs(x = "Principal components", y = "Explained variance") +
+    scale_x_continuous(n.breaks = 10) +
+    scale_y_continuous(labels = scales::label_percent()) +
+    theme_light(base_size = 6)
+
+a <- pca2$x |> as_tibble() |> 
+    select(PC1:PC2) |> 
+    mutate(iso = df_un_reduced |> pull(iso3) |> unique() ) |> 
+    ggplot(aes(PC1, PC2)) +
+    geom_text(aes(label = iso), size = 2) +
+    coord_equal() +
+    theme_light(base_size = 6)
+
+b <- ggplot() +
+    geom_segment(
+        data = pca2$rotation |>
+            as_tibble(rownames = "factor") |> 
+            mutate(year = str_sub(factor, start = 1L, end = 4L),
+                   factor = str_remove(factor, pattern = "\\d{4}_"))  |> 
+            select(year, factor, PC1:PC2),
+        aes(xend = PC1, yend = PC2, color = factor, alpha = as.numeric(year)),
+        x = 0, y = 0,
+        arrow = arrow(length = unit(0.1,"cm")), linewidth = 0.3) +
+    scale_color_brewer("Variable",palette = "Set3",
+        guide = guide_legend(ncol = 2, title.position = "top")) +
+    scale_alpha("Year", guide = guide_legend(ncol = 1, title.position = "top")) +
+    xlim(c(-0.1, 0.15)) +
+    theme_light(base_size = 6)
+
+a + b + plot_layout(guides = "collect") & 
+    theme_light(base_size = 10) &
+    theme(legend.position = "bottom")
+
+# ggsave(
+#     filename = "pca_wb-data.png", path = "figures/", dpi = 400, bg = "white",
+#     device = "png", width = 7, height = 4,
+#     plot = (a + b + plot_layout(guides = "collect") & 
+#     theme_light(base_size = 6) &
+#     theme(legend.position = "bottom") )
+# )
+
+```
+
 ### Multiple factor analysis
@@
 ```{r}
+df_un_reduced |> 
+    pivot_longer(3:last_col(), names_to = "variable", values_to = "value") |> 
+    separate(variable, into = c("year", "indicator"), sep = "_") |> 
+    skimr::skim() # 19 years, 9 variables
+
 mfa1 <- df_un_reduced |> 
-    select(-iso3, -country, -NA_NA) |> 
+    select(-iso3, -country) |> 
     as.matrix(dimnames = list(df_un_reduced |> pull(iso3) |> unique())) |> 
-    FactoMineR::MFA(group = rep(19, 10), type = rep("s", 10)) # 18 years times 6 variables
+    FactoMineR::MFA(group = rep(19, 9), type = rep("s", 9)) # 19 years times 9 variables
 ```
@@
 ```{r}
-mfa1 |> summary()
+# df_dat$year |> unique() |> length() # 19yrs
+
+mfa2 <- df_dat |> 
+    select(-country_name, -country, -sptinc992j_p0p100) |> 
+    filter(!is.na(rptinc992j_p0p100)) |> 
+    pivot_longer(cols = `Access to electricity (% of population)`:last_col(),
+                 values_to = "value", names_to = "variable") |> #pull(variable) |> unique() 12 vars
+    unite(col = "var", year, variable, sep = "_") |> 
+    pivot_wider(names_from = var, values_from = value) |> 
+    ungroup() |> 
+    select(-country_code) |> 
+    as.matrix(dimnames = list(df_dat |> pull(country_code) |> unique())) |> 
+    FactoMineR::MFA(group = rep(19, 12), type = rep("s", 12)) # 19 years times 9 variables
 ```
@@
 
+```{r}
+ggbiplot::ggbiplot(
+    pca2,
+    labels = df_un_reduced |> pull(iso3) |> unique(),
+    circle = TRUE,
+    varname.abbrev = TRUE)
+```
+
+Extract the weight per country on the relevant components to see how they move on PCA or MFA space.
+
+```{r}
+mfa1 |> summary()
+mfa2$summary.quanti
+```
+
 ## To dos:
@@
 ```{r}
-pca3 <- df_dat |> 
-    ungroup() |> 
-    mutate(year = as_factor(year)) |> 
-    split(~year) |>
-    map(function(x) select(x, where(is.numeric))) |> 
-    map(function(x) princomp(x, center = TRUE, scale. = TRUE))
-```
-
-```{r}
-pca3 |> map(biplot)
+# pca3 <- df_dat |> 
+#     ungroup() |> 
+#     mutate(year = as_factor(year)) |> 
+#     split(~year) |>
+#     map(function(x) select(x, where(is.numeric))) |> 
+#     map(function(x) princomp(x, center = TRUE, scale. = TRUE))
+```
+
+```{r}
+# pca3 |> map(biplot)
 ```

### 2023-10-09 figures.R

@@ -1,10 +1,11 @@
-library(ggplot2)
+library(tidyverse)
 library(patchwork)
 library(plot3D)
 library(magrittr)
 
 
 load("data/cleaned_trilemma-20230614.RData")
+load("data/trilemma_static.Rda")
 
 #### Conceptual figure ####
 old_par <- par()
@@ -379,12 +380,15 @@ ggsave(
 )
 
 #### surface ####
+old_par <- par()
 
+quartz(width = 3.5, height = 3.5, pointsize = 7)
+#par(mar = c(0,0,0,0))
 
 with(df_dat, {
     
     #linear regression:
-    fit <- glm(EFconsPerCap  ~ gni2 + gini + I(gni2^2) + I(gini^2), 
+    fit <- glm(EFconsPerCap  ~ gni2 * gini + I(gni2^2) + I(gini^2), 
                family = "gaussian", data = df_dat)
     
     ## predict values
@@ -397,12 +401,20 @@ with(df_dat, {
     
     ## plot
     scatter3D(
-        z = EFconsPerCap, y = gini, x = gni2, pch = 1, theta = 0, phi = 20, 
+        z = EFconsPerCap, y = gini, x = gni2, pch = 1, theta = -30, phi = 20, 
         ticktype = "detailed", zlab = "Ecological footprint", ylab = "Inequality (Gini)",
         xlab = "Prosperity (GNI[log])",
         surf = list(z = EF_pred, y = gini_pred, x = gni_pred, alpha = 0.5, col = "grey40")
     )
 })
+
+quartz.save(
+    file="figures/sm_cube.png", type = "png", dpi = 500,
+    width = 3.5, height = 3.5, pointsize = 7, bg = "white"
+)
+dev.off()
+par(old_par)
+
 
 with(df_dat, {
     scatter3D(
@@ -447,25 +459,159 @@ null3 <- glm(EFconsPerCap ~ gini + I(gini^2) ,
 
 summary(nokc)
 summary(null3)
+
+
 #### concpetual figures SI ####
 
-ggplot(data = tibble(x = seq(-10,10,0.1)), aes(x)) +
-    geom_function(fun = function(x) -x^2 )
+a <- ggplot(data = tibble(x = seq(-10,10,0.1)), aes(x)) +
+    geom_function(fun = function(x) -x^2 ) +
+    annotate(geom = "segment", x = -10, xend = 0, y =0, yend = 0, color = "orange", linetype = 2) +
[…]

### 2024-01-16 inequality-SDGs.Rmd
back up

@@
     pivot_longer(cols = `Access to electricity (% of population)`:last_col(),
-                 values_to = "value", names_to = "variable") |> #pull(variable) |> unique() 12 vars
+                 values_to = "value", names_to = "variable") |> 
+    arrange(variable, year) |> 
     unite(col = "var", year, variable, sep = "_") |> 
@@
 
+```{r}
+# save(pca1, pca2, mfa1, mfa2, file = "data/ordination_results.Rda")
+```
+
 ## To dos:
 
-- one PCA per year, see how PCA1 change over time.
+- one PCA per year, see how PCA1 change over time?
 
@@
 ```
+
+- Clustering analysis on the reduced dimensions (PCA, MFA)
+- Plot ineq against biodiv proxies, does the path foward differ from the path backwards = hysteresis?
 

### 2024-01-16 paper/manuscript.Rmd
back up

@@
 
+```{r data, include=FALSE}
+library(tidyverse)
+library(here)
+
+## word counts
+## Abstract 150
+## Intro 697
+## Method 957
+## Results 621
+## Discussion 1434
+## Conclusion 267
+## Total main 3976 + 150 abstract
+## 
+## Targets: 
+# 
+# - NatComms: 5000w | 150w abstract | 4 figs or tables | Methods end 3000w | Refs
+# 70max
+# 
+# - WorldDev: | 4500-10k words including all (refs, captions, tables, etc) | 300w
+# abstract
+```
+
 # Outline
@@
 / People and Nature?
-/ PNAS / ERL [4000w]
+/ PNAS / ERL / NatSust?
+[4000w]
 
@@
 
--
+-   download and check world inequality db again
 
@@
 
-```{r echo = FALSE}
+```{r echo = FALSE, message=FALSE, warning=FALSE}
 dat <- read_csv(file = "~/Documents/Projects/DATA/WorldBank/SDG_csv/SDGData.csv") |> 
@@
 We focused subsequent analysis only on country level indicators for SDGs 10, 14 and 15, related to inequality and the biosphere respectively.
+We complemented the SDGs datasets with inequality data from the World Inequality database (WID), using their estimates of the pre-tax national income computed as ratio or Gini.
 
 ```{r echo = FALSE}
-load(file = "data/cleaned_sdg.RData")
+load(file = paste0(here(), "/data/cleaned_sdg.RData"))
 # `df_dat` is the data from World Bank, and `df_un_reduced` is the data from UN SDGs
@@
 We discarded indicators for which time series contained more than 30% of missing values, or less than 45 countries.
-Missing values were then imputed using a cubic spline, leaving us with `r df_dat |> pull(country_code) |> unique() |> length()` countries, `r df_dat |> pull(year) |> unique() |> length()` years of data across 9 indicators for the WB dataset; and `r df_un_reduced |> pull(series_code) |> unique() |> length()` indicators, `r df_un_reduced |> pull(iso3) |> unique() |> length()` countries over `r df_un_reduced |> pull(year) |> unique() |> length()` years for the UN dataset.
-Table X summarizes our variable selection, their units and available ranges.
+Missing values were then imputed using a cubic spline, leaving us with `r df_dat |> pull(country_code) |> unique() |> length()` countries, `r df_dat |> pull(year) |> unique() |> length()` years of data across 9 indicators for the WB dataset; and `r df_un_reduced |> pull(series_name) |> unique() |> length()` indicators, `r df_un_reduced |> pull(iso3) |> unique() |> length()` countries over `r df_un_reduced |> pull(year) |> unique() |> length()` years for the UN dataset.
+The UN dataset was further reduced to 50 countries because the WID does not report inequality timeseries for Fiji.
 
-*Ordination:* We used multiple factor analysis and principal component analysis over time to ...
+*Ordination:* We used multiple factor analysis and principal component analysis to reduce the dimensionality of the data and explore similarities and differences across countries.
 
 Notes:
-
--   Add Gini to the datasets!
-    repeat the analysis
-
--   Add the SM figures, they are ready
 
@@
 
-\# References {.unnumbered}
+# References {.unnumbered}
 
@@
 
-\\pagebreak
+\pagebreak
 
 # Supplementary Material {#sec:SM}
+
+```{=tex}
+\renewcommand\thefigure{S\arabic{figure}}
+\renewcommand\thetable{S\arabic{table}}
+\setcounter{table}{0}
+\setcounter{figure}{0}
+```
+```{=tex}
+\begin{figure*}[ht]
+\centering
+\includegraphics[width = 6in, height = 6in]{WB_environment_SDGs}
+\caption{\textbf{World Bank environmental SDGs} }
+\label{fig:SM_WB_env}
+\end{figure*}
+```
+```{=tex}
+\begin{figure*}[ht]
+\centering
+\includegraphics[width = 6in, height = 6in]{WB_environment_SDGs_othervars}
+\caption{\textbf{World Bank environmental SDGs additional variables} }
+\label{fig:SM_WB_env2}
+\end{figure*}
+```
+```{=tex}
+\begin{figure*}[ht]
+\centering
+\includegraphics[width = 6in, height = 6in]{WB_selected_vars}
+\caption{\textbf{World Bank selected variables} }
+\label{fig:SM_WB_select}
+\end{figure*}
+```
+```{=tex}
+\begin{figure*}[ht]
+\centering
+\includegraphics[width = 6in, height = 5in]{WB_correlogram}
+\caption{\textbf{Correlogram of World Bank selected variables} }
+\label{fig:SM_WBcorr}
+\end{figure*}
+```
+```{=tex}
+\begin{figure*}[ht]
+\centering
+\includegraphics[width = 6in, height = 5in]{UN_SDG_10-14-15}
+\caption{\textbf{United Nations SDGs 10, 14 and 25 dataset} }
+\label{fig:SM_UNdata}
+\end{figure*}
+```
+```{=tex}
+\begin{figure*}[ht]
+\centering
+\includegraphics[width = 6in, height = 5in]{UN_selected_vars}
+\caption{\textbf{UN selected variables} }
+\label{fig:SM_UNselect}
+\end{figure*}
+```
+```{=tex}
+\begin{figure*}[ht]
+\centering
+\includegraphics[width = 6in, height = 5in]{UN_reduced_vars}
+\caption{\textbf{UN reduced variables} }
+\label{fig:SM_UNred}
+\end{figure*}
+```
+```{=tex}
+\begin{figure*}[ht]
+\centering
+\includegraphics[width = 6in, height = 5in]{UN_correlogram}
+\caption{\textbf{Correlogram of United Nations selected variables} }
+\label{fig:SM_UNcorr}
+\end{figure*}
+```

### 2024-01-16 03-clustering.R

@@ -1,0 +1,28 @@
+library(tidyverse)
+
+load("data/ordination_results.Rda")
+
+m <- "ward.D2" # minimize the total within-cluster variance
+
+clust_num <- NbClust::NbClust( 
+    data = (pca1$x[,1:10] %>% as.data.frame() %>% 
+                select(where(is.numeric)) ),
+    #distance = "maximum",
+    min.nc = 2, max.nc = 10, 
+    method = m, 
+    index = 'all')
+
+## Stability & Internal validation
+stab <- clValid::clValid(
+    obj = pca1$x[,1:10] ,
+    nClust = 2:10,
+    clMethods = c(
+        "hierarchical", "kmeans",  "som", 
+        "model", "diana", "sota", "pam", "clara", "agnes"),
+    validation = c('stability', "internal"),
+    #metric = "manhattan",
+    method = "ward.D2",
+    verbose = FALSE
+) ## Hierarchical is the optimal method
+
+clValid::optimalScores(stab)

### 2024-01-16 04-exploration.R

@@ -1,0 +1,29 @@
+library(tidyverse)
+
+load("data/cleaned_sdg.RData")
+
+df_dat |> 
+    ggplot(aes(year, rptinc992j_p0p100)) +
+    geom_line(aes(group = country_code))
+
+df_dat |> 
+    ggplot(aes(rptinc992j_p0p100, `Access to electricity (% of population)`)) +
+    geom_point(aes(color = year), size = 0.5) +
+    geom_line(aes(group = country), linewidth = 0.4)
+
+df_un_reduced |>
+    pivot_longer(3:last_col(), names_to = "variable", values_to = "value") |> 
+    mutate(
+        year = str_sub(variable, 1L, 4L),
+        variable = str_sub(variable, 6L, -1L),
+        year = as.numeric(year)
+    ) |> 
+    pivot_wider(names_from = variable, values_from = value) |> 
+    ggplot(aes(`rptinc992j_p0p100`, `Average proportion of Terrestrial Key Biodiversity Areas (KBAs) covered by protected areas (%)` )) +
+    geom_point(aes(color = year)) + 
+    geom_line(aes(group = iso3))
+
+ineq_sum <- df_dat |> 
+    group_by(country_name) |> 
+    summarize(model = lm(rptinc992j_p0p100~year))
+

### 2024-01-16 05-test_idea.R

@@ -1,0 +1,18 @@
+library(tidyverse)
+
+load("data/cleaned_trilemma-20230614.RData")
+
+lms <- df_dat |> group_by(iso3) |> 
+    summarize(model = list(lm(EFconsPerCap ~ year)))
+
+map(.x=lms$model, ~coef(.x)[2])
+
+lms <- lms |> 
+    mutate(slope = map_dbl(model, function(x) coef(x)[2]))
+
+df_dat |> 
+    left_join(lms |> select(-model)) |> 
+    ggplot(aes(gini, EFconsPerCap)) +
+    geom_line(aes(group = iso3, color = slope)) +
+    scale_color_viridis_c() +
+    theme_light()

### 2024-01-16 figures.R

@@ -215,14 +215,16 @@ c3 <- df_dat |>
     # scale_fill_viridis_c() + 
     #geom_density_2d(color = "white", size = 0.2) +
     geom_density_2d_filled(alpha = 0.95, show.legend = FALSE) +
-    geom_point(aes(color = gni_log), size = 0.05, alpha = 0.25) +
-    geom_path(aes(group = iso3, color = gni_log), size = 0.05, alpha = 0.75) +
+    geom_point(aes(color = gni_log), size = 0.01, alpha = 0.15) +
+    geom_path(aes(group = iso3, color = gni_log), size = 0.01, alpha = 0.5) +
     scico::scale_color_scico(
         "Prosperity (GNI [log])", palette = "grayC", direction = -1,
         guide = guide_colorbar(title.position = "top")) +
     geom_hline(yintercept = 1.68, linetype = 2, color = "gray90", size = 0.2) +
     geom_vline(xintercept = 30, linetype = 2, color = "gray90", size = 0.2) +
     labs(x = "Inequality (Gini)", y = "Ecological footprint") +
+    annotate("text", x = c(32, 50), y = c(10,5), label = c("Regime 1", "Regime 2"), 
+             colour = "white", size = 2) +
     theme_light(base_size = 7) +
     theme(legend.position = "top", panel.grid = element_blank(),
           legend.key.width = unit(7, "mm"), legend.key.height = unit(2,"mm"))
@@ -230,14 +232,16 @@ c3 <- df_dat |>
 c2 <- df_dat |> 
     ggplot(aes(gini,gni2)) +
     geom_density_2d_filled(alpha = 0.95, show.legend = FALSE) +
-    geom_point(aes(color = EFconsPerCap), size = 0.01, alpha = 0.25) +
-    geom_path(aes(group = iso3, color = EFconsPerCap), size = 0.05, alpha = 0.75) +
+    geom_point(aes(color = EFconsPerCap), size = 0.01, alpha = 0.15) +
+    geom_path(aes(group = iso3, color = EFconsPerCap), size = 0.01, alpha = 0.5) +
     geom_hline(yintercept = (12546), linetype = 2, color = "gray90", size = 0.2) +
     geom_vline(xintercept = 30, linetype = 2, color = "gray90", size = 0.2) +
     scico::scale_color_scico(
         "Ecological footprint",palette = "grayC", direction = 1,
         guide = guide_colorbar(title.position = "top")) +
[…]

### 2024-03-05 inequality-SDGs.Rmd
national analysis first steps

@@
     labs(x = "Mean proportion of missing values across countries", y = "Indicator name", tag = "A")+
-    theme_light(base_size = 6)
+    theme_light(base_size = 8)
 
@@
         "Proportion of missing values",
-        guide = guide_colorbar(barwidth = unit(2,"cm"), barheight = unit(2,"mm"))) + 
+        guide = guide_colorbar(barwidth = unit(2,"cm"), barheight = unit(5,"mm"))) + 
     labs(x = "Country code", y = "Indicator name", tag = "B") +
-    theme_light(base_size = 6) +
+    theme_light(base_size = 8) +
     theme(axis.text.x = element_text(angle = 90, size = 3), 
@@
 # ggsave(
-#     plot = p/q, filename = "WB_selected_vars.png", path = "figures/", device = "png",
-#     width = 7, height = 7, dpi = 400, bg = "white"
+#     plot = p/q, filename = "WB_selected_vars.png", path = "paper/figures/", device = "png",
+#     width = 7.5, height = 7, dpi = 400, bg = "white"
 # )
@@
 ```
-`m` is total, `a` is average. All other codes are documented here: [https://wid.world/codes-dictionary/#packages](https://wid.world/codes-dictionary/#packages).
+`m` is total, `a` is average. All other codes are documented here: [https://wid.world/codes-dictionary/#packages](https://wid.world/codes-dictionary/#packages). The selected variables are:
+
+- `rptinc992j_p0p100` which means the ratio of the pretax national income adults (pop > 20yrs) of the top 10% / the bottom 50%
+
+- `sptinc992j_p99p100`: share of pretax national income adults equal-split for the top 1%
+
+- `gptinc992j_p0p100`: Gini coefficient for the pretax national income for the adult population
 
@@
     # upper = list(continuous = GGally::wrap("cor", size = 2))
-    ) + theme_light(base_size = 4)
+    ) + theme_light(base_size = 8)
 
@@
 # ggsave(
-#     plot = crg, filename = "WB_correlogram.png", path = "figures/", device = "png",
+#     plot = crg, filename = "WB_correlogram.png", path = "paper/figures/", device = "png",
 #     width = 7, height = 7, dpi = 400, bg = "white"
@@
          x = "Year", y = "Indicator name", tag = "A") +
-    theme_light(base_size = 6)
+    theme_light(base_size = 8)
 a
@@
     labs(title = "SDG14: Life under water", x = "Year", y = "Indicator name", tag = "B") +
-    theme_light(base_size = 6)
+    theme_light(base_size = 8)
 b
@@
     labs(title = "SDG15: Life on land", x = "Year", y = "Indicator name", tag = "C") +
-    theme_light(base_size = 6)
+    theme_light(base_size = 8)
 c
@@
 # ggsave(
-#     plot = a/b/c, filename = "UN_selected_vars.png", path = "figures/", device = "png",
-#     width = 7, height = 7, dpi = 400, bg = "white"
+#     plot = a/b/c + plot_layout(heights = c(1,1,1.5)), 
+#     filename = "UN_selected_vars.png", path = "paper/figures/", device = "png",
+#     width = 7, height = 8, dpi = 400, bg = "white"
 # )
@@
     filter(missing_countries < 45) |> 
+    mutate(series_name = str_replace(series_name, " covered", "\ncovered"),
+           series_name = str_replace(series_name, " certification", "\ncertification")) |> 
     ggplot(aes(missing_countries, series_name))+
@@
     labs(tag = "A", y = "Series name", x = "Proportion of missing countries") +
-    theme_light(base_size = 6)
+    theme_light(base_size = 8)
 p2
@@
     filter(!any(is.na(missing))) |> 
+    mutate(series_name = str_replace(series_name, " covered", "\ncovered"),
+           series_name = str_replace(series_name, " certification", "\ncertification")) |> 
     ggplot(aes(iso3, series_name)) +
@@
     labs(tag = "B", y = "Series name", x = "Countries") +
-    theme_light(base_size = 6) +
-    theme(axis.text.x = element_text(angle = 90))
+    theme_light(base_size = 8) +
+    theme(axis.text.x = element_text(angle = 90, size = 4))
 
@@
 # ggsave(
-#     plot = p2/q2, filename = "UN_reduced_vars.png", path = "figures/", device = "png",
-#     width = 7, height = 7, dpi = 400, bg = "white"
+#     plot = p2/q2, filename = "UN_reduced_vars.png", 
+#     path = "paper/figures/", device = "png",
+#     width = 7, height = 7.5, dpi = 400, bg = "white"
 # )
@@
 # ggsave(
-#     plot = crg, filename = "UN_correlogram.png", path = "figures/", device = "png",
+#     plot = crg, filename = "UN_correlogram.png", path = "paper/figures/", device = "png",
 #     width = 7, height = 7, dpi = 400, bg = "white"
@@
     geom_line(aes(color = variable)) +
-    facet_wrap(~PC) +
-    theme_light(base_size = 8)
-
+    labs(y = "Component loadings", x = "Year", title = "Top 4 principal components",
+         caption = "Data source: World Bank and World Inequality Database") +
+    facet_wrap(~PC) + scale_color_brewer("Variables", palette = "Set3") +
+    theme_light(base_size = 10)
+p
 # ggsave(
-#     plot = p, file = "pca-wb-time.png", path = "figures/", device = "png",
+#     plot = p, file = "pca-wb-time.png", path = "paper/figures/", device = "png",
 #     width = 7, height = 4, bg = "white", dpi = 400
 # )
-p + scale_color_brewer(palette = "Set3")
+
 ```
@@
     scale_y_continuous(labels = scales::label_percent()) +
-    theme_light(base_size = 6)
+    theme_light(base_size = 10)
 
@@
     coord_equal() +
-    theme_light(base_size = 6)
+    theme_light(base_size = 10)
 
@@
     coord_equal()+
-    theme_light(base_size = 6)
+    theme_light(base_size = 10)
 
@@
 # ggsave(
-#     filename = "pca_wb-data.png", path = "figures/", dpi = 400, bg = "white",
+#     filename = "pca_wb-data.png", path = "paper/figures/", dpi = 400, bg = "white",
 #     device = "png", width = 7, height = 4,
@@
 For the UN dataset, the area based nature proxies have zero variance (SD), which means that despite completness, it is not informative.
+
+**240220** To do: code categorical variables so they appear on the order of Inequality, life on land, life on water.
+Do the clustering routine on the PCA coordinates, use all components, or the first 10.
+Show the clusters on fig A.
 
@@
     as.matrix(dimnames = list(df_un_reduced |> pull(iso3) |> unique())) |> 
-    FactoMineR::MFA(group = rep(19, 9), type = rep("s", 9)) # 19 years times 9 variables
+    FactoMineR::MFA(group = rep(19, 9), type = rep("s", 9), ncp = 20) # 19 years times 9 variables, 20 components (PCAs)
 ```
@@
 ```{r}
-mfa1$group
+#mfa1$group
 ```
@@
     as.matrix(dimnames = list(df_dat |> pull(country_code) |> unique())) |> 
-    FactoMineR::MFA(group = rep(19, 12), type = rep("s", 12)) # 19 years times 9 variables
+    FactoMineR::MFA(group = rep(19, 12), type = rep("s", 12), ncp = 20) # 19 years times 9 variables
 ```
@@
 ```{r}
-mfa1 |> summary()
-mfa2$summary.quanti
-```
-
-```{r}
-# save(pca1, pca2, mfa1, mfa2, file = "data/ordination_results.Rda")
+# mfa1 |> summary()
+# mfa2$summary.quanti
+```
+
+```{r}
+#save(pca1, pca2, mfa1, mfa2, file = "data/ordination_results.Rda")
+```
+
+```{r}
+df_un_reduced |> 
+    pivot_longer(3:last_col(), names_to = "variables", values_to = "values") |> 
+    separate(variables, sep = "_", into = c("year", "variable")) |> 
+    mutate(year= as.numeric(year)) |> 
+    filter(variable == "gptinc992j") |> 
+    ggplot(aes(year, values)) +
+    geom_line(aes(group = iso3))
+```
+
+```{r}
+
+df_un_reduced |> 
+    pivot_longer(3:last_col(), names_to = "variables", values_to = "values") |> 
+    separate(variables, sep = "_", into = c("year", "variable")) |> 
+    mutate(year= as.numeric(year)) |> #pull(variable) |> unique()
+    mutate(variable = case_when(
+        variable == "Average proportion of Freshwater Key Biodiversity Areas (KBAs) covered by protected areas (%)" ~ "freshwater",
+        variable == "Average proportion of Marine Key Biodiversity Areas (KBAs) covered by protected areas (%)"~"marine",
+        variable == "Average proportion of Mountain Key Biodiversity Areas (KBAs) covered by protected areas (%)" ~ "mountain",
+        variable == "Average proportion of Terrestrial Key Biodiversity Areas (KBAs) covered by protected areas (%)" ~ "terrestrial", 
+        variable == "Forest area certified under an independently verified certification scheme (thousands of hectares)" ~ "forest",
+        variable == "gptinc992j" ~ "Gini",
+        variable == "rptinc992j"~ "Ratio 10/50",
+        variable == "sptinc992j" ~ "Share 1%",
+        .default = variable
+    )) |> 
+    filter(variable %in% c("freshwater", "marine", "terrestrial", "mountain", "forest", "Gini", "Ratio 10/50", "Share 1%")) |> #pull(variable) |> unique()
+    pivot_wider(names_from = variable, values_from = values) |> 
+    ggplot(aes(Gini, freshwater)) + 
+    geom_line(aes(group = iso3, color = year))
+
 ```

### 2024-03-05 paper/manuscript.Rmd
national analysis first steps

@@
 
--   Figure with missing values from WB dataset
+-   Figure with missing values from WB dataset: **improve resolution**
 
--   Figure with missing alues from UN dataset
+-   Figure with missing values from UN dataset: **improve resolution**
 
@@
 We focused subsequent analysis only on country level indicators for SDGs 10, 14 and 15, related to inequality and the biosphere respectively.
-We complemented the SDGs datasets with inequality data from the World Inequality database (WID), using their estimates of the pre-tax national income computed as ratio or Gini.
+We complemented the SDGs datasets with inequality data from the World Inequality database (WID), using their estimates of the ratio of pre-tax national income for working adults (population > 20 years old) computed as the share of the top 10% over the share of the bottom 50% (`rptinc992j_p0p100`), the share of the 1% (`sptinc992j_p99p100`), and the Gini coefficient (`gptinc992j_p0p100`).
 
@@
 We discarded indicators for which time series contained more than 30% of missing values, or less than 45 countries.
-Missing values were then imputed using a cubic spline, leaving us with `r df_dat |> pull(country_code) |> unique() |> length()` countries, `r df_dat |> pull(year) |> unique() |> length()` years of data across 9 indicators for the WB dataset; and `r df_un_reduced |> pull(series_name) |> unique() |> length()` indicators, `r df_un_reduced |> pull(iso3) |> unique() |> length()` countries over `r df_un_reduced |> pull(year) |> unique() |> length()` years for the UN dataset.
+Missing values were then imputed using a cubic spline, leaving us with `r df_dat |> pull(country_code) |> unique() |> length()` countries, `r df_dat |> pull(year) |> unique() |> length()` years of data across 9 indicators for the WB dataset; and `r df_un_reduced |> pivot_longer(cols = 3:last_col(), names_to = "var", values_to= "value") |> separate(var, into = c("year", "series_name"), sep = "_") |>  pull(series_name) |> unique() |> length()` indicators, `r df_un_reduced |> pull(iso3) |> unique() |> length() + 1` countries over `r df_un_reduced |> pivot_longer(cols = 3:last_col(), names_to = "var", values_to= "value") |> separate(var, into = c("year", "series_name"), sep = "_") |> pull(year) |> unique() |> length()` years for the UN dataset.
 The UN dataset was further reduced to 50 countries because the WID does not report inequality timeseries for Fiji.
+Table X summarizes our selected variables, their units and available ranges.
+A list of the countries analysed is presented in the supplementary information (SI).
 
-*Ordination:* We used multiple factor analysis and principal component analysis to reduce the dimensionality of the data and explore similarities and differences across countries.
+*Ordination:* We used multiple factor analysis (MFA) and principal component analysis (PCA) to reduce the dimensionality of the data and explore similarities and differences across countries.
+MFA enables to specify the nested structure of our data and account for repeated observations of our variables over time.
+We recover some of the qualitative results with PCA as robustness check, but these results are presented in the SI.
+We also performed a clustering routine following the protocols by Charrad [] and Brooks [].
+In short, we tested over 10 clustering techniques and compared them across \>30 performance metrics to infer from the data what are the optimal numbers of clusters to fit and preferable algorithms.
+
+*Analysis of trajectories:* With the results from the MFA we extracted trajectories of the countries of analysis on the reduced dimensions space.
+We were particularly interested on the principal components related to inequality and performance on SDGs related to life on land and water.
+The use of trajectories in the MFA space allows us to test whether countries can achieve low inequality and high environmental standards over time, or whether trade offs emerge in improving one dimension at the cost of another.
+
+# Results
+
+```{=tex}
+\begin{figure*}[ht]
+\centering
+\includegraphics[width = 7in, height = 4in]{fig_mfa_wb}
+\caption{\textbf{Multiple factor analysis with World Bank data}  }
+\label{fig:wb}
+\end{figure*}
+```
+
+```{=tex}
+\begin{figure*}[ht]
+\centering
+\includegraphics[width = 7in, height = 4in]{fig_UN_ordination}
+\caption{\textbf{Multiple factor analysis with United Nations data}  }
+\label{fig:un}
+\end{figure*}
+```
 
@@
 -   Paper figures:
-
-    -   Ordination plot for each dataset (nicer arrows) using iso codes instead of numbers.
-        Show clustering groups in colors.
 
@@
 \centering
-\includegraphics[width = 6in, height = 6in]{WB_environment_SDGs}
+\includegraphics[width = 6in, height = 4in]{WB_environment_SDGs}
 \caption{\textbf{World Bank environmental SDGs} }

### 2024-03-05 03-clustering.R

@@ -1,19 +1,30 @@
 library(tidyverse)
+library(NbClust)
+library(clValid)
+library(mclust) # this needs to be loaded due to error on clValid
 
 load("data/ordination_results.Rda")
 
+# WB dataset: pca1, mfa2
+# UN dataset: pca2, mfa1
+
 m <- "ward.D2" # minimize the total within-cluster variance
 
-clust_num <- NbClust::NbClust( 
+## pca1
+
+plot(pca1)
+
+# using first 10 components
+clust_num <- NbClust( 
     data = (pca1$x[,1:10] %>% as.data.frame() %>% 
                 select(where(is.numeric)) ),
     #distance = "maximum",
     min.nc = 2, max.nc = 10, 
     method = m, 
-    index = 'all')
+    index = 'all') # 3 recommended clusters
 
 ## Stability & Internal validation
-stab <- clValid::clValid(
+stab <- clValid(
     obj = pca1$x[,1:10] ,
     nClust = 2:10,
     clMethods = c(
@@ -21,8 +32,11 @@ stab <- clValid::clValid(
         "model", "diana", "sota", "pam", "clara", "agnes"),
     validation = c('stability', "internal"),
     #metric = "manhattan",
-    method = "ward.D2",
+    method = "ward",
     verbose = FALSE
 ) ## Hierarchical is the optimal method
 
-clValid::optimalScores(stab)
+optimalScores(stab)
+summary(stab)
+
+

### 2024-03-05 05-figures.R

@@ -1,0 +1,331 @@
+## Final figures for the paper
+## some chunks come from inequality-SDG note book but quality adusted for publication
+
+library(tidyverse)
+library(ggforce) # for circles
+library(NbClust)
+library(clValid)
+
+load("data/cleaned_sdg.RData") # cleaned datasets
+load("data/ordination_results.Rda")
+
+# WB dataset: pca1, mfa2; data: df_dat
+# UN dataset: pca2, mfa1; data: df_un_reduced
+
+#### multiple factor analysis ####
+## World bank dataset: 
+plot(mfa2)
+
+mfa2$ind |> names()
+mfa2$ind$contrib
+# contry codes on the same order that they entered the MFA
+country_codes <- df_dat |> 
+    select(-country_name, -country, -sptinc992j_p0p100) |> 
+    filter(!is.na(rptinc992j_p0p100)) |> 
+    pivot_longer(cols = `Access to electricity (% of population)`:last_col(),
+                 values_to = "value", names_to = "variable") |> 
+    arrange(variable, year) |> 
+    unite(col = "var", year, variable, sep = "_") |> 
+    pivot_wider(names_from = var, values_from = value) |> 
+    ungroup() |> 
+    select(country_code)
+
+mfa2$global.pca$eig # the first 10 components explain 94.8 of the variability
+
+# clustering:
+# using first 10 components
+clust_num <- NbClust( 
+    data = (mfa2$ind$coord[,1:10] %>% as.data.frame() %>% 
+                select(where(is.numeric)) ),
+    #distance = "maximum",
+    min.nc = 2, max.nc = 10, 
+    method = m, 
+    index = 'all') # 3 recommended clusters
+
+## Stability & Internal validation
+stab <- clValid(
+    obj = mfa2$ind$coord[,1:10] ,
+    nClust = 3,
+    clMethods = c(
+        "hierarchical", "kmeans",  "som", 
+        "model", "diana", "sota", "pam", "clara", "agnes"),
+    validation = c('stability', "internal"),
+    #metric = "manhattan",
+    method = "ward",
+    verbose = FALSE
+) ## Diana is the optimal method, but the Best.partition is better
+
+# clus <- diana(
+#     x = mfa2$ind$coord[,1:10], diss = FALSE,
+#     stop.at.k = 3, metric = 'manhattan',
+#     keep.diss = FALSE, keep.data = FALSE
[…]

### 2024-05-21 inequality-SDGs.Rmd
ms draft

@@
     select(-country_name, -country, -sptinc992j_p0p100) |> 
-    filter(!is.na(rptinc992j_p0p100)) |> 
+    filter(!is.na(rptinc992j_p0p100)) |>#  pull(country_code) |> unique() |> length()
     pivot_longer(cols = `Access to electricity (% of population)`:last_col(),

### 2024-05-21 05-figures.R

@@ -5,6 +5,9 @@ library(tidyverse)
 library(ggforce) # for circles
 library(NbClust)
 library(clValid)
+library(ggnewscale)
+library(gganimate)
+library(patchwork)
 
 load("data/cleaned_sdg.RData") # cleaned datasets
 load("data/ordination_results.Rda")
@@ -33,7 +36,8 @@ country_codes <- df_dat |>
 mfa2$global.pca$eig # the first 10 components explain 94.8 of the variability
 
 # clustering:
-# using first 10 components
+# using first 10 components 
+m <- "ward.D2" 
 clust_num <- NbClust( 
     data = (mfa2$ind$coord[,1:10] %>% as.data.frame() %>% 
                 select(where(is.numeric)) ),
@@ -103,16 +107,16 @@ si_wb1 <- mfa2$group$contrib|>
     labs(y = "Variables", x = "Dimension", tag = "A") +
     theme_light(base_size = 6) +
     theme(legend.position = "bottom") 
-
+si_wb1
 
 ## Note on interpretation (source: Multiple Factor Analysis by example using R, pg 138)
 ## Lg and RV are measurements that help understand relationships between variables.
 ## They are 0 iff each variable of one group is uncorrelated with every variable of the other.
 ## RV is a measurement of the relationship between two groups of multidimensional variables, in 
-## cases when one group is one-dimensiona, Lg is more suitable. In both measurements, we evaluate
+## cases when one group is one-dimensional, Lg is more suitable. In both measurements, we evaluate
 ## the importance of the structure common to two groups of variables:
-## - In RV, wihtout paying attention to the diminsion of this common structure
-## - In Lg, taking into account the dimension of this common structure and its innertia
+## - In RV, without paying attention to the dimension of this common structure
+## - In Lg, taking into account the dimension of this common structure and its inertia
 ## relative to that of the groups. Lg is in some ways a "number of common dimensions", each
 ## weighted by its inertia.
 mfa2$group$cos2 |> 
@@ -144,7 +148,7 @@ b <- mfa2$quanti.var$coord |>
     coord_equal() +
     theme_light(base_size = 7) +
     theme(legend.position = "bottom") 
-
+b
[…]

### 2025-03-21 notes_retreat_2024.Rmd
subnational analysis draft

---
title: "Notes: retreat 2024"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

## Mechanisms & feedbacks session

1. Fairness and perceptions
2. Market concentration
3. Aspiration
    - conspicuous consumtpion
    - social norms
4. Cooperation
5. Environmental change
6. Environmental shocks

**Literature review workflow**

- map research: who, what, etc + key topics
- effects social inequality on environment

Problems:

- Levels, aggregation
- Measures
- direct vs indirect effects
- intertwinness
- positive feedbacks vs negative ones

Policy = P; Inequality = I; Enviornment = E

```{r}
library(DiagrammeR)
mermaid(
    "graph TB
    P --> I
    I --> E
    "
)

mermaid(
    "graph TB
    P <--> I
    I <--> E
    P <--> E
    "
)
```

Environmental justice: colonialims etc

Potential way of coding:

### 2025-03-21 paper/manuscript.Rmd
subnational analysis draft

@@
 # abstract
+
+# Notes:
+
+# how to make it different from the trilemma?# 
+# clearer story
+# how much of the patterns hold constant if we remove indicators?
+# sensitivity analysis for MFA / ordination and clustering.
 ```
@@
 
+-   Why inequality and the biosphere / as opposed to other SDGs
+
 -   Do the trade-offs hold at national scale with SDG data?
@@
 -   Clustering
+-   Trajectories
 
@@
 
--   download and check world inequality db again
+-   We confirm the results of Wu with three clusters of countries in two different SDGs datastets.
+-   We confirm the existence of inequality traps or regime shifts in inequality.
+-   We confirm the path dependency: the pathway from one regime to another is not random and offer insights on the type of policies to favour to achieve multiple goals.
+It also points to risks that need to be manage to avoid trade-offs.
 
 Discussion {1000w}
+
+-   Poor data quality overall.
+    Biodiversity metrics either don't change or are focused on conservation.
+    Thus it speaks of intentions not actual state variables.
+
+-   Despite data challenges we confirm the existence of inequality regimes.
+
+-   Some of the mechanisms by which inequality and the biosphere might be coupled occur at the scale of individuals or households.
+
+-   Future research gaps: do the patterns hold at subnational scales?
+    Do the patterns hold with biodiversity variables that actually reflects the state of biodiversity?
+    Would the patterns be the same at local scales?
+    Would they be different in human dominated systems such as cities?
 
@@
 
--
+\pagebreak
+
+\begin{abstract}
+
+\end{abstract}
 
@@
 It is unclear if there are structural constraints that limits a country capacity to achieve particular SDGs, or if there are milestones that need to be unlocked before it can achieve the desired performance in a particular dimension of sustainability.
-In other words, these correlations do not account for the path dependency inherent of the process of development.
+In other words, these correlations do not account for the path dependency inherent on the process of development.
+
+-   Add a paragraph on lit review on SDGs synergies and trade-offs
 
@@
 
-Recent work shows that a trilemma exist where countries struggle to achieve high prosperity, high environmental standards while reducing inequality.
+Recent work shows that a trilemma exist where countries struggle to achieve high prosperity, high environmental standards while reducing inequality [@Wu_2024].
 Using data from environmental footprint, the gross national product, and the Gini coefficient time series, 53 countries were clustered and a typology of trajectories identified.
@@
 Countries trajectories on the trilemma space, however, suggest the hypothesis that social progress can be achieved without compromising the biosphere.
+
 It remains an open questions if these patterns hold across different datasets and scales.
-Here we address such gap by studying synergies and trade-offs between SDGs related to inequality and the biosphere.
+If the hypothesis of the inequality trap is true, we should observe bimodal or multimodal distributions in inequality and environmental variables.
+Each mode would correspond to an inequality regime, and the transitions probability of staying within one regime should be much higher than the probability of shifting regimes.
+Here we address such questions by studying synergies and trade-offs between SDGs related to inequality and the biosphere.
+
+-   Add a figure that describes the rationale of the methods and hypotheses
 
@@
 We focused subsequent analysis only on country level indicators for SDGs 10, 14 and 15, related to inequality and the biosphere respectively.
-We complemented the SDGs datasets with inequality data from the World Inequality database (WID), using their estimates of the ratio of pre-tax national income for working adults (population > 20 years old) computed as the share of the top 10% over the share of the bottom 50% (`rptinc992j_p0p100`), the share of the 1% (`sptinc992j_p99p100`), and the Gini coefficient (`gptinc992j_p0p100`).
+We complemented the SDGs datasets with inequality data from the World Inequality database (WID), using their estimates of the ratio of pre-tax national income for working adults (population \> 20 years old) computed as the share of the top 10% over the share of the bottom 50% (`rptinc992j_p0p100`), the share of the 1% (`sptinc992j_p99p100`), and the Gini coefficient (`gptinc992j_p0p100`).
 
@@
 We discarded indicators for which time series contained more than 30% of missing values, or less than 45 countries.
-Missing values were then imputed using a cubic spline, leaving us with `r df_dat |> pull(country_code) |> unique() |> length()` countries, `r df_dat |> pull(year) |> unique() |> length()` years of data across 9 indicators for the WB dataset; and `r df_un_reduced |> pivot_longer(cols = 3:last_col(), names_to = "var", values_to= "value") |> separate(var, into = c("year", "series_name"), sep = "_") |>  pull(series_name) |> unique() |> length()` indicators, `r df_un_reduced |> pull(iso3) |> unique() |> length() + 1` countries over `r df_un_reduced |> pivot_longer(cols = 3:last_col(), names_to = "var", values_to= "value") |> separate(var, into = c("year", "series_name"), sep = "_") |> pull(year) |> unique() |> length()` years for the UN dataset.
+Missing values were then imputed using a cubic spline, leaving us with `r df_dat |> filter(!is.na(rptinc992j_p0p100)) |> pull(country_code) |> unique() |> length()` countries, `r df_dat |> pull(year) |> unique() |> length()` years of data across 9 indicators for the WB dataset; and `r df_un_reduced |> pivot_longer(cols = 3:last_col(), names_to = "var", values_to= "value") |> separate(var, into = c("year", "series_name"), sep = "_") |>  pull(series_name) |> unique() |> length()` indicators, `r df_un_reduced |> pull(iso3) |> unique() |> length() + 1` countries over `r df_un_reduced |> pivot_longer(cols = 3:last_col(), names_to = "var", values_to= "value") |> separate(var, into = c("year", "series_name"), sep = "_") |> pull(year) |> unique() |> length()` years for the UN dataset.
 The UN dataset was further reduced to 50 countries because the WID does not report inequality timeseries for Fiji.
@@
 We also performed a clustering routine following the protocols by Charrad [] and Brooks [].
-In short, we tested over 10 clustering techniques and compared them across \>30 performance metrics to infer from the data what are the optimal numbers of clusters to fit and preferable algorithms.
+We tested over 10 clustering techniques and compared them across \>30 performance metrics to infer from the data what are the optimal numbers of clusters to fit and preferable algorithms.
 
-*Analysis of trajectories:* With the results from the MFA we extracted trajectories of the countries of analysis on the reduced dimensions space.
-We were particularly interested on the principal components related to inequality and performance on SDGs related to life on land and water.
-The use of trajectories in the MFA space allows us to test whether countries can achieve low inequality and high environmental standards over time, or whether trade offs emerge in improving one dimension at the cost of another.
+*Analysis of trajectories:* With the results from the MFA we identified candidate variables where synergies or trade-offs are expected.
+A trade-off in the reduced dimensional space occurs when improving on the direction of one indicator (e.g. reducing inequality) implies declining on the direction of another indicator.
+Similarly, a synergy would be when increasing in one indicator seems to pull a country on the direction of improving in another indicator.
+We extracted trajectories of the countries of analysis for some of these candidate variables.
+If the hypothesis of inequality traps or inequality regimes is true, we would expect to find multimodal distributions in inequality indicators with respect to other variables; we expect the modes of these distributions to be correlated to the countries groups identified via clustering analysis; and we expect that most countries stay within their region of the parameter space and perhaps a few of them able to move between regimes.
 
@@
 
-```{=tex}
+Despite the differences in coverage with respect to countries, time, and indicators tracked, the ordination in both data sets results in three clusters of countries (Fig \ref{fig:wb}, Fig \ref{fig:un}).
+For the WB data set, countries along the first principal component are differentiated by high levels of economic inequality, high energy intensity but low carbon emissions (negative values of Dim 1, green cluster), versus countries with relatively low gender inequality, high agricultural productivity, high carbon emissions, high urbanization and internet access (positive values in Dim 1, yellow and blue clusters).
+Forest related variables have the lowest loading on the first two components and do not change over time, while access to internet or inclusion of women in leadership roles have the highest variability over time.
+
 \begin{figure*}[ht]
@@
 \includegraphics[width = 7in, height = 4in]{fig_mfa_wb}
-\caption{\textbf{Multiple factor analysis with World Bank data}  }
+\caption{\textbf{Multiple factor analysis with World Bank data} The first 10 principal componets explain 94.8\% of the varation, the first two (A) explain 53\%.
+The 10 first components were used to cluster 151 countries resulting in three clusters (A).
+The correlation circle across explanatory variables is presented in (B) along their loadings on the first two components of the ordination.}
 \label{fig:wb}
 \end{figure*}
-```
 
-```{=tex}
+The UN dataset offers a similar ordination, where countries with low values along the first axes but higher values along the second axis have the highest inequalities (e.g Mexico, Fig \ref{fig:un}).
+The inequality variables are highly correlated but also explain large amounts of the variance.
+Contrary to the WB data, here forest related variables do show variability over time, but variables related to biodiversity loss (Red list index) do not change much over the time period of the data (2000-2015).
+Interestingly, the ordination suggests that places with low inequality would have high Red list index.
+Japan, Italy or Ireland are examples of countries on the region of the parameter space where inequality is relatively low but conservation metrics per country are relatively high.
+
 \begin{figure*}[ht]
@@
 \end{figure*}
-```
+
+We find evidence of multi-modal distributions in inequality and environmental variables.
+This finding supports the hypothesis of the existence of inequality regimes, particularly when inequality is measured as the share of the top 1% (Fig \ref{fig:bim}).
+However, the lack of variability in environmental variables and smaller sample size in the UN data set prevent us to statistically derive modes in the distribution, except for the red list index (Fig \ref{fig:SM_UNbim}).
+
+\begin{figure*}[ht]
+\centering
+\includegraphics[width = 7in, height = 4in]{bimodal}
+\caption{\textbf{Inequality regimes} We find bimodal distributions for some dimensions of inequality against environmental factors in the WB dataset.
+Countries trajectories in A using the same cluster groups as in Fig \ref{fig:un}.
+Bimodal distributions are more common for the share of the top 1\% than other inequality variables (B).
+The y-axis has been rescaled to the range 0-1 to ease comparison.
+Supplementary figure \ref{fig:SM_UNbim} shows a similar plot for the UN data, however lack of variability in the environmental data prevents the identification of multiple modes in the distribution.}
+\label{fig:bim}
+\end{figure*}
+
+# Discussion
+
+-   Biodiversity variables are either limited to conservation, or resource production.
+    Other biodiversity related values such as cultural values (intrinsic, recreational, spiritual) are not captured by SDG indicators.
+
+-   Data gaps undermine the analysis of trajectories over time.
+    For some of the variables where information exist, the statistics do not change (e.g. % forest area).
+    Thus the indicators fail at capturing progress or lack of it towards the sustainable goals.
+
+-   We find evidence for bistability in inequality, further supporting the hypothesis of inequality traps.
+    However, data gaps and lack of variability in biodiversity indicators prevent us to further test features of regime shift dynamics such as hysteresis.
+    Further work can look into this hypothesis by looking into other datasets or longer time series.
+
+# Conclusion
 
@@
 
-    -   Plot with PCA selection (2-3 components?) and then change over time of the components.
-        Do clustering on trayectories as well?
-        -- and then add a map with the clusters for each dataset.
+    -   Plot with PCA selection (2-3 components?) and then change over time of the components. Do clustering on trayectories as well? -- and then add a map with the clusters for each dataset.
 
@@
 
-```{=tex}
 \renewcommand\thefigure{S\arabic{figure}}
@@
 \setcounter{figure}{0}
-```
-```{=tex}
 \begin{figure*}[ht]
@@
 \end{figure*}
-```
-```{=tex}
 \begin{figure*}[ht]
@@
 \includegraphics[width = 6in, height = 5in]{UN_SDG_10-14-15}
-\caption{\textbf{United Nations SDGs 10, 14 and 25 dataset} }
+\caption{\textbf{United Nations SDGs 10, 14 and 15 dataset} }
 \label{fig:SM_UNdata}
 \end{figure*}
-```
-```{=tex}
 \begin{figure*}[ht]
@@
 \end{figure*}
-```
+\begin{figure*}[ht]
+\centering
+\includegraphics[width = 6in, height = 5in]{sm_wb_ordinations}
+\caption{\textbf{Variable importance in ordination on WB data} }
+\label{fig:SM_WBord}
+\end{figure*}
+\begin{figure*}[ht]
+\centering
+\includegraphics[width = 4in, height = 3.5in]{bimodal_UN}
+\caption{\textbf{Bimodality in UN data} The quality of the data does not allow for detection of bimodality in most UN variables.
+The lack of variation in biosphere related data does not allow the identification of areas in the parameter space where trajectories converge, except for the red list index, where we find multimodal distributions }
+\label{fig:SM_UNbim}
+\end{figure*}

### 2025-03-21 subnational.Rmd
subnational analysis draft

---
title: "Subnational inequality"
author: "Juan Rocha"
date: "Updated `r Sys.Date()`"
output:
  html_notebook:
    toc: yes
    toc_float: yes
    highlight: tango
    code_folding: hide
    df_print: paged
    theme: 
      bootswatch: cosmo
      code_font:
        google: Fira Code
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  cache = TRUE, warning = FALSE, message = FALSE, echo = TRUE,
  fig.width = 8, fig.height = 5
)

library(tidyverse)
library(naniar)
library(plotly)
library(patchwork)
library(fs)
library(tictoc)

# time series
library(tsibble)
library(fable)
library(imputeTS)
library(slider)

## spatial
library(terra)
library(tidyterra)
library(ggspatial)
library(stars)

# # clustering
# library(vegan)
# library(NbClust)
# library(clValid)
# library(mclust)

#theme_set(theme_light(base_size = 12))
```

## Inequality

Exploring subnational inequality data from Kummu et al. Load data:

```{r}

fls <- dir_ls("~/Documents/Projects/DATA/Kummu_inequality/transfer_294976_files_744c217c/")
dat <- rast(fls[2])
dat

dat <- vect(fls[5])
#terra::plot(dat$gini_disp_2021)

tmap::tm_polygons(dat)

ggplot(data = dat) +
    geom_spatvector(aes(fill = `2020`), color = "white", linewidth = 0.1) +
    scale_fill_viridis_c()
   # scale_fill_gradient2() 
   # 
dat |> as_tibble() |> 
    ggplot(aes(`1990`, `2021`)) +
    geom_point(aes(color = slope)) +
    geom_abline(slope = 1, intercept = 0) +
    scale_color_gradient2()
```

Time series

```{r}
dat |> 
    as_tibble() |> 
    pivot_longer(cols = `1990`:`2021`, names_to = "year", values_to = "Gini") |> 
    ggplot(aes(year, Gini)) +
    geom_line(aes(group = iso3), alpha = 0.5)
```

## Human footprint

```{r}
hfp <- dir_ls("/Users/juanrocha/Documents/Projects/DATA/HII_human_impact_index/first-gen/")

fp1 <- stars::read_stars(hfp[10])
fp1 <- rast(hfp[20])

terra::plot(fp1, type = "continuous", #col = map.pal("viridis", 100),
            range = c(-100,5000),
            col = c("#224f1a","#a3ff76","#feff6f","#a09568","#ffa802","#f7797c","#fb0102","#d87136","#a90086","#7a1ca5","#421137","#000000")
            )
#click(n=1)

# tic()
# m <- max(fp1, na.rm = TRUE)
# toc() # 496s

m

fp1[,10000,] |> range(na.rm= TRUE)
```

From Google Earth engine 
```{javascript eval=FALSE}
var hiiviz = {min: 5, max: 5000, palette: ["224f1a","a3ff76","feff6f","a09568","ffa802","f7797c","fb0102","d87136","a90086","7a1ca5","421137","000000"]
```

## Spatial joints

```{r}
#dat <- dat |> st_as_sf()
#plgns <- dat |> select(admID, iso3, geometry)

tic()
fp <- extract(x = fp1, y = dat, fun = mean, bind = TRUE, na.rm = TRUE)
toc() #
```

### 2025-05-06 paper/manuscript.Rmd
implementing comments paper, updating analysis SDG un

@@
 
-Countries around the world have committed to achieve 17 sustainable development goals.
-The ambitious agenda is materialized in 169 targets and indicators, yet not all targets are monitored and not all countries report them.
+Countries around the world have committed to achieve 17 sustainable development goals as proposed by the United Nations.
+The ambitious agenda is materialized in 169 targets and indicators, yet not all targets are monitored or properly measured, and not all countries report them, presenting significant data and knowledge gaps.
 An open question in sustainability science is if these targets are achievable simultaneously?
 Previous work on synergies and trade-offs between sustainable development goals (SDGs) suggest that both options are possible, yet this work is limited to expert elicitation and correlational studies.
-It is unclear if there are structural constraints that limits a country capacity to achieve particular SDGs, or if there are milestones that need to be unlocked before it can achieve the desired performance in a particular dimension of sustainability.
-In other words, these correlations do not account for the path dependency inherent on the process of development.
+It is unclear if there are system or structural constraints that limit a country's capacity to achieve particular SDGs, or if there are milestones that need to be unlocked before it can achieve the desired performance in a particular dimension of sustainability.
+In other words, these correlations do not account for the path dependency and causal inference inherent on the process of development.
 
--   Add a paragraph on lit review on SDGs synergies and trade-offs
+When the SDGs were adopted in 2015, work began to create a comprehensive database of SDG indicators [@schmidt-traub2017; @sachs2024], to be updated annually.
+Based on these data, previous analyses found that goals associated with poverty alleviation, well-being, economic development, and innovation (e.g.
+SDGs 1, 3, 7, 8, 9) tended to be synergistic with other goals, while goals related to responsible consumption, climate action, natural resources, and cooperation showed the most trade-offs (e.g.
+SDGs 11, 12, 13, 14, 16, 17) [@pradhan2017; @s41599-019-0335-5].
+Drawing upon relevant World Bank data, Lusseau and Mancini [@s41893-019-0231-4] find these patterns to be modulated by the overall income level of countries, with low-income countries showing synergies across all goals, while trade-offs start to appear in higher income categories.
+More recently, [@s41467-023-44679-w] analysed transboundary SDG interactions and found that high income countries play a disproportionate role in influencing the achievement of SDGs in other countries.
 
 Here we address the question of synergies and trade-offs by focusing on goals related to biosphere and inequality (SDGs 10, 14 and 15).
-The motivation is twofold.
+The motivation is threefold.
 First, the Convention for Biological Diversity is currently negotiating and agreeing on the next set of goals and ambitions to bend the curve of biodiversity loss.
 Understanding which countries are well positioned to achieve biodiversity goals or not will be necessary to design strategies that support countries with low capabilities, that address the dimensions of justice, and design targets with plausibility in mind.
-Second, within country inequality has been rising in the last decades even in high income countries, and as consequence of the Covid pandemic, in between country inequality raised for first time in a generation.
+Second, within-country inequality has been rising in the last decades even in high income countries, and as a consequence of the Covid pandemic, in between country inequality raised for first time in a generation.
 These statistics set back progress on the inequality SDG by at least a decade, and undermines the mantra of leaving no one behind.
+And third, recent conceptual and theoretical work has proposed mechanisms by which increase in inequality can impact the environment, and changes in the environment can feedback into further inequalities [@hamann2018; @dade2022; @leach2018].
 
-Recent work shows that a trilemma exist where countries struggle to achieve high prosperity, high environmental standards while reducing inequality [@Wu_2024].
+Recent work also shows that a trilemma exists where countries struggle to achieve high prosperity, high environmental standards while reducing inequality [@Wu_2024].
 Using data from environmental footprint, the gross national product, and the Gini coefficient time series, 53 countries were clustered and a typology of trajectories identified.
-No country achieved the three goals simultenously, and Latin American countries seem to have dynamics characterized by an inequality trap.
-Countries trajectories on the trilemma space, however, suggest the hypothesis that social progress can be achieved without compromising the biosphere.
+No country achieved the three goals simultaneously, and Latin American countries seem to have dynamics characterized by an inequality trap.
+Countries trajectories on the trilemma space, however, suggest the hypothesis that social progress can be achieved without compromising the biosphere [@Wu_2024].
 
-It remains an open questions if these patterns hold across different datasets and scales.
+It remains an open question if these patterns hold across different datasets and scales.
 If the hypothesis of the inequality trap is true, we should observe bimodal or multimodal distributions in inequality and environmental variables.
@@
 un_countries <- read_csv2(file = "~/Documents/Projects/DATA/SDGs_UNStats/UNSD — Methodology.csv")
-sdg <- read_csv(file = "~/Documents/Projects/DATA/SDGs_UNStats/20180817102044132_juan.rocha@su.se_data.csv") |> janitor::clean_names()
 
-sdg <- sdg |> 
-    left_join(
-        select(un_countries, country = `Country or Area`, starts_with("ISO")),
-        by = c("geo_area_name" = "country")) |> 
-    filter(value != "-")|> 
-    mutate(value = as.numeric(value))
+# old UN dataset
+# sdg <- read_csv(file = "~/Documents/Projects/DATA/SDGs_UNStats/20180817102044132_juan.rocha@su.se_data.csv") |> janitor::clean_names()
+
+## new UN dataset:
+fls <- dir_ls("data/sdgs/")
+
+sdg <- list()
+tic()
+for (i in seq_along(fls)){
+    load(fls[i])
+    sdg[[i]] <- d
+}
+toc() # 21s
+
+tic()
+sdg <- sdg |> bind_rows() |> 
+    unnest(c(goal, target, indicator))
+toc() # 12.1s
+
+# sdg <- sdg |> 
+#     left_join(
+#         select(un_countries, country = `Country or Area`, starts_with("ISO")),
+#         by = c("geo_area_name" = "country")) |> 
+#     filter(value != "-")|> 
+#     mutate(value = as.numeric(value))
 
@@
 The WB dataset offers `r dat$indicator_name |> unique() |> length()` indicators, with time series from 1990 to 2019 for `r dat$country_name |> unique() |> length()` countries or administrative areas (N = 2'013.791 observations).
-The UN dataset offers time series from 2000 to 2019 for 17 SDGs, 112 targets, and 356 indicators, in 310 administrative areas (N = 1'072.731 observations).
+The UN dataset offers time series from 1963 to 2025 for 17 SDGs, 168 targets, and 247 indicators, 687 time series, in 413 administrative areas (N = `r sdg |> nrow()` observations).
 Despite their coverage, both datasets contain high proportion of missing values and the sample is unbalanced across countries and time, meaning some countries have better temporal coverage than others.
-We focused subsequent analysis only on country level indicators for SDGs 10, 14 and 15, related to inequality and the biosphere respectively.
+We focused subsequent analysis only on country level indicators for SDGs 5, 6, 10, 13, 14 and 15, related to inequality and the biosphere respectively.
 We complemented the SDGs datasets with inequality data from the World Inequality database (WID), using their estimates of the ratio of pre-tax national income for working adults (population \> 20 years old) computed as the share of the top 10% over the share of the bottom 50% (`rptinc992j_p0p100`), the share of the 1% (`sptinc992j_p99p100`), and the Gini coefficient (`gptinc992j_p0p100`).
+We also used their estimates for net wealth inequality computed as the share of the top 10% over the share of the bottom 50% (`rhweal992j_p0p100`), the share of the 1% (`shweal992j_p99p100`), and the Gini coefficient(`ghweal992j_p0p100`).
 
 ```{r echo = FALSE}
+rm(dat, sdg, fls) # remove unnecessary objects
+load(file = paste0(here(), "/data/cleaned_SDGs_2025.Rda")) 
+# dat is the most recent UN dataset 2025, keys is the var names
 load(file = paste0(here(), "/data/cleaned_sdg.RData"))
@@
 We discarded indicators for which time series contained more than 30% of missing values, or less than 45 countries.
-Missing values were then imputed using a cubic spline, leaving us with `r df_dat |> filter(!is.na(rptinc992j_p0p100)) |> pull(country_code) |> unique() |> length()` countries, `r df_dat |> pull(year) |> unique() |> length()` years of data across 9 indicators for the WB dataset; and `r df_un_reduced |> pivot_longer(cols = 3:last_col(), names_to = "var", values_to= "value") |> separate(var, into = c("year", "series_name"), sep = "_") |>  pull(series_name) |> unique() |> length()` indicators, `r df_un_reduced |> pull(iso3) |> unique() |> length() + 1` countries over `r df_un_reduced |> pivot_longer(cols = 3:last_col(), names_to = "var", values_to= "value") |> separate(var, into = c("year", "series_name"), sep = "_") |> pull(year) |> unique() |> length()` years for the UN dataset.
+Missing values were then imputed using a cubic spline, leaving us with `r df_dat |> filter(!is.na(rptinc992j_p0p100)) |> pull(country_code) |> unique() |> length()` countries, `r df_dat |> pull(year) |> unique() |> length()` years of data across 9 indicators for the WB dataset; and `r keys |> filter(series %in% unique(dat$series)) |>  pull(series) |> unique() |> length()` series capturing `r keys |> filter(series %in% unique(dat$series)) |>  pull(indicator) |> unique() |> length()` indicators, `r dat |> pull(geoAreaName) |> unique() |> length()` countries over `r dat |> pull(timePeriodStart) |> unique() |> length()` years for the UN dataset.
 The UN dataset was further reduced to 50 countries because the WID does not report inequality timeseries for Fiji.
@@
 A list of the countries analysed is presented in the supplementary information (SI).
+
+```{r echo=FALSE}
+# create table: source, series code, name, 
+```
+
+VOY AQUI
 
@@
 If the hypothesis of inequality traps or inequality regimes is true, we would expect to find multimodal distributions in inequality indicators with respect to other variables; we expect the modes of these distributions to be correlated to the countries groups identified via clustering analysis; and we expect that most countries stay within their region of the parameter space and perhaps a few of them able to move between regimes.
+
+```{r table1, include=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
+
+library(kableExtra)
+
+un_vars <- df_un_reduced |> names() |> str_subset("^\\d{4}_") |> str_remove("\\d{4}_") |> unique()
+un_vars <- un_vars[-c(7:9)]
+wb_vars <- names(df_dat)[4:12]
+iq_vars <-  names(df_dat)[c(14,16,17)]
+
+tbl <- tibble(
+    Source = c(rep("UN",length(un_vars)), rep("WB", length(wb_vars)), rep("WII",length(iq_vars))),
+    Variable = c(un_vars, wb_vars, iq_vars)
+    #Units = c(rep("Percentage", 4), "Thousands of hectares", "0-1")
+)
+
+tbl <- tbl |> mutate(
+    Units = str_extract(Variable, "\\(.\\)$"),
+    Units = str_remove_all(Units, "\\(|\\)")
+) |> mutate(
+        Variable = str_remove(Variable, " \\(%\\)"),
+        Units = str_extract(Variable, "\\(.*\\)$"),
+        Variable = str_remove(Variable, " \\(.*\\)$"),
+        Units = str_remove_all(Units, "\\(|\\)"))
+
+tbl$Units[1:4] <- "Percentage"
+tbl$Units[6] <- "0 to 1"
+tbl$Units[16:18] <- "0 to 1"
+tbl$Variable[16:18] <- c(
+    "Ratio of pre-tax national income for working adults (rptinc992j_p0p100)", "Gini coefficient (gptinc992j_p0p100)", "Share of the top 1% (sptinc992j_p99p100)"
+)
+
+knitr::kable(tbl, format = "latex", booktabs = TRUE, caption = "Summary of variables used") %>%
+  kable_styling(latex_options = c("striped", "hold_position"), 
+                full_width = TRUE, font_size = 8) %>%
+  column_spec(1, width = "15mm") |> 
+  column_spec(2, width = "10cm") 
+```
 

### 2025-05-06 wii-db.Rmd
implementing comments paper, updating analysis SDG un

@@
     unique() |> 
-    str_subset("(r|g|s)ptinc")
+    str_subset("(r|g|s)ptinc|(r|g|s)hweal")
 
@@
     bind_rows()
-toc()
+toc() # 90s
 

### 2025-05-06 SDG_explore.R

@@ -1,0 +1,59 @@
+#### exploration ####
+library(tidyverse)
+library(skimr)
+library(fs)
+
+
+fls <- dir_ls("data/sdgs/")
+fls
+
+load(fls[10])
+
+extract_summary <- function(x){
+    
+    load(x)
+    
+    x <- d |> 
+        unnest(c(goal, target, indicator)) |> 
+        select(goal:series, seriesDescription, source, geoAreaName:value) |> 
+        #pull(source) |> head()
+        mutate(timePeriodStart = as.numeric(timePeriodStart), value = as.numeric(value)) |>
+        group_by(goal, target, indicator, series, geoAreaName, timePeriodStart) |>
+        summarize(value = mean(value)) |> 
+        ungroup()
+    x |> group_by(goal, target, indicator, series) |> 
+        summarize(years = max(timePeriodStart) - min(timePeriodStart),
+                  countries = length(unique(geoAreaName)),
+                  n = n(), 
+                  completness = n/(years*countries)) |> 
+        ungroup()
+    
+}
+safe_extract <- safely(extract_summary)
+
+out <- map(
+    .x = fls,
+    .f = safe_extract,
+    .progress = TRUE
+)
+
+out <- transpose(out)
+is_ok <- map(out$error, is.null) |> unlist()
+
+sdgs <- out$result |> 
+    bind_rows() |> 
+    unique()
+
+series <- jsonlite::fromJSON("https://unstats.un.org/sdgs/UNSDGAPIV5/v1/sdg/Series/List") |> 
+    as_tibble()
+
+sdgs <- sdgs |> 
+    rename(seriesCode = series) |> 
+    left_join(
+        series |> unnest(c(goal, target, indicator)) |> 
+            select(goal, target, indicator, seriesCode = code, description) |> 
+            unique()
+        )
+
+write_csv(sdgs, file = "data/sdgs_completness.csv")
+

### 2025-05-06 ordination.R

@@ -1,0 +1,380 @@
+library(tidyverse)
+library(fs)
+library(tictoc)
+library(gganimate)
+library(patchwork)
+#### Import and clean data ####
+## list of countries
+un_countries <- read_csv2(file = "~/Documents/Projects/DATA/SDGs_UNStats/UNSD — Methodology.csv") |> 
+    janitor::clean_names()
+countries <- read_csv(
+    file = "~/Documents/Projects/DATA/WorldBank/SDG_csv/SDGCountry.csv") |> 
+    janitor::clean_names()
+
+countries_list <- countries |> filter(!is.na(currency_unit)) |> pull(country_code)
+## correct Namibia, it is not an NA in x2_alpha_code
+#countries$x2_alpha_code[countries$country_code == "NAM"] <- "NA"
+
+countries <- un_countries |># names()
+    select(country_or_area:iso_alpha3_code) |> 
+    filter(iso_alpha3_code %in% countries_list)
+
+
+## load files
+fls <- dir_ls("data/sdgs/")
+# 
+# sdg <- map(
+#     .x = fls[1],
+#     .f = function(x) load(x),
+#     .progress = TRUE
+# )
+
+dat <- list()
+tic()
+for (i in seq_along(fls)){
+    load(fls[i])
+    dat[[i]] <- d
+}
+toc() # 21s
+
+fls[which(map(dat, nrow) |> unlist() >= 50000)] # perhaps re-download with larger page size
+
+tic()
+dat <- dat |> bind_rows() |> 
+    unnest(c(goal, target, indicator))
+toc() # 12.1s
+
+dat # 2.8M obs
+
+keys <- dat |> 
+    select(goal, target, indicator, series, seriesDescription) |> 
+    unique()
+
+df_countries <- dat |> 
+    select(starts_with("geoArea")) |> 
+    unique()
+
+df_countries <- df_countries |> 
+    mutate(l = str_length(geoAreaCode)) |> 
+    mutate(m49_code = case_when(
+        l == 1 ~ paste0("00", geoAreaCode),
+        l == 2 ~ paste0("0", geoAreaCode),
+        .default = geoAreaCode)
+    ) |> select(-l)
+
+un_countries <- un_countries |> 
+    select(country_or_area, m49_code, contains("iso"))
+
+### Reduce to countries only
+df_countries <- un_countries |> 
+    filter(iso_alpha3_code %in% countries_list) |> 
+    left_join(df_countries)
+
[…]

### 2025-05-11 paper/manuscript.Rmd
paper updates: new UN data

@@
 
-```{r data, include=FALSE}
+```{r setup, include=FALSE}
 library(tidyverse)
@@
 
--   Add a figure that describes the rationale of the methods and hypotheses
-
 # Methods
@@
 ```{r echo = FALSE, message=FALSE, warning=FALSE}
-dat <- read_csv(file = "~/Documents/Projects/DATA/WorldBank/SDG_csv/SDGData.csv") |> 
-    janitor::clean_names()
-
-un_countries <- read_csv2(file = "~/Documents/Projects/DATA/SDGs_UNStats/UNSD — Methodology.csv")
+# dat <- read_csv(file = "~/Documents/Projects/DATA/WorldBank/SDG_csv/SDGData.csv") |> 
+#     janitor::clean_names()
+# 
+# un_countries <- read_csv2(file = "~/Documents/Projects/DATA/SDGs_UNStats/UNSD — Methodology.csv")
 
@@
 ## new UN dataset:
-fls <- dir_ls("data/sdgs/")
-
-sdg <- list()
-tic()
-for (i in seq_along(fls)){
-    load(fls[i])
-    sdg[[i]] <- d
-}
-toc() # 21s
-
-tic()
-sdg <- sdg |> bind_rows() |> 
-    unnest(c(goal, target, indicator))
-toc() # 12.1s
-
+# fls <- fs::dir_ls(paste0(here(), "/data/sdgs/"))
+# 
+# sdg <- list()
+# 
+# for (i in seq_along(fls)){
+#     load(fls[i])
+#     sdg[[i]] <- d
+# }
+# 
+# 
+# sdg <- sdg |> bind_rows() |> 
+#     unnest(c(goal, target, indicator))
+# 
 # sdg <- sdg |> 
@@
 *Datasets:* To that end, we used the SDGs datasets made openly available by the World Bank (WB) and United Nations (UN).
-The WB dataset offers `r dat$indicator_name |> unique() |> length()` indicators, with time series from 1990 to 2019 for `r dat$country_name |> unique() |> length()` countries or administrative areas (N = 2'013.791 observations).
-The UN dataset offers time series from 1963 to 2025 for 17 SDGs, 168 targets, and 247 indicators, 687 time series, in 413 administrative areas (N = `r sdg |> nrow()` observations).
+The WB dataset offers 403 indicators, with time series from 1990 to 2019 for 263 countries or administrative areas (N = 2'013.791 observations).
+The UN dataset offers time series from 1963 to 2025 for 17 SDGs, 168 targets, and 247 indicators, 687 time series, in 413 administrative areas (N = 2'821.669 observations).
 Despite their coverage, both datasets contain high proportion of missing values and the sample is unbalanced across countries and time, meaning some countries have better temporal coverage than others.
@@
 ```{r echo = FALSE}
-rm(dat, sdg, fls) # remove unnecessary objects
+#rm(dat, sdg, fls) # remove unnecessary objects
 load(file = paste0(here(), "/data/cleaned_SDGs_2025.Rda")) 
-# dat is the most recent UN dataset 2025, keys is the var names
-load(file = paste0(here(), "/data/cleaned_sdg.RData"))
-# `df_dat` is the data from World Bank, and `df_un_reduced` is the data from UN SDGs
-df_dat <- df_dat |> ungroup()
-df_un_reduced <- df_un_reduced |> ungroup()
+un_dat <- ungroup(un_dat)
+wb_dat <- ungroup(wb_dat)
+
 ```
@@
 We discarded indicators for which time series contained more than 30% of missing values, or less than 45 countries.
-Missing values were then imputed using a cubic spline, leaving us with `r df_dat |> filter(!is.na(rptinc992j_p0p100)) |> pull(country_code) |> unique() |> length()` countries, `r df_dat |> pull(year) |> unique() |> length()` years of data across 9 indicators for the WB dataset; and `r keys |> filter(series %in% unique(dat$series)) |>  pull(series) |> unique() |> length()` series capturing `r keys |> filter(series %in% unique(dat$series)) |>  pull(indicator) |> unique() |> length()` indicators, `r dat |> pull(geoAreaName) |> unique() |> length()` countries over `r dat |> pull(timePeriodStart) |> unique() |> length()` years for the UN dataset.
-The UN dataset was further reduced to 50 countries because the WID does not report inequality timeseries for Fiji.
-Table X summarizes our selected variables, their units and available ranges.
+Missing values were then imputed using a cubic spline, leaving us with `r wb_dat |> pull(country_code) |> unique() |> length()` countries, `r wb_dat |> pull(year) |> unique() |> length()` years of data across 9 indicators for the WB dataset; and `r un_keys |> filter(series %in% unique(un_dat$series)) |>  pull(series) |> unique() |> length()` series capturing `r un_keys |> filter(series %in% unique(un_dat$series)) |>  pull(indicator) |> unique() |> length()` indicators, `r un_dat |> pull(geoAreaName) |> unique() |> length()` countries over `r un_dat |> pull(timePeriodStart) |> unique() |> length()` years for the UN dataset.
+The UN dataset was further reduced to 67 countries because the WID does not report inequality time series for Fiji.
+Table \ref{tbl1} summarizes our selected variables, their units and available ranges.
 A list of the countries analysed is presented in the supplementary information (SI).
-
-```{r echo=FALSE}
-# create table: source, series code, name, 
-```
-
-VOY AQUI
 
@@
 We recover some of the qualitative results with PCA as robustness check, but these results are presented in the SI.
-We also performed a clustering routine following the protocols by Charrad [] and Brooks [].
+We also performed a clustering routine following the protocols by Charrad [@charrad2014] and Brooks [@brock2008].
 We tested over 10 clustering techniques and compared them across \>30 performance metrics to infer from the data what are the optimal numbers of clusters to fit and preferable algorithms.
@@
 
-un_vars <- df_un_reduced |> names() |> str_subset("^\\d{4}_") |> str_remove("\\d{4}_") |> unique()
-un_vars <- un_vars[-c(7:9)]
-wb_vars <- names(df_dat)[4:12]
-iq_vars <-  names(df_dat)[c(14,16,17)]
+load(paste0(here(),"/data/wb_var_defs.Rda"))
+load(paste0(here(),"/data/wii_small-data.RData"))
 
-tbl <- tibble(
-    Source = c(rep("UN",length(un_vars)), rep("WB", length(wb_vars)), rep("WII",length(iq_vars))),
-    Variable = c(un_vars, wb_vars, iq_vars)
-    #Units = c(rep("Percentage", 4), "Thousands of hectares", "0-1")
+un_vars <- un_keys |> 
+    filter(series %in% unique(un_dat$series)) |> 
+    select(Goal = goal, Series = series, Variable = seriesDescription) |> 
+    mutate(Source = "UN")
+
+wb_vars <- wb_key |> 
+    filter(indicator_name %in% colnames(wb_dat)[4:12]) |> 
+    select(Series = series_code, Variable = indicator_name) |> 
+    mutate(Goal = c("15", "10", rep("13", 2), "10", rep("15",2),"10","10" ),
+           Source = "WB")
+
+iq_vars <- tibble(
+    Source = "WII", 
+    Goal = "10", 
+    Series = ineq_dat |> 
+        unite(col = "series", variable, percentile, sep = "_") |>
+        # removed due to zero variance
+        filter(!series %in% c("sptinc992j_p0p100","shweal992j_p0p100")) |>
+        pull(series) |> unique(),
+    Variable = c(
+        "Ratio of pre-tax national income for working adults",
+        "Ratio of wealth for working adults",
+        "Share of the top 1% of wealth",
+        "Gini coefficient of wealth",
+        "Gini coefficeint of pre-tax income for working adults",
+        "Share of the 1% of pre-tax income for working adults"
     )
+)
+
+tbl <- bind_rows(wb_vars, un_vars, iq_vars) |> 
+    select(Source, Goal, Series, Variable) |> 
+    mutate(Goal = as.numeric(Goal)) |> 
+    arrange(Goal)
 
@@
 
-tbl$Units[1:4] <- "Percentage"
+tbl$Units[3:4] <- "0 to 1"
+tbl$Units[5] <- "US$ per cubic meter"
 tbl$Units[6] <- "0 to 1"
-tbl$Units[16:18] <- "0 to 1"
-tbl$Variable[16:18] <- c(
-    "Ratio of pre-tax national income for working adults (rptinc992j_p0p100)", "Gini coefficient (gptinc992j_p0p100)", "Share of the top 1% (sptinc992j_p99p100)"
-)
+tbl$Units[18:23] <- "0 to 1"
+tbl$Units[29:32] <- "0 to 1"
+tbl$Units[c(7,9,13)] <- "sq. km"
 
-knitr::kable(tbl, format = "latex", booktabs = TRUE, caption = "Summary of variables used") %>%
+knitr::kable(
+    tbl, format = "latex", booktabs = TRUE, caption = "Summary of variables used", label = "tbl1") %>%
   kable_styling(latex_options = c("striped", "hold_position"), 
-                full_width = TRUE, font_size = 8) %>%
-  column_spec(1, width = "15mm") |> 
-  column_spec(2, width = "10cm") 
+                full_width = TRUE, font_size = 8) |> 
+  column_spec(c(1,2), width = "5mm") |> column_spec(3, width = "25mm") |>
+  column_spec(4, width = "8cm") 
 ```
@@
 
-Despite the differences in coverage with respect to countries, time, and indicators tracked, the ordination in both data sets results in three clusters of countries (Fig \ref{fig:wb}, Fig \ref{fig:un}).
-For the WB data set, countries along the first principal component are differentiated by high levels of economic inequality, high energy intensity but low carbon emissions (negative values of Dim 1, green cluster), versus countries with relatively low gender inequality, high agricultural productivity, high carbon emissions, high urbanization and internet access (positive values in Dim 1, yellow and blue clusters).
-Forest related variables have the lowest loading on the first two components and do not change over time, while access to internet or inclusion of women in leadership roles have the highest variability over time.
+Despite the differences in coverage with respect to countries, time, and indicators tracked, the ordination in both data sets results in two clusters of countries (Fig \ref{fig:wb}, Fig \ref{fig:un}).
+For the WB data set, countries along the first principal component are differentiated by high levels of economic inequality, high energy intensity but low carbon emissions (positive values of Dim 1, eg.
+green cluster Mexico: MEX or South Africa: ZAF), versus countries with relatively low gender inequality, high agricultural productivity, high carbon emissions, high urbanization and internet access (negative values in Dim 1, yellow cluster).
+Forest related variables have the lowest loading on the first two components and do not change much over time, while access to internet or inclusion of women in leadership roles have the highest variability over time.
+The first dimension on the ordination is best explained by variability in the inequality variables including women in buisness, access to internet and electricity, while the second dimension is best explained by urbanization and carbon emissions (Fig \ref{fig:SM_WBord}).
 
@@
 \caption{\textbf{Multiple factor analysis with World Bank data} The first 10 principal componets explain 94.8\% of the varation, the first two (A) explain 53\%.
-The 10 first components were used to cluster 151 countries resulting in three clusters (A).
-The correlation circle across explanatory variables is presented in (B) along their loadings on the first two components of the ordination.}
+The 10 first components were used to cluster 151 countries resulting in two clusters (A).
+The correlation circle across explanatory variables is presented in (B) along their loadings on the first two components of the ordination.
+Variables in the legend are ordered and coloured according to the SDGs used (e.g.
+orange for gender equality, blues for water and sanitation, reds for inequality, and greens for life on land).}
 \label{fig:wb}
@@
 
-The UN dataset offers a similar ordination, where countries with low values along the first axes but higher values along the second axis have the highest inequalities (e.g Mexico, Fig \ref{fig:un}).
+The UN dataset offers a similar ordination, where countries with high values along the first axes and higher values along the second axis have the highest inequalities (e.g Fig \ref{fig:un}).
 The inequality variables are highly correlated but also explain large amounts of the variance.
-Contrary to the WB data, here forest related variables do show variability over time, but variables related to biodiversity loss (Red list index) do not change much over the time period of the data (2000-2015).
-Interestingly, the ordination suggests that places with low inequality would have high Red list index.
-Japan, Italy or Ireland are examples of countries on the region of the parameter space where inequality is relatively low but conservation metrics per country are relatively high.
+Contrary to the WB data, here forest related variables do show variability over time, but variables related to biodiversity loss (Red list index) or some of the area based indicators for water related SDGs do not change much over the time period of the data (2000-2021).
+Places with lower economic inequality also tend to have better opportunities for woman to participate in political decision making.
+Interestingly, high values on the red list index correlate with lower levels of inequality.
+The first dimension on the ordination is best explained by variability in the inequality, while the second dimension is best explained by key biodiversity areas in terrestrial systems, reservoir statistics, as well as women participation in parliament (Fig \ref{fig:SM_UNord}).
 
@@
 \centering
-\includegraphics[width = 7in, height = 4in]{fig_UN_ordination}
-\caption{\textbf{Multiple factor analysis with United Nations data}  }
+\includegraphics[width = 7in, height = 4.5in]{fig_UN_ordination}
+\caption{\textbf{Multiple factor analysis with United Nations data} The first 10 principal components used in the ordination explain 72\% of the variation, the first two (A) explain 29.9\%.
+The first 10 components were used to cluster 67 countires resulting in two clusters (A).
+The correlation circle across explanatory variables is presented in (B) along their loadings on teh first two components of the ordination.
+Variables in the legend are ordered and coloured according to the SDGs used (e.g.
+reds are inequality, greens are life in land).}
 \label{fig:un}
@@
 We find evidence of multi-modal distributions in inequality and environmental variables.
-This finding supports the hypothesis of the existence of inequality regimes, particularly when inequality is measured as the share of the top 1% (Fig \ref{fig:bim}).
-However, the lack of variability in environmental variables and smaller sample size in the UN data set prevent us to statistically derive modes in the distribution, except for the red list index (Fig \ref{fig:SM_UNbim}).
+This finding supports the hypothesis of the existence of inequality regimes both in income and wealth, particularly when inequality is measured as the share of the top 1% (Fig \ref{fig:bim}).
+However, the lack of variability in environmental variables and smaller sample size in the UN data set prevent us to statistically derive modes in the distribution or all our variables (Table \\ref{tab}).
+As a result, we only report bimodal distributions for cereal yields, carbon emissions, and energy intensity (WB data, Fig \ref{fig:bim}}), and the red list index (UN data, Fig \ref{fig:SM_UNbim}).
+We also confirm that, as expected, these inequality regimes are related with the country typologies identified.
 
@@
 
--   Biodiversity variables are either limited to conservation, or resource production.
-    Other biodiversity related values such as cultural values (intrinsic, recreational, spiritual) are not captured by SDG indicators.
+Most countries in the world are committed to achieve progress on sustainable development goals by 2030.
+Despite several of these aspirational goals wont be meet by 2030; they concentrate political willingness and action that has been translated into some progress in most countries.
+One key question is whether there are synergies and trade-offs between these goals.
+Here we addressed this question by investigating SDGs related to inequality and the biosphere (SDGs 5, 6, 10, 13, 14, and 15) through ordination methods.
 
--   Data gaps undermine the analysis of trajectories over time.
-    For some of the variables where information exist, the statistics do not change (e.g. % forest area).
-    Thus the indicators fail at capturing progress or lack of it towards the sustainable goals.
+We find evidence of synergies and trade-offs.
+For example, almost all inequality metrics are correlated, except the ratio of wealth which is much more noisy and sometimes anti-correlated with inequality metrics on income.
+For example, Sweden traditionally has had low income inequality but high wealth inequality, although the former has been increasing as well.
+Similarly, gender equity measured as share of women in parliament is generally better in places with lower income inequality but not necessarily low wealth inequality.
+Interestingly, countries with high gender equality perform better in area based indicators of ecological and biological conservation.
+Higher levels of urbanization are correlated with improvements on access to electricity and internet usage, but also with higher carbon emissions and lower energy intensity.
 
--   We find evidence for bistability in inequality, further supporting the hypothesis of inequality traps.
+We had an explicit interest in the interactions between inequality and biosphere related SDGs.
+Biodiversity variables are either limited to conservation, or resource production.
+Most of the proxies of ecosystem variables are area based (e.g. % forest area), meaning they change very little over time.
+The lack of variability in the SDGs datasets and a large proportion of missing values (Figs \ref{fig:SM_WB_env}, \ref{fig:SM_WB_env2}, \ref{fig:SM_UNdata} ) forced us to drop many indicators.
+It also questions whether the indicators currently used are good proxies of progress or lack of in the goals.
+If a variable does not change over time at a scale at which information can feedback political decisions, then is not very useful as proxy of progress a desired direction.
+Other biodiversity related values such as cultural values (intrinsic, recreational, spiritual) are not currently captured by SDG indicators.
+An interesting avenue for future research is to design comparable observables that can be monitored by third parties (not reported by governments) that are sensitive enough to capture progress or lack of it towards the goals.
+Some examples include the essential biodiversity variables innitiative advanced by GeoBON, the human footprint index, or recent developments on functional integrity.
+All of them offer time series at varied levels of resolution that enable subnational monitoring.
+
+Despite these limitations, we find evidence for bistability in inequality, further supporting the hypothesis of inequality traps [@Wu_2024].
 However, data gaps and lack of variability in biodiversity indicators prevent us to further test features of regime shift dynamics such as hysteresis.
-    Further work can look into this hypothesis by looking into other datasets or longer time series.
+To further test the existence of hysteresis we need a better understanding of the potential mechanisms at the country scale that might be generating the alternative regimes, including a controlling parameter at which the shift from one regime to the other should result on a different break point.
+Further work can look into this hypothesis by looking into other datasets or longer time series, or by exploring the plausibility of alternative mechanisms through modelling.
+Generative models for wealth inequality exist inspired in biological mechanisms of heredity, but less is known on inequality mechanisms for income or gender dimensions.
 
@@
 \setcounter{figure}{0}
+
+List of countries analyzed in the UN dataset after removing time series with too many missing values or too few countries: `r un_dat |> pull(geoAreaName) |> unique() |> str_flatten_comma(last = " and ")`
+
+List of countries analyzed in the WB dataset after removing time series with too many missing values or too few countries: `r wb_dat |> pull(country_name) |> unique() |> str_flatten_comma(last = " and ")`
+
 \begin{figure*}[ht]
@@
 \end{figure*}
+
+\begin{figure*}[ht]
+\centering
+\includegraphics[width = 6in, height = 5in]{sm_un_ordinations}
+\caption{\textbf{Variable importance in ordination on UN data} }
+\label{fig:SM_UNord}
+\end{figure*}
+
 \begin{figure*}[ht]

### 2025-05-11 02-data_cleaning.R

@@ -1,0 +1,332 @@
+library(tidyverse)
+library(fs)
+library(tictoc)
+library(gganimate)
+library(patchwork)
+
+#### UN SDGs ####
+##list of countries
+un_countries <- read_csv2(file = "~/Documents/Projects/DATA/SDGs_UNStats/UNSD — Methodology.csv") |> 
+    janitor::clean_names()
+countries <- read_csv(
+    file = "~/Documents/Projects/DATA/WorldBank/SDG_csv/SDGCountry.csv") |> 
+    janitor::clean_names()
+
+countries_list <- countries |> filter(!is.na(currency_unit)) |> pull(country_code)
+## correct Namibia, it is not an NA in x2_alpha_code
+#countries$x2_alpha_code[countries$country_code == "NAM"] <- "NA"
+
+countries <- un_countries |># names()
+    select(country_or_area:iso_alpha3_code) |> 
+    filter(iso_alpha3_code %in% countries_list)
+
+
+## load files
+fls <- dir_ls("data/sdgs/")
+# 
+# sdg <- map(
+#     .x = fls[1],
+#     .f = function(x) load(x),
+#     .progress = TRUE
+# )
+
+dat <- list()
+tic()
+for (i in seq_along(fls)){
+    load(fls[i])
+    dat[[i]] <- d
+}
+toc() # 21s
+
+fls[which(map(dat, nrow) |> unlist() >= 50000)] # perhaps re-download with larger page size
+
+tic()
+dat <- dat |> bind_rows() |> 
+    unnest(c(goal, target, indicator))
+toc() # 12.1s
+
+dat # 2.8M obs
+
+un_keys <- dat |> 
+    select(goal, target, indicator, series, seriesDescription) |> 
+    unique()
+
+df_countries <- dat |> 
+    select(starts_with("geoArea")) |> 
+    unique()
+
+df_countries <- df_countries |> 
+    mutate(l = str_length(geoAreaCode)) |> 
+    mutate(m49_code = case_when(
+        l == 1 ~ paste0("00", geoAreaCode),
+        l == 2 ~ paste0("0", geoAreaCode),
+        .default = geoAreaCode)
+    ) |> select(-l)
+
+un_countries <- un_countries |> 
+    select(country_or_area, m49_code, contains("iso"))
+
+### Reduce to countries only
+df_countries <- un_countries |> 
+    filter(iso_alpha3_code %in% countries_list) |> 
+    left_join(df_countries)
+
+## Reduce dataset to SDGs of interest. 
[…]

### 2025-05-11 05-figures.R

@@ -5,51 +5,47 @@ library(tidyverse)
 library(ggforce) # for circles
 library(NbClust)
 library(clValid)
+library(mclust)
 library(ggnewscale)
 library(gganimate)
 library(patchwork)
+library(FactoMineR)
+library(RColorBrewer)
 
-load("data/cleaned_sdg.RData") # cleaned datasets
+load("data/cleaned_SDGs_2025.Rda") # cleaned datasets
 load("data/ordination_results.Rda")
-
-# WB dataset: pca1, mfa2; data: df_dat
-# UN dataset: pca2, mfa1; data: df_un_reduced
+load("data/wii_small-data.RData")
+countries <- read_csv(
+    file = "~/Documents/Projects/DATA/WorldBank/SDG_csv/SDGCountry.csv") |> 
+    janitor::clean_names()
 
 #### multiple factor analysis ####
 ## World bank dataset: 
-plot(mfa2)
+plot(wb_mfa)
 
-mfa2$ind |> names()
-mfa2$ind$contrib
+wb_mfa$ind |> names()
+wb_mfa$ind$contrib
 # contry codes on the same order that they entered the MFA
-country_codes <- df_dat |> 
-    select(-country_name, -country, -sptinc992j_p0p100) |> 
-    filter(!is.na(rptinc992j_p0p100)) |> 
-    pivot_longer(cols = `Access to electricity (% of population)`:last_col(),
-                 values_to = "value", names_to = "variable") |> 
-    arrange(variable, year) |> 
-    unite(col = "var", year, variable, sep = "_") |> 
-    pivot_wider(names_from = var, values_from = value) |> 
-    ungroup() |> 
-    select(country_code)
 
-mfa2$global.pca$eig # the first 10 components explain 94.8 of the variability
+country_codes <- wb_dat |> filter(!is.na(rhweal992j_p0p100), !is.na(rptinc992j_p0p100)) |> pull(country_code) |> unique() #149 countries
+
+wb_mfa$global.pca$eig 
 
 # clustering:
 # using first 10 components 
 m <- "ward.D2" 
-clust_num <- NbClust( 
-    data = (mfa2$ind$coord[,1:10] %>% as.data.frame() %>% 
+clust_wb <- NbClust( 
+    data = (wb_mfa$ind$coord[,1:10] %>% as.data.frame() %>% 
                 select(where(is.numeric)) ),
     #distance = "maximum",
     min.nc = 2, max.nc = 10, 
     method = m, 
[…]

### 2025-05-11 ordination.R

@@ -3,255 +3,25 @@ library(fs)
 library(tictoc)
 library(gganimate)
 library(patchwork)
-#### Import and clean data ####
-## list of countries
-un_countries <- read_csv2(file = "~/Documents/Projects/DATA/SDGs_UNStats/UNSD — Methodology.csv") |> 
-    janitor::clean_names()
+
+
+load("data/wii_small-data.RData")
+load("data/cleaned_SDGs_2025.Rda")
+
 countries <- read_csv(
     file = "~/Documents/Projects/DATA/WorldBank/SDG_csv/SDGCountry.csv") |> 
     janitor::clean_names()
 
-countries_list <- countries |> filter(!is.na(currency_unit)) |> pull(country_code)
-## correct Namibia, it is not an NA in x2_alpha_code
-#countries$x2_alpha_code[countries$country_code == "NAM"] <- "NA"
 
-countries <- un_countries |># names()
-    select(country_or_area:iso_alpha3_code) |> 
-    filter(iso_alpha3_code %in% countries_list)
-
-
-## load files
-fls <- dir_ls("data/sdgs/")
-# 
-# sdg <- map(
-#     .x = fls[1],
-#     .f = function(x) load(x),
-#     .progress = TRUE
-# )
-
-dat <- list()
-tic()
-for (i in seq_along(fls)){
-    load(fls[i])
-    dat[[i]] <- d
-}
-toc() # 21s
-
-fls[which(map(dat, nrow) |> unlist() >= 50000)] # perhaps re-download with larger page size
-
-tic()
-dat <- dat |> bind_rows() |> 
-    unnest(c(goal, target, indicator))
-toc() # 12.1s
-
-dat # 2.8M obs
-
-keys <- dat |> 
-    select(goal, target, indicator, series, seriesDescription) |> 
-    unique()
-
-df_countries <- dat |> 
-    select(starts_with("geoArea")) |> 
-    unique()
-
-df_countries <- df_countries |> 
-    mutate(l = str_length(geoAreaCode)) |> 
-    mutate(m49_code = case_when(
-        l == 1 ~ paste0("00", geoAreaCode),
-        l == 2 ~ paste0("0", geoAreaCode),
-        .default = geoAreaCode)
-    ) |> select(-l)
-
-un_countries <- un_countries |> 
-    select(country_or_area, m49_code, contains("iso"))
-
-### Reduce to countries only
-df_countries <- un_countries |> 
[…]

### 2025-05-19 notes_retreat_2024.Rmd
implementing comments paper, updating analysis SDG un
updates from retreat

@@
 ---
+
+## BYS retreat 2025
+
+- Session National SDGs:
+    - Discussion comparsion to other studies, analysis from Matti and Chinese papers
+    - List factors or mechanisms that can create the regimes
+    - Hypothesize and test causes that distinguishes clusters: wealth, income, corruption
+
+- Futures discussion
+    - Mistra: case studies in Indonesia with Robert, mining with Exeter,
+    - Writers: Maike, Emilie, Tom, Caro, Peter, Juan, Tong(?)
+
+## Notes meeting 2025
+
+- We re-visit the selection of variables and downloaded all SDGs data from the UN API
+    - script `sdg_api.R`
+- We ran again into issues of missing values, took us long to solve it but we did some progress
+- Decision: to include SDGs `c("5", "6", "10", "13","14", "15")` so add water and climate
 

### 2025-05-19 paper/manuscript.Rmd
updates from retreat

@@
 ---
-title: "Trade-offs between biosphere goals and inequality at the national scale"
+title: "Biosphere Stewardship in an Unequal World: A Sustainable Development Goals Approach"
 output:
@@
 Countries around the world have committed to achieve 17 sustainable development goals as proposed by the United Nations.
-The ambitious agenda is materialized in 169 targets and indicators, yet not all targets are monitored or properly measured, and not all countries report them, presenting significant data and knowledge gaps.
+The ambitious agenda is materialized in 169 targets and indicators, yet not all targets are monitored or properly measured, and not all countries report them, presenting significant data and knowledge gaps [@min2024].
 An open question in sustainability science is if these targets are achievable simultaneously?
-Previous work on synergies and trade-offs between sustainable development goals (SDGs) suggest that both options are possible, yet this work is limited to expert elicitation and correlational studies.
+Previous work on synergies and trade-offs between sustainable development goals (SDGs) suggest that both options are possible [@schmidt-traub2017; @sachs2024; @pradhan2017; @kroll2019], yet this work is limited to expert elicitation and correlational studies.
 It is unclear if there are system or structural constraints that limit a country's capacity to achieve particular SDGs, or if there are milestones that need to be unlocked before it can achieve the desired performance in a particular dimension of sustainability.
@@
 
-When the SDGs were adopted in 2015, work began to create a comprehensive database of SDG indicators [@schmidt-traub2017; @sachs2024], to be updated annually.
+When the SDGs were adopted in 2015, the United Nations began to create a comprehensive database of SDG indicators [@schmidt-traub2017; @sachs2024], to be updated annually.
 Based on these data, previous analyses found that goals associated with poverty alleviation, well-being, economic development, and innovation (e.g.
 SDGs 1, 3, 7, 8, 9) tended to be synergistic with other goals, while goals related to responsible consumption, climate action, natural resources, and cooperation showed the most trade-offs (e.g.
-SDGs 11, 12, 13, 14, 16, 17) [@pradhan2017; @s41599-019-0335-5].
-Drawing upon relevant World Bank data, Lusseau and Mancini [@s41893-019-0231-4] find these patterns to be modulated by the overall income level of countries, with low-income countries showing synergies across all goals, while trade-offs start to appear in higher income categories.
-More recently, [@s41467-023-44679-w] analysed transboundary SDG interactions and found that high income countries play a disproportionate role in influencing the achievement of SDGs in other countries.
+SDGs 11, 12, 13, 14, 16, 17) [@pradhan2017; @kroll2019].
+Drawing upon relevant World Bank data, Lusseau and Mancini [@lusseau2019] find these patterns to be modulated by the overall income level of countries, with low-income countries showing synergies across all goals, while trade-offs start to appear in higher income categories.
+More recently, [@xiao2024] analysed transboundary SDG interactions and found that high income countries play a disproportionate role in influencing the achievement of SDGs in other countries.
+These differences suggest that it may be worthwhile investigating if the nature of the mechanisms linking SDG to each other could be dependent on income level.
 
-Here we address the question of synergies and trade-offs by focusing on goals related to biosphere and inequality (SDGs 10, 14 and 15).
+Here we address the question of synergies and trade-offs by focusing on goals related to biosphere and inequality (SDGs 5, 6, 10, 13, 14 and 15).
 The motivation is threefold.
 First, the Convention for Biological Diversity is currently negotiating and agreeing on the next set of goals and ambitions to bend the curve of biodiversity loss.
-Understanding which countries are well positioned to achieve biodiversity goals or not will be necessary to design strategies that support countries with low capabilities, that address the dimensions of justice, and design targets with plausibility in mind.
-Second, within-country inequality has been rising in the last decades even in high income countries, and as a consequence of the Covid pandemic, in between country inequality raised for first time in a generation.
+Understanding which countries are well positioned to achieve biodiversity goals or not will be necessary to design strategies that support countries with low capabilities, that address the dimensions of justice, and design targets with plausibility in mind [@díaz2020].
+Second, within-country inequality has been rising in the last decades even in high income countries [@milanovic2024], and as a consequence of the Covid pandemic, in between country inequality raised for first time in a generation [@wade2023world].
 These statistics set back progress on the inequality SDG by at least a decade, and undermines the mantra of leaving no one behind.
-And third, recent conceptual and theoretical work has proposed mechanisms by which increase in inequality can impact the environment, and changes in the environment can feedback into further inequalities [@hamann2018; @dade2022; @leach2018].
+And third, recent conceptual and theoretical work has proposed mechanisms by which increase in inequality can impact the environment, while changes in the environment can feedback into further inequalities [@hamann2018; @dade2022; @leach2018].
 
-Recent work also shows that a trilemma exists where countries struggle to achieve high prosperity, high environmental standards while reducing inequality [@Wu_2024].
+Recent work also shows that a trilemma exists where countries struggle to simultaneously achieve high prosperity, high environmental standards while reducing inequality [@Wu_2024].
 Using data from environmental footprint, the gross national product, and the Gini coefficient time series, 53 countries were clustered and a typology of trajectories identified.
 No country achieved the three goals simultaneously, and Latin American countries seem to have dynamics characterized by an inequality trap.
-Countries trajectories on the trilemma space, however, suggest the hypothesis that social progress can be achieved without compromising the biosphere [@Wu_2024].
+Some countries trajectories on the trilemma space, however, suggest the hypothesis that social progress can be achieved without compromising the biosphere [@Wu_2024].
+While no country has simultaneously achieved these three goals, some countrie are indeed moving on the right direction [@Wu_2024; @chrisendo2024; @liu2025]
 
-It remains an open question if these patterns hold across different datasets and scales.
+It remains open questions whether these patterns are robust across different datasets and scales, or whether there are specific driving factors and feedback mechanisms that underlie inequality traps.
+The dichotomy between low-income countries exhibiting synergies across all goals, while trade-offs starting to appear in higher-income countries [@lusseau2019] motivates the need to study mechanisms explaining these trade-offs and how they may differ due to countries’ level of income.
 If the hypothesis of the inequality trap is true, we should observe bimodal or multimodal distributions in inequality and environmental variables.
 Each mode would correspond to an inequality regime, and the transitions probability of staying within one regime should be much higher than the probability of shifting regimes.
+If there are nonlinear dynamics in inequality keeping countries trapped on a particular regime, then we should also observe hysteresis or different break points between regimes.
 Here we address such questions by studying synergies and trade-offs between SDGs related to inequality and the biosphere.
@@
 The UN dataset offers time series from 1963 to 2025 for 17 SDGs, 168 targets, and 247 indicators, 687 time series, in 413 administrative areas (N = 2'821.669 observations).
-Despite their coverage, both datasets contain high proportion of missing values and the sample is unbalanced across countries and time, meaning some countries have better temporal coverage than others.
+Despite their coverage, both datasets contain a high proportion of missing values and the sample is unbalanced across countries and time, meaning some countries have better temporal coverage than others.
 We focused subsequent analysis only on country level indicators for SDGs 5, 6, 10, 13, 14 and 15, related to inequality and the biosphere respectively.
@@
 *Ordination:* We used multiple factor analysis (MFA) and principal component analysis (PCA) to reduce the dimensionality of the data and explore similarities and differences across countries.
-MFA enables to specify the nested structure of our data and account for repeated observations of our variables over time.
+MFA enables us to specify the nested structure of our data and account for repeated observations of our variables over time.
 We recover some of the qualitative results with PCA as robustness check, but these results are presented in the SI.
-We also performed a clustering routine following the protocols by Charrad [@charrad2014] and Brooks [@brock2008].
+We also performed a clustering sensitivity analysis following the protocols by Charrad [@charrad2014] and Brooks [@brock2008].
 We tested over 10 clustering techniques and compared them across \>30 performance metrics to infer from the data what are the optimal numbers of clusters to fit and preferable algorithms.
@@
 *Analysis of trajectories:* With the results from the MFA we identified candidate variables where synergies or trade-offs are expected.
-A trade-off in the reduced dimensional space occurs when improving on the direction of one indicator (e.g. reducing inequality) implies declining on the direction of another indicator.
-Similarly, a synergy would be when increasing in one indicator seems to pull a country on the direction of improving in another indicator.
+A trade-off in the reduced dimensional space occurs when improving on the direction of one indicator (e.g. reducing inequality) implies a decline in the direction of another indicator.
+Similarly, a synergy would be when increasing in one indicator coincides with a country towards improving in another indicator.
 We extracted trajectories of the countries of analysis for some of these candidate variables.
@@
 
-```{r table1, include=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
+```{r table1, include=TRUE, echo=FALSE, warning=FALSE, message=FALSE, opts.label="tbl1"}
 
@@
 This finding supports the hypothesis of the existence of inequality regimes both in income and wealth, particularly when inequality is measured as the share of the top 1% (Fig \ref{fig:bim}).
-However, the lack of variability in environmental variables and smaller sample size in the UN data set prevent us to statistically derive modes in the distribution or all our variables (Table \\ref{tab}).
+However, the lack of variability in environmental variables and smaller sample size in the UN data set prevent us to statistically derive modes in the distribution or all our variables (Table \ref{tbl1}).
 As a result, we only report bimodal distributions for cereal yields, carbon emissions, and energy intensity (WB data, Fig \ref{fig:bim}}), and the red list index (UN data, Fig \ref{fig:SM_UNbim}).
@@
 The lack of variability in the SDGs datasets and a large proportion of missing values (Figs \ref{fig:SM_WB_env}, \ref{fig:SM_WB_env2}, \ref{fig:SM_UNdata} ) forced us to drop many indicators.
-It also questions whether the indicators currently used are good proxies of progress or lack of in the goals.
-If a variable does not change over time at a scale at which information can feedback political decisions, then is not very useful as proxy of progress a desired direction.
+It also questions whether the indicators currently used are good proxies of progress, or lack of it, in the goals.
+If a variable does not change over time at a scale at which information can feedback political decisions, then is not very useful as proxy of progress towards a desired direction.
 Other biodiversity related values such as cultural values (intrinsic, recreational, spiritual) are not currently captured by SDG indicators.
 An interesting avenue for future research is to design comparable observables that can be monitored by third parties (not reported by governments) that are sensitive enough to capture progress or lack of it towards the goals.
-Some examples include the essential biodiversity variables innitiative advanced by GeoBON, the human footprint index, or recent developments on functional integrity.
-All of them offer time series at varied levels of resolution that enable subnational monitoring.
+Some examples include the essential biodiversity variables initiative advanced by GeoBON, the human footprint index, or recent developments on functional integrity.
+All of them offer time series at varied levels of resolution that enable sub national monitoring.
 
@@
 Further work can look into this hypothesis by looking into other datasets or longer time series, or by exploring the plausibility of alternative mechanisms through modelling.
-Generative models for wealth inequality exist inspired in biological mechanisms of heredity, but less is known on inequality mechanisms for income or gender dimensions.
+Generative models for wealth inequality exist inspired in biological mechanisms of heredity [@arrow2009], but less is known on inequality mechanisms for income or gender dimensions, although some hypothesis have been put forward [@chancel2020unsustainable].
 
@@
 \end{figure*}
-
 \begin{figure*}[ht]

### 2025-05-19 05-figures.R

@@ -167,7 +167,7 @@ b <- wb_mfa$quanti.var$coord |>
     mutate(var_name = str_replace(var_name, "j_p", "j-p")) |> 
     separate(var_name, into = c("year", "var_name"), sep = "_" ) |> #slice(172:200) 
     mutate(var_name = str_replace(var_name, "j-p", "j_p")) |> 
-    left_join(wb_short_names) |> 
+    left_join(wb_short_names) |> filter(!str_detect(var_name, "tinc|hwea")) |> 
     mutate(year = as.numeric(year)) |># pull(var_name) |> unique()
     ggplot() +
     geom_hline(yintercept = 0, linetype = 2, color = "grey50", linewidth = 0.5) +
@@ -176,9 +176,9 @@ b <- wb_mfa$quanti.var$coord |>
         aes(xend = Dim.1, x = 0, yend = Dim.2, y = 0, color = short_name, alpha = year), 
         arrow = arrow(length = unit(0.02, "npc"))) +
     geom_circle(aes(x0=0, y0=0, r = 1),  color = "grey24", linewidth = 0.01 ) +
-    scale_color_manual("Variables from the World Bank and World Inequality databases", 
-                       values = clrs,
-                       guide = guide_legend(ncol = 4, title.position = "top")) +
+    # scale_color_manual("Variables from the World Bank and World Inequality databases", 
+    #                    values = clrs,
+    #                    guide = guide_legend(ncol = 4, title.position = "top")) +
     scale_alpha("Year", guide = guide_legend(ncol = 1, title.position = "top")) +
     labs(x = glue::glue("Dim 1 (", round(wb_mfa$global.pca$eig[1,2], 2), "% of variance explained)" ),
          y = glue::glue("Dim 2 (", round(wb_mfa$global.pca$eig[2,2], 2), "% of variance explained)" ), tag = "B") +
@@ -359,7 +359,7 @@ bb <- un_mfa$quanti.var$coord |>
     mutate(var_name = str_replace(var_name, "_", "-")) |> #change the first one
     separate(var_name, into = c("year", "var_name"), sep = "-" ) |> #slice(172:200)
     mutate(year = as.numeric(year)) |> 
-    left_join(un_short_names) |> 
+    left_join(un_short_names) |> #filter(str_detect(var_name, "tinc|hwea")) |> 
     ggplot() +
     geom_hline(yintercept = 0, linetype = 2, color = "grey50", linewidth = 0.5) +
[…]

### 2025-05-19 corruption.R

@@ -1,0 +1,142 @@
+library(tidyverse)
+
+## Corruption datasets from Gothemburg
+
+## Standard dataset: 12.391 obs, 2008 variables
+crpt <- read_csv("https://www.qogdata.pol.gu.se/data/qog_std_ts_jan25.csv")
+crpt |> names()
+
+## Basic dataset: 12.391 obs, 251 most commonly used variables
+crpt <- read_csv("https://www.qogdata.pol.gu.se/data/qog_bas_ts_jan25.csv")
+crpt
+
+crpt |> select(year, ccodealp, ti_cpi) |> # corruption perception index
+    ggplot(aes(year, ti_cpi)) +
+    geom_path(aes(group = ccodealp))
+
+crpt_dat <- crpt |> 
+    select(year, ccodealp, ti_cpi) |> # corruption perception index
+    group_by(ccodealp) |> 
+    filter(!all(is.na(ti_cpi))) |> # remove full missing ts
+    summarize(
+        mean_crpt = mean(ti_cpi, na.rm = TRUE),
+        trend_crpt = slope(ti_cpi)
+    )
+
+wb <- read_csv(
+    "~/Documents/Projects/DATA/WorldBank/SDG_csv/SDGData.csv") %>% 
+    janitor::clean_names() |> 
+    filter(indicator_name == "GNI per capita (US$)") |> select(-x35) |> 
+    pivot_longer(cols = starts_with("x"), names_to = "year", values_to = "gni") |> 
+    mutate(year = str_remove(year, "x"), year = as.numeric(year)) |> 
+    select(-indicator_code, -indicator_name) |> 
+    group_by(country_code) |> 
+    filter(!all(is.na(gni))) |> 
+    summarize(
+        mean_gni = mean(gni, na.rm = TRUE), 
+        trend_gni = slope(gni)
+    )
+    
+
+
+un_dat
+
+slope <- function(x){
+    fit <- lm(x~ seq_along(x))
+    return(coefficients(fit)[2])
+}
+
+## static:
+df_dat <- wb_dat |> 
+    filter(!is.na(rptinc992j_p0p100), !is.na(rhweal992j_p0p100)) |> 
+    #filter(country_code == "AFG") |> select(ghweal992j_p0p100, year)
+    select(country_name, country_code, year,
+           emissions = 'CO2 emissions (metric tons per capita)', gini_in = gptinc992j_p0p100,
[…]

### 2025-05-19 ordination.R

@@ -192,3 +192,68 @@ a + b + plot_layout(guides = "collect") &
 
 
 ## Improve the graph, use nice names
+
+#### Animation ####
+# new UN dataset:
+fls <- fs::dir_ls("data/sdgs/")
+##list of countries
+un_countries <- read_csv2(file = "~/Documents/Projects/DATA/SDGs_UNStats/UNSD — Methodology.csv") |> 
+    janitor::clean_names()
+sdg <- list()
+
+for (i in seq_along(fls)){
+    load(fls[i])
+    sdg[[i]] <- d
+}
+
+
+sdg <- sdg |> bind_rows() |>
+    unnest(c(goal, target, indicator))
+
+sdg <- sdg |>
+    left_join(
+        select(un_countries, country = country_or_area, starts_with("iso") ),
+        by = c("geoAreaName" = "country")) |>
+    filter(value != "-")|>
+    mutate(value = as.numeric(value))
+
+## For visualization use all the data.
+tic()
+sdg <- sdg |>
+    filter(valueType != "String") |>
+    mutate(value = as.numeric(value)) |> filter(!is.na(value)) |>
+    filter(geoAreaName %in% df_countries$geoAreaName) #|>
+# group_by(goal, target, indicator, series, geoAreaName, timePeriodStart ) |>
+# summarize(value = mean(value, na.rm = TRUE))
+toc() # 3s
+
+## Dont expand: this recover implicit missing values
+tic()
+sdg <- right_join(
+    sdg |> select(-c(time_detail:dimensions)),
+    sdg |>
+        ungroup() |>
+        expand(series, geoAreaName, timePeriodStart))
+toc() # 6s
+
+tic()
+sdg |> ungroup() |> 
+    # filter(geoAreaName %in% c("Colombia", "Peru", "Ecuador", "Sweden", "South Africa", 
+    #                           "France", "United States of America", "China", "Cameroon",
+    #                           "Japan")) |> #, "Peru", "Ecuador"
+    ggplot(aes(timePeriodStart, series)) +
+    geom_tile(aes(fill = value)) +
+    geom_text(aes(label = geoAreaName), x = 1975, y = 600, color = "white", size = 8) +
+    scale_fill_viridis_c(na.value = "orange") +
+    labs(title = "Sustainable Development Goals",
+         subtitle = "Everything orange are missing values", caption = "Data source: UN SDGs API",
[…]
